### 2014-03-21 manuscript.tex
Change the template for Geophysics

\documentclass[manuscript]{geophysics}
\usepackage[nomarkers]{endfloat}
% For a two-column format that looks like the final product, uncomment bellow
% and comment the two lines above
%\documentclass[paper,twocolumn,twoside]{geophysics}

\usepackage{graphicx}

\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\mat}[1]{\mathbf{#1}}
\newcommand{\comp}[1]{#1^{\alpha\beta}}
\newcommand{\norm}[1]{\left|\left|#1\right|\right|}

\begin{document}

\title{
    Tesseroids:
    command-line tools for
    gravity field forward modeling
    in spherical coordinates
}

% manuscript number
\ms{}

\address{
    \footnotemark[1]
    Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil,
    e-mail: \\
    \footnotemark[2]
    Observat\'orio Nacional, Rio de Janeiro, Brazil,
    e-mail:
}

\author{Leonardo Uieda\footnotemark[1]\footnotemark[2],
        Vanderlei C. Oliveira Jr\footnotemark[2],
        and
        Val\'eria C. F. Barbosa\footnotemark[2]}

\lefthead{Uieda et al.}
\righthead{Gravity forward modeling with Tesseroids}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
pellentesque, quis molestie sem vestibulum.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Citing articles in text \citet{Asgharzadeh2007} and with parenthesis
\citep{Braitenberg2011}.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
pellentesque, quis molestie sem vestibulum.

Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
cursus ligula.

Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
non, pharetra varius justo.

Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
at pellentesque nisi lobortis a.

Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
tempus nibh pulvinar, ornare risus nec, suscipit eros.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
pellentesque, quis molestie sem vestibulum.

\begin{equation}
    \vect{d}^{\alpha\beta} = \sum\limits_{j=1}^{M} p_j \vect{a}_j^{\alpha\beta}.
    \label{eq:dalphabeta_sum}
\end{equation}

Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
cursus ligula.

Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
non, pharetra varius justo.

Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
at pellentesque nisi lobortis a.

Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
tempus nibh pulvinar, ornare risus nec, suscipit eros.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

%\begin{figure}
    %\centering
    %\subfigure[]{
        %\includegraphics[width=0.4\textwidth]{figs/sub1.png}\label{fig:sub1}}
    %\subfigure[]{
        %\includegraphics[width=0.4\textwidth]{figs/sub2.png}\label{fig:sub2}}
  %\caption{This is a figure caption.}
  %\label{fig:composite}
%\end{figure}

%This Figure \ref{fig:composite} is made up of Figure \ref{fig:sub1} and
%\ref{fig:sub2}.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
pellentesque, quis molestie sem vestibulum.

Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
cursus ligula.

Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
non, pharetra varius justo.

Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
at pellentesque nisi lobortis a.

Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
tempus nibh pulvinar, ornare risus nec, suscipit eros.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
pellentesque, quis molestie sem vestibulum.

Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
cursus ligula.

Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
non, pharetra varius justo.

Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
at pellentesque nisi lobortis a.

Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
tempus nibh pulvinar, ornare risus nec, suscipit eros.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
pellentesque, quis molestie sem vestibulum.

Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
cursus ligula.

Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
non, pharetra varius justo.

Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
at pellentesque nisi lobortis a.

Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
tempus nibh pulvinar, ornare risus nec, suscipit eros.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
pellentesque, quis molestie sem vestibulum.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{seg}
\bibliography{references}

\end{document}

### 2014-03-26 manuscript.tex
Added some refs and formula to text

@@
-\documentclass[manuscript]{geophysics}
-\usepackage[nomarkers]{endfloat}
+%\documentclass[manuscript]{geophysics}
+%\usepackage[nomarkers]{endfloat}
 % For a two-column format that looks like the final product, uncomment bellow
 % and comment the two lines above
-%\documentclass[paper,twocolumn,twoside]{geophysics}
+\documentclass[paper,twocolumn,twoside]{geophysics}
 
 \usepackage{graphicx}
+\usepackage{amsmath}
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
-et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
-tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
-Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
-eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
-fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
-non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
-consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
-convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
-nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
-pellentesque, quis molestie sem vestibulum.
-
-Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
-primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
-leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
-eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
-nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
-gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
-Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
-ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
-cursus ligula.
-
-Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
-posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
-ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
-consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
-molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
-Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
-facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
-non, pharetra varius justo.
-
-Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
-orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
-dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
-congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
-ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
-Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
-libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
-vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
-pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
-pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
-facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
-at pellentesque nisi lobortis a.
-
-Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
-Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
-suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
-elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
-hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
-gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
-amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
-tempus nibh pulvinar, ornare risus nec, suscipit eros.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Formula by \citet{Grombein2013}.
+
+First appears in \citet{Asgharzadeh2007}.
+
+Studied by \citet{Wild-Pfeiffer2008}.
+
+Numerical accuracy by \citet{Ku1977}.
+
+\subsection{Software implementation}
+
 Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
-et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
-tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
-Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
-eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
-fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
-non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
-consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
-convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
-nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
-pellentesque, quis molestie sem vestibulum.
+
+\citet{Barrera-Figueroa2006}.
+
+\subsection{Comparison with a spherical half-shell}
+
+The potential of a spherical half-shell is
 
 \begin{equation}
-    \vect{d}^{\alpha\beta} = \sum\limits_{j=1}^{M} p_j \vect{a}_j^{\alpha\beta}.
-    \label{eq:dalphabeta_sum}
+    V(r) = 2\pi G \rho \left[ \dfrac{l^3 + {r'}^3}{3r} - 0.5 {r'}^2 \right]
+           \Biggr \rvert_{r'=r_1}^{r'=r_2}
+    \label{eq:halfshell-pot}
 \end{equation}
 
-Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
-primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
-leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
-eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
-nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
-gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
-Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
-ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
-cursus ligula.
+\noindent
+where $\rho$ is the density,
+$l = \sqrt{r^2 + {r'}^2}$,
+$r$ is the radial coordinate of the observation point,
+and
+$r_1$ and $r_2$ are the bottom and top of the spherical shell,
+respectively.
 
-Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
-posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
-ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
-consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
-molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
-Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
-facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
-non, pharetra varius justo.
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/spherical-shell.png}
+    \caption{This is a figure caption.}
+    \label{fig:shell}
+\end{figure}
 
-Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
-orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
-dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
-congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
-ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
-Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
-libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
-vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
-pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
-pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
-facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
-at pellentesque nisi lobortis a.
+This is a reference to Figure~\ref{fig:shell}.
 
-Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
-Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
-suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
-elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
-hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
-gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
-amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
-tempus nibh pulvinar, ornare risus nec, suscipit eros.
+The $g_z$ component of the gravitational attraction for this shell is the
+radial derivative of the potential.
+
+The sign should be inverted because our convention is that z points down
+(toward the center of the Earth)
+
+\begin{equation}
+    g_z(r) = -\dfrac{\partial V}{\partial r} = 2\pi G \rho \left[ \dfrac{l^3 +
+            {r'}^3}{3r^2} - l \right] \Biggr \rvert_{r'=r_1}^{r'=r_2}
+    \label{eq:halfshell-gz}
+\end{equation}
+
+The gravity gradient tensor are the second derivatives of $V$:
+
+\begin{equation}
+    g_{xx}(r) = g_{yy}(r) = -\dfrac{1}{2} g_{zz}
+\end{equation}
+
+\begin{equation}
+    g_{xy}(r) = g_{xz}(r) = g_{yz}(r) = 0
+\end{equation}
+
+\begin{equation}
+    g_{zz}(r) = 2\pi G \rho \left[ 2\dfrac{l^3 + {r'}^3}{3r^3} - \dfrac{l}{r} +
+    \dfrac{r}{l} \right] \Biggr \rvert_{r'=r_1}^{r'=r_2}
+\end{equation}
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
-et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
-tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
-Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
-eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
-fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
-non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
-consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
-convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
-nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
-pellentesque, quis molestie sem vestibulum.
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/error-20deg}
+    \caption{This is a figure caption.}
+    \label{fig:error20}
+\end{figure}
 
-Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
-primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
-leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
-eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
-nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
-gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
-Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
-ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
-cursus ligula.
-
-Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
-posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
-ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
-consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
-molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
-Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
-facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
-non, pharetra varius justo.
-
-Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
-orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
-dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
-congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
-ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
-Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
-libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
-vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
-pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
-pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
-facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
-at pellentesque nisi lobortis a.
-
-Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
-Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
-suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
-elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
-hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
-gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
-amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
-tempus nibh pulvinar, ornare risus nec, suscipit eros.
+This is a reference to Figure \ref{fig:error20}.
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
-et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
-tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
-Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
-eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
-fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
-non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
-consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
-convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
-nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
-pellentesque, quis molestie sem vestibulum.
-
-Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
-primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
-leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
-eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
-nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
-gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
-Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
-ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
-cursus ligula.
-
-Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
-posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
-ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
-consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
-molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
-Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
-facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
-non, pharetra varius justo.
-
-Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
-orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
-dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
-congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
-ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
-Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
-libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
-vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
-pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
-pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
-facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
-at pellentesque nisi lobortis a.
-
-Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
-Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
-suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
-elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
-hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
-gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
-amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
-tempus nibh pulvinar, ornare risus nec, suscipit eros.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
-et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
-tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
-Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
-eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
-fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
-non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
-consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
-convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
-nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
-pellentesque, quis molestie sem vestibulum.
-
-Ut quis justo non leo malesuada molestie a ut eros. Vestibulum ante ipsum
-primis in faucibus orci luctus et ultrices posuere cubilia Curae; Morbi eros
-leo, elementum in dictum et, vulputate vel enim. Mauris pharetra lacinia augue
-eget placerat. Sed id massa ut massa blandit convallis. Quisque vel dolor eu
-nulla volutpat imperdiet. Nunc bibendum sem id elit accumsan, ac luctus leo
-gravida. Proin cursus justo sem, vulputate hendrerit libero pulvinar in.
-Suspendisse auctor sem at lorem pretium gravida. Vivamus aliquet cursus metus,
-ac pellentesque nunc auctor vel. Cras quis nibh ullamcorper, vestibulum est et,
-cursus ligula.
-
-Nunc euismod nunc nec lectus tempus, ultricies gravida quam rhoncus. Sed
-posuere vestibulum nulla, ut accumsan justo venenatis a. Etiam consequat in leo
-ac posuere. Fusce mattis ac sem non tempus. Aenean sollicitudin massa eu velit
-consequat, sed congue metus tincidunt. Pellentesque posuere, urna a consequat
-molestie, turpis libero rhoncus mi, lobortis accumsan nulla lectus at nunc.
-Suspendisse fermentum augue vel ornare cursus. Suspendisse sit amet eros
-facilisis, placerat nunc et, dapibus metus. Proin dolor enim, tempus at orci
-non, pharetra varius justo.
-
-Maecenas eget turpis vitae lorem pharetra venenatis in sed eros. Pellentesque
-orci lorem, egestas eu metus nec, hendrerit condimentum diam. Morbi quam neque,
-dictum ac erat a, lacinia facilisis risus. Phasellus ligula quam, suscipit eu
-congue ut, condimentum id leo. Donec convallis libero et aliquam mattis. Lorem
-ipsum dolor sit amet, consectetur adipiscing elit. Morbi sed vestibulum neque.
-Praesent nec nunc pellentesque, suscipit mi a, congue magna. Pellentesque et
-libero vitae velit aliquet congue eget id lorem. Etiam dolor sapien, ultrices
-vel neque vel, ultrices tempus quam. Nulla ultricies vehicula sagittis. Nunc
-pulvinar eros lacus, sit amet viverra enim ultricies sed. Ut sagittis
-pellentesque tristique. Proin facilisis tempor enim eu iaculis. In odio nulla,
-facilisis non aliquet nec, facilisis non sapien. Sed auctor tincidunt mauris,
-at pellentesque nisi lobortis a.
-
-Integer ullamcorper ipsum in metus lobortis, eget venenatis tellus adipiscing.
-Integer mi nulla, accumsan nec auctor vitae, tempus in erat. In at nisl
-suscipit, malesuada nulla ac, suscipit erat. Maecenas placerat neque eget lacus
-elementum dictum. Praesent lobortis nisi at arcu vulputate dapibus. Ut quis
-hendrerit magna. Donec ut fringilla elit. Suspendisse varius, eros eget aliquet
-gravida, purus justo hendrerit est, in fermentum nisl augue a ligula. Nulla sit
-amet magna risus. Vestibulum semper dolor et mauris sodales pretium. Nullam
-tempus nibh pulvinar, ornare risus nec, suscipit eros.
 

### 2014-03-28 manuscript.tex
Better shell fig and figs with tess splitting
Made a figure with examples of splitting a tesseroid using different
ratios.

In shell figure, removed parallels and meridians and made whole shell
discretized into tesseroids.

Using relative paths for both figures in Inkscape.
Added figure of tesseroid in coord system
Also added a caption to the split figure.
Caption for tesseroid figure

@@
 Numerical accuracy by \citet{Ku1977}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/tesseroid}
+    \caption{
+        View of a tesseroid,
+        the integration point $Q$,
+        a geocentric coordinate system $(X, Y, Z)$,
+        the computation $P$ and it's local coordinate system $(x, y, z)$.
+        $r$, $\phi$, $\lambda$ are
+        the radius, latitude, and longitude, respectively, of point $P$.
+    }
+    \label{fig:tesseroid}
+\end{figure}
+
+This is a reference to Figure~\ref{fig:tesseroid}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/tesseroid-split.png}
+    \caption{
+        Adaptive discretization for a computation point P
+        of the tesseroid shown in (a) using distance/size ratios of
+        (b) 0.5, (c) 1, and (d) 2.
+        $d$ is the distance between the tesseroid and P.
+    }
+    \label{fig:split}
+\end{figure}
+
+This is a reference to Figure~\ref{fig:split}.
 

### 2014-03-28 manuscript.tex
Added formula by Grombein

@@
 Numerical accuracy by \citet{Ku1977}.
+
+The gravitational potential of a tesseroid is
+
+\begin{equation}
+    V(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        \frac{1}{\ell} \kappa  dr' d\phi' d\lambda'
+\end{equation}
+
+The gravitational attraction
+can be calculated using the formula
+\citep{Grombein2013}:
+
+\begin{equation}
+    g_{\alpha}(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        \frac{\Delta_{\alpha}}{\ell^3} \kappa dr' d\phi' d\lambda'
+        \ \ \ \alpha \in \{x,y,z\}
+\end{equation}
+
+The gravity gradients can be calculated
+using the general formula
+\citep{Grombein2013}:
+
+\begin{multline}
+    g_{\alpha\beta}(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        I_{\alpha\beta}({r'}, {\phi'}, {\lambda'})
+        dr' d\phi' d\lambda'
+        \\
+        \alpha,\beta \in \{x,y,z\}
+\end{multline}
+
+\begin{equation}
+    I_{\alpha\beta}({r'}, {\phi'}, {\lambda'}) =
+        \left(
+            \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
+            \frac{\delta_{\alpha\beta}}{\ell^3}
+        \right)
+        \kappa \ \ \ \alpha,\beta \in \{x,y,z\}
+\end{equation}
+
+where $\rho$ is density,
+$\{x, y, z\}$ correspond to the local coordinate system
+of the computation point P,
+$\delta_{\alpha\beta}$ is the Kronecker delta, and
+
+\begin{eqnarray}
+    \Delta_x &=& r' K_{\phi} \\
+    \Delta_y &=& r' \cos \phi' \sin(\lambda' - \lambda) \\
+    \Delta_z &=& r' \cos \psi - r\\
+    \ell &=& \sqrt{r'^2 + r^2 - 2 r' r \cos \psi} \\
+    \cos\psi &=& \sin\phi\sin\phi' + \cos\phi\cos\phi'
+                 \cos(\lambda' - \lambda) \\
+    K_{\phi} &=& \cos\phi\sin\phi' - \sin\phi\cos\phi'
+                 \cos(\lambda' - \lambda)\\
+    \kappa &=& {r'}^2 \cos \phi'
+\end{eqnarray}
+
+$\phi$ is latitude,
+$\lambda$ is longitude, and
+$r$ is radius.
 

### 2014-03-28 manuscript.tex
Traded pngs for eps figures

@@
 Numerical accuracy by \citet{Ku1977}.
-
+\\
 The gravitational potential of a tesseroid is
@@
 This is a reference to Figure~\ref{fig:tesseroid}.
+\\
+The above integrals are solved using the Gauss-Legendre Quadrature rule
+\citep{Asgharzadeh2007}
+
+\begin{multline}
+    g_{\alpha\beta}(r,\phi,\lambda) \approx G \rho
+        \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8}
+        \\
+        \sum\limits_{k=1}^{N^{\lambda}}
+        \sum\limits_{j=1}^{N^{\phi}}
+        \sum\limits_{i=1}^{N^r}
+        W^r_i W^{\phi}_j W^{\lambda}_k
+        I_{\alpha\beta}({r'}_i, {\phi'}_j, {\lambda'}_k )
+        \\
+        \alpha,\beta \in \{1,2,3\}
+\end{multline}
+
+where $W_i^r$, $W_j^{\phi}$, and $W_k^{\lambda}$
+are weighting coefficients
+and $N^r$, $N^{\phi}$, and $N^{\lambda}$
+are the number of quadrature nodes
+(i.e., the order of the quadrature),
+for the radius, latitude, and longitude, respectively.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid-split.png}
+    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
     \caption{
@@
     \centering
-    \includegraphics[width=\columnwidth]{figs/spherical-shell.png}
+    \includegraphics[width=\columnwidth]{figs/spherical-shell}
     \caption{This is a figure caption.}

### 2014-04-01 manuscript.tex
Writing methods until adaptive discretization part
Added Hildebrand reference for the GLQ and edited equations.

--- 2014-03-28 manuscript.tex
@@
     Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil,
-    e-mail: \\
+    e-mail: leouieda@gmail.com;\\
     \footnotemark[2]
-    Observat\'orio Nacional, Rio de Janeiro, Brazil,
-    e-mail:
+    Observat\'orio Nacional, Rio de Janeiro, Brazil;
 }
@@
 
-Formula by \citet{Grombein2013}.
+The gravitational potential,
+gravitational acceleration,
+and Marussi (gravity gradient) tensor
+caused by a homogeneous tesseroid
+with density $\rho$
+are \citep{Grombein2013}
 
-First appears in \citet{Asgharzadeh2007}.
+\begin{equation}
+    V(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        \frac{1}{\ell} \kappa  dr' d\phi' d\lambda',
+    \label{eq:tesspot}
+\end{equation}
+\begin{equation}
+    g_{\alpha}(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        \frac{\Delta_{\alpha}}{\ell^3} \kappa dr' d\phi' d\lambda'
+        \ \ \ \alpha \in \{x,y,z\},
+    \label{eq:tessgrav}
+\end{equation}
 
-Studied by \citet{Wild-Pfeiffer2008}.
+\noindent
+and
 
-Numerical accuracy by \citet{Ku1977}.
+\begin{equation}
+    g_{\alpha\beta}(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        I_{\alpha\beta}
+        dr' d\phi' d\lambda',
+    \label{eq:tesstensor}
+\end{equation}
+
+\noindent
+in which
+$(\phi, \lambda, r)$ are
+the latitude, longitude, and radius
+coordinates of computation point P (Figure~\ref{fig:tesseroid}),
+$\alpha,\beta \in \{x,y,z\}$
+(the local coordinate system of point P),
+and
+
+\begin{eqnarray}
+    I_{\alpha\beta} &=&
+        \left(
+            \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
+            \frac{\delta_{\alpha\beta}}{\ell^3}
+        \right) \kappa, \\
+    \Delta_x &=& r' K_{\phi} , \\
+    \Delta_y &=& r' \cos \phi' \sin(\lambda' - \lambda) , \\
+    \Delta_z &=& r' \cos \psi - r, \\
+    \ell &=& \sqrt{r'^2 + r^2 - 2 r' r \cos \psi} , \\
+    \cos\psi &=& \sin\phi\sin\phi' + \cos\phi\cos\phi'
+                 \cos(\lambda' - \lambda) , \\
+    K_{\phi} &=& \cos\phi\sin\phi' - \sin\phi\cos\phi'
+                 \cos(\lambda' - \lambda), \\
+    \kappa &=& {r'}^2 \cos \phi'.
+\end{eqnarray}
+
+\noindent
+$\delta_{\alpha\beta}$ is the Kronecker delta,
+i.e. $\delta_{\alpha\beta}=1$ if $\alpha=\beta$
+and $\delta_{\alpha\beta}=0$ otherwise.
 
@@
         $r$, $\phi$, $\lambda$ are
-        the radius, latitude, and longitude, respectively, of point $P$.
+        the radius, latitude, and longitude, respectively, of point $P$,
+        and $\ell$ is the Cartesian distance between $P$ and $Q$.
     }
@@
 
-This is a reference to Figure~\ref{fig:tesseroid}.
+The above integrals
+have to be solved numerically
+\citep{Wild-Pfeiffer2008}
+and different approaches have been proposed.
+\citet{Heck2007}
+used a Taylor series expansion,
+while \citet{Asgharzadeh2007}
+used the Gauss-Legendre Quadrature (GLQ).
+\citet{Wild-Pfeiffer2008} investigated
+the use of both methods,
+along with alternative mass elements,
+and found the GLQ to be the most efficient method.
+Therefore, we have opted to use
+the 3D GLQ in our computations.
+
+The Gauss-Legendre Quadrature
+approximates the integral by
+a weighted sum
+\citep{Hildebrand1987},
+e.g.
+
+\begin{equation}
+    \int\limits_a^b f(x) dx \approx
+    \frac{b-a}{2}\sum\limits_{i=1}^N W_i f(x_i),
+\end{equation}
+
+\noindent
+where the discretization points (nodes) $x_i$
+are the roots of the $N$th order Legendre polynomial $P_N$
+and the weights $W_i$ are \citep{Hildebrand1987}
+
+\begin{equation}
+    W_i = \frac{2}{(1 - x_i^2)(P'_N(x_i))^2}.
+\end{equation}
+
+The triple integrals in equations
+\ref{eq:tesspot},
+\ref{eq:tessgrav},
+and
+\ref{eq:tesstensor}
+become
+
+\begin{equation}
+    V(r,\phi,\lambda) \approx
+        A
+        \sum\limits_{k=1}^{N^{\lambda}}
+        \sum\limits_{j=1}^{N^{\phi}}
+        \sum\limits_{i=1}^{N^r}
+        W^r_i W^{\phi}_j W^{\lambda}_k
+        \frac{1}{\ell} \kappa,
+\end{equation}
+\begin{equation}
+    g_{\alpha}(r,\phi,\lambda) \approx
+        A
+        \sum\limits_{k=1}^{N^{\lambda}}
+        \sum\limits_{j=1}^{N^{\phi}}
+        \sum\limits_{i=1}^{N^r}
+        W^r_i W^{\phi}_j W^{\lambda}_k
+        \frac{\Delta_{\alpha}}{\ell^3} \kappa,
+\end{equation}
+
+\noindent
+and
+
+\begin{equation}
+    g_{\alpha\beta}(r,\phi,\lambda) \approx
+        A
+        \sum\limits_{k=1}^{N^{\lambda}}
+        \sum\limits_{j=1}^{N^{\phi}}
+        \sum\limits_{i=1}^{N^r}
+        W^r_i W^{\phi}_j W^{\lambda}_k
+        I_{\alpha\beta},
+\end{equation}
+
+\noindent
+where
+
+\begin{equation}
+    A = G \rho
+    \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8},
+\end{equation}
+
+\noindent
+$W_i^r$, $W_j^{\phi}$, and $W_k^{\lambda}$
+are the weights
+and $N^r$, $N^{\phi}$, and $N^{\lambda}$
+are the number of nodes
+for the radial, latitudinal, and longitudinal dimensions, respectively.
+
+The accuracy of the numerical integration
+using the GLQ
+was investigated by \citet{Ku1977}
+for the case of the right-rectangular prism.
+\citet{Ku1977} found that the accuracy
+depends on the ratio between
+the average distance between GLQ nodes
+and the distance to the observation point P.
+As a rule-of-thumb,
+\citet{Ku1977} suggests that
+the distance between nodes
+should not be greater than
+the distance to point P.
+\citet{Li2011} devised a scheme that
+recursively subdivides the tesseroids
+into smaller ones,
+effectively decreasing the distance between nodes.
+This scheme guarantees that
+the rule-of-thumb is not violated.
+Hence,
+optimal accuracy of the numerical integration
+is achieved automatically.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid-split.png}
+    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
     \caption{
-        Adaptive discretization for a computation point P
-        of the tesseroid shown in (a) using distance/size ratios of
+        Adaptive discretization
+        of the tesseroid shown in (a)
+        for a computation point P
+        using distance/size ratios of
         (b) 0.5, (c) 1, and (d) 2.
@@
 \end{figure}
-
-This is a reference to Figure~\ref{fig:split}.
 

### 2014-04-01 manuscript.tex
Sect about shell comparison comes before software
Finished the theory section before and started on the shell comparison
section.

@@
 
+We use a modified version
+of the recursive scheme of \citet{Li2011}.
+We fix the number of nodes in the GLQ
+$N^\lambda=N^\phi=N^r=2$.
+In this case, the maximum distance between nodes
+is approximately the largest dimension of the tesseroid ($L$)
+\citep{Wild-Pfeiffer2008}.
+For a computation point $P$
+at a distance $d$ from
+the center of the top face of the tesseroid
+(Figure~\ref{fig:ratio}a),
+the tesseroid is divided in eight parts if
+
+\begin{equation}
+    \frac{d}{L} < R,
+    \label{eq:ratio}
+\end{equation}
+
+\noindent
+where $R$ is the minimum distance/size ratio desired.
+In case the tesseroid is divided,
+the inequality in equation \ref{eq:ratio}
+is checked for each of the eight parts.
+This process is repeated recursively
+until no divisions are required,
+ensuring that equation \ref{eq:ratio}
+holds for all tesseroids created.
+The gravitational effect
+of the original tesseroid
+is then calculated as
+the sum of the effect
+of the smaller tesseroids.
+
+The rule-of-thumb of \citet{Ku1977}
+used by \citet{Li2011}
+is equivalent to setting $R=1$.
+Figure~\ref{fig:ratio}
+illustrates the subdivision of a tesseroid
+(Figure~\ref{fig:ratio}a)
+for computation point $P$
+using increasing values of $R$.
+Note that increasing $R$
+results is a finer division of the tesseroid
+around the computation point
+and thus a more accurate GLQ integration.
+However, higher values of $R$
+also result in
+larger computation times.
+For example,
+the tesseroid in Figure~\ref{fig:ratio}a
+is replaced by
+50 tesseroids for $R=0.5$ (Figure~\ref{fig:ratio}b),
+344 tesseroids for $R=1$ (Figure~\ref{fig:ratio}c),
+and 2472 tesseroids for  $R=2$ (Figure~\ref{fig:ratio}d).
+
 \begin{figure}
@@
         for a computation point P
-        using distance/size ratios of
+        using $R$ equal to
         (b) 0.5, (c) 1, and (d) 2.
-        $d$ is the distance between the tesseroid and P.
+        $d$ is the distance between
+        the center of the top face of the tesseroid
+        and $P$.
+        $L$ is the largest dimension of the tesseroid.
     }
-    \label{fig:split}
+    \label{fig:ratio}
 \end{figure}
 
-\subsection{Software implementation}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-
-\citet{Barrera-Figueroa2006}.
+Clearly,
+there is need to determine
+an optimal value of $R$
+that yields the desired accuracy
+in the GLQ integration.
+In the following section,
+we compare the computed gravitational effects
+of a tesseroid model
+with that of a spherical half-shell
+(Figure~\ref{fig:shell}),
+for which exist analytical expressions.
 
@@
 
-The potential of a spherical half-shell is
+The gravitational potential
+of a homogeneous spherical half-shell
+at a computation point P
+located at a height $h$
+from the center of the shells top face
+(Figure~\ref{fig:shell})
+is
 
@@
     V(r) = 2\pi G \rho \left[ \dfrac{l^3 + {r'}^3}{3r} - 0.5 {r'}^2 \right]
-           \Biggr \rvert_{r'=r_1}^{r'=r_2}
+           \Biggr \rvert_{r'=r_1}^{r'=r_2},
     \label{eq:halfshell-pot}
@@
 \end{equation}
+
+\subsection{Software implementation}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
+
+\citet{Barrera-Figueroa2006}.
 

### 2014-04-01 manuscript.tex
Section on comparison with shell
Finished most of it and updated the figure to include the height of the
computation point.

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methodology}
@@
 
-The gravitational potential
-of a homogeneous spherical half-shell
-at a computation point P
-located at a height $h$
-from the center of the shells top face
-(Figure~\ref{fig:shell})
-is
+The gravitational potential,
+gravitational attraction,
+and
+Marussi (gravity gradient) tensor
+caused by the homogeneous spherical half-shell
+shown in Figure~\ref{fig:shell}
+at a computation point
+$P = (\lambda=90^\circ, \phi=0^\circ, r=r_2+h)$
+are
 
 \begin{equation}
-    V(r) = 2\pi G \rho \left[ \dfrac{l^3 + {r'}^3}{3r} - 0.5 {r'}^2 \right]
+    V(h) = 2\pi G \rho \left[
+        \dfrac{l^3 + {r'}^3}{3(r_2 + h)} - 0.5 {r'}^2 \right]
          \Biggr \rvert_{r'=r_1}^{r'=r_2},
-    \label{eq:halfshell-pot}
+    \label{eq:shellpot}
+\end{equation}
+\begin{equation}
+    g_z(h) = 2\pi G \rho \left[
+        \dfrac{l^3 + {r'}^3}{3(r_2 + h)^2} - l \right]
+        \Biggr \rvert_{r'=r_1}^{r'=r_2},
+    \label{eq:shellgz}
+\end{equation}
+\begin{equation}
+    g_{xx}(h) = g_{yy}(h) = -\dfrac{1}{2} g_{zz}(h),
+\end{equation}
+\begin{equation}
+    g_{xy}(h) = g_{xz}(h) = g_{yz}(h) = 0,
+\end{equation}
+
+\noindent
+and
+
+\begin{equation}
+    g_{zz}(h) = 2\pi G \rho \left[
+        2\dfrac{l^3 + {r'}^3}{3(r_2 + h)^3}
+        - \dfrac{l}{r_2 + h} + \dfrac{r_2 + h}{l}
+        \right]
+        \Biggr \rvert_{r'=r_1}^{r'=r_2},
+    \label{eq:shellgzz}
 \end{equation}
@@
 where $\rho$ is the density,
-$l = \sqrt{r^2 + {r'}^2}$,
-$r$ is the radial coordinate of the observation point,
+$l = \sqrt{(r_2 + h)^2 + {r'}^2}$,
 and
-$r_1$ and $r_2$ are the bottom and top of the spherical shell,
+$r_1$ and $r_2$ are radial coordinates of
+the bottom and top of the spherical shell,
 respectively.
@@
     \includegraphics[width=\columnwidth]{figs/spherical-shell}
-    \caption{This is a figure caption.}
+    \caption{A spherical half-shell
+        spanning the ranges $0^{\circ} \le \lambda \le 180^\circ$,
+        $-90^\circ \le \phi \le 90^\circ$,
+        and $r_1 \le r \le r_2$,
+        the corresponding tesseroid model,
+        the computation point $P$
+        at $(\lambda=90^\circ,\phi=0^\circ,r=r_2+h)$,
+        and its local coordinate system $(x, y, z)$.
+    }
     \label{fig:shell}
@@
 
-This is a reference to Figure~\ref{fig:shell}.
-
-The $g_z$ component of the gravitational attraction for this shell is the
-radial derivative of the potential.
-
-The sign should be inverted because our convention is that z points down
-(toward the center of the Earth)
-
-\begin{equation}
-    g_z(r) = -\dfrac{\partial V}{\partial r} = 2\pi G \rho \left[ \dfrac{l^3 +
-            {r'}^3}{3r^2} - l \right] \Biggr \rvert_{r'=r_1}^{r'=r_2}
-    \label{eq:halfshell-gz}
-\end{equation}
-
-The gravity gradient tensor are the second derivatives of $V$:
-
-\begin{equation}
-    g_{xx}(r) = g_{yy}(r) = -\dfrac{1}{2} g_{zz}
-\end{equation}
-
-\begin{equation}
-    g_{xy}(r) = g_{xz}(r) = g_{yz}(r) = 0
-\end{equation}
-
-\begin{equation}
-    g_{zz}(r) = 2\pi G \rho \left[ 2\dfrac{l^3 + {r'}^3}{3r^3} - \dfrac{l}{r} +
-    \dfrac{r}{l} \right] \Biggr \rvert_{r'=r_1}^{r'=r_2}
-\end{equation}
+For our comparison,
+we used a spherical half-shell with
+$\rho=2.7\ g.cm^{-3}$,
+$r_1=6378.137\ km$ (mean Earth radius),
+and
+$r_2 = r_1 + 50\ km$.
+We discretized the half-shell
+into tesseroids of equal size
+and calculated the gravitational effects of
+the half-shell
+(using equations \ref{eq:shellpot}-\ref{eq:shellgzz})
+and the corresponding tesseroid model
+at various heights.
+These calculations were repeated
+for varying values of
+the distance/size ratio $R$.
+This analysis
+allows us to establish
+the optimal value of $R$
+that achieves a desired error level.
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-
 \citet{Barrera-Figueroa2006}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
-
-%\begin{figure}
-    %\centering
-    %\subfigure[]{
-        %\includegraphics[width=0.4\textwidth]{figs/sub1.png}\label{fig:sub1}}
-    %\subfigure[]{
-        %\includegraphics[width=0.4\textwidth]{figs/sub2.png}\label{fig:sub2}}
-  %\caption{This is a figure caption.}
-  %\label{fig:composite}
-%\end{figure}
-
-%This Figure \ref{fig:composite} is made up of Figure \ref{fig:sub1} and
-%\ref{fig:sub2}.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Acknowledgments}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{seg}

### 2014-04-01 manuscript.tex
Started implementation section

@@
 
+Data and source code to produce the results of this comparison
+are provided by \textbf{(cite notebook on figshare)}.
+
 \subsection{Software implementation}
 
-\citet{Barrera-Figueroa2006}.
+The methods described above
+were implemented the software package
+Tesseroids version 1.1.1.
+The package consists
+of a range of command-line programs
+written in the C programming language.
+Tesseroids is free and open-source software
+under the terms of
+the 3-clause BSD license.
+
+Each program performs a single type of calculation.
+Programs receive input from
+the standard input stream (\texttt{stdin})
+and command-line arguments.
+Output is made in plain text
+through the standard output stream (\texttt{stdout}).
+
+The main programs in Tesseroids are
+\texttt{tesspot},
+\texttt{tessgx},
+\texttt{tessgy},
+\texttt{tessgz},
+\texttt{tessgxx},
+\ldots ,
+\texttt{tessgzz}
+(henceforth referred to collectively as \texttt{tess*} programs).
+These programs
+calculate the gravitational potential,
+gravitational attraction,
+and Marussi tensor
+caused by a tesseroid model.
+The computation points are passed
+through the standard input stream
+and the tesseroid model
+is passed through
+a simple model file.
+All programs receive
+optional command-line arguments
+that control the GLQ order (default is $N^\lambda=N^\phi=N^r=2$)
+and distance/size ratio $R$
+for the recursive division scheme.
+
+The complete list of programs provided in Tesseroids is:
+
+\begin{itemize}
+    \item \texttt{tesspot}, \texttt{tessgx}, \texttt{tessgy}, \texttt{tessgz},
+        \texttt{tessgxx}, \texttt{tessgxy}, \texttt{tessgxz},
+        \texttt{tessgyy}, \texttt{tessgyz}, \texttt{tessgzz}:
+        Calculate the gravitational potential, gravitational attraction,
+        and Marussi tensor caused by a tesseroid model.
+    \item \texttt{prismpot},
+        \texttt{prismgx}, \texttt{prismgy}, \texttt{prismgz},
+        \texttt{prismgxx}, \texttt{prismgxy}, \texttt{prismgxz},
+        \texttt{prismgyy}, \texttt{prismgyz}, \texttt{prismgzz}:
+        Calculate the gravitational potential, gravitational attraction,
+        and Marussi tensor caused by a right-rectangular prism model
+        in Cartesian coordinates
+        using the formula of \citet{Nagy2000}.
+    \item \texttt{prismpots}, \texttt{prismgs}, \texttt{prismggts}:
+        Calculate the gravitational potential, gravitational attraction,
+        and Marussi tensor caused by a right-rectangular prism model
+        in spherical coordinates
+        using the formula of \citet{Wild-Pfeiffer2008}.
+    \item \texttt{tess2prism}: Converts a tesseroid model into
+        right-rectangular prisms of equal mass.
+    \item \texttt{tessgrd}: Generate regular grids.
+    \item \texttt{tessdefaults}: Output the default values
+        and values of physical constants used.
+    \item \texttt{tessmodgen}: Generate a tesseroid model
+        from a gridded relief (e.g., a digital elevation model).
+    \item \texttt{tesslayers}: Generate a tesseroid model of
+        a series of stacked layers.
+    \item \texttt{tessmass}: Calculate the total mass of a tesseroid model.
+\end{itemize}
+
+Tesseroids is freely available
+online\footnote{www.leouieda.com/tesseroids}.
+\textbf{(cite archive on figshare)} provides a persistent
+archive for the source code and compiled binaries of Tesseroids.
 

### 2014-08-20 manuscript.tex
Moved manuscript to GJI template

@@
-%\documentclass[manuscript]{geophysics}
-%\usepackage[nomarkers]{endfloat}
-% For a two-column format that looks like the final product, uncomment bellow
-% and comment the two lines above
-\documentclass[paper,twocolumn,twoside]{geophysics}
+%\documentclass[extra,mreferee]{gji}
+% Use the option below to make a nicely formatted manuscript
+\documentclass[extra]{gji}
+\usepackage{timet}
+\usepackage{graphicx}
 
-\usepackage{graphicx}
-\usepackage{amsmath}
-
-\newcommand{\vect}[1]{\mathbf{#1}}
-\newcommand{\mat}[1]{\mathbf{#1}}
-\newcommand{\comp}[1]{#1^{\alpha\beta}}
-\newcommand{\norm}[1]{\left|\left|#1\right|\right|}
+\title[Optimal forward modeling using tesseroids]{
+    Optimal forward modeling of gravitational fields
+    using tesseroids with adaptive discretization
+}
+\author[Uieda et al.]{
+    Leonardo Uieda$^{1,2}$,
+    Vanderlei C. Oliveira Jr$^2$,
+    Val\'eria C. F. Barbosa$^2$
+    \\
+    $^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil.
+    $^2$Observat\'orio Nacional, Rio de Janeiro, Brazil.
+}
 
@@
 
-\title{
-    Tesseroids:
-    command-line tools for
-    gravity field forward modeling
-    in spherical coordinates
-}
+\label{firstpage}
+\maketitle
 
-% manuscript number
-\ms{}
-
-\address{
-    \footnotemark[1]
-    Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil,
-    e-mail: leouieda@gmail.com;\\
-    \footnotemark[2]
-    Observat\'orio Nacional, Rio de Janeiro, Brazil;
-}
-
-\author{Leonardo Uieda\footnotemark[1]\footnotemark[2],
-        Vanderlei C. Oliveira Jr\footnotemark[2],
-        and
-        Val\'eria C. F. Barbosa\footnotemark[2]}
-
-\lefthead{Uieda et al.}
-\righthead{Gravity forward modeling with Tesseroids}
-
-\begin{abstract}
+\begin{summary}
 Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
@@
 pellentesque, quis molestie sem vestibulum.
-\end{abstract}
+\end{summary}
 
@@
 the latitude, longitude, and radius
-coordinates of computation point P (Figure~\ref{fig:tesseroid}),
+coordinates of computation point P (Fig.~\ref{fig:tesseroid}),
 $\alpha,\beta \in \{x,y,z\}$
@@
 the center of the top face of the tesseroid
-(Figure~\ref{fig:ratio}a),
+(Fig.~\ref{fig:ratio}a),
 the tesseroid is divided in eight parts if
@@
 is equivalent to setting $R=1$.
-Figure~\ref{fig:ratio}
+Fig.~\ref{fig:ratio}
 illustrates the subdivision of a tesseroid
-(Figure~\ref{fig:ratio}a)
+(Fig.~\ref{fig:ratio}a)
 for computation point $P$
@@
 For example,
-the tesseroid in Figure~\ref{fig:ratio}a
+the tesseroid in Fig.~\ref{fig:ratio}a
 is replaced by
-50 tesseroids for $R=0.5$ (Figure~\ref{fig:ratio}b),
-344 tesseroids for $R=1$ (Figure~\ref{fig:ratio}c),
-and 2472 tesseroids for  $R=2$ (Figure~\ref{fig:ratio}d).
+50 tesseroids for $R=0.5$ (Fig.~\ref{fig:ratio}b),
+344 tesseroids for $R=1$ (Fig.~\ref{fig:ratio}c),
+and 2472 tesseroids for  $R=2$ (Fig.~\ref{fig:ratio}d).
 
@@
 with that of a spherical half-shell
-(Figure~\ref{fig:shell}),
+(Fig.~\ref{fig:shell}),
 for which exist analytical expressions.
@@
 caused by the homogeneous spherical half-shell
-shown in Figure~\ref{fig:shell}
+shown in Fig.~\ref{fig:shell}
 at a computation point
@@
     V(h) = 2\pi G \rho \left[
-        \dfrac{l^3 + {r'}^3}{3(r_2 + h)} - 0.5 {r'}^2 \right]
-         \Biggr \rvert_{r'=r_1}^{r'=r_2},
+        \frac{l^3 + {r'}^3}{3(r_2 + h)} - 0.5 {r'}^2 \right]
+         \Biggr|_{r'=r_1}^{r'=r_2},
     \label{eq:shellpot}
@@
     g_z(h) = 2\pi G \rho \left[
-        \dfrac{l^3 + {r'}^3}{3(r_2 + h)^2} - l \right]
-        \Biggr \rvert_{r'=r_1}^{r'=r_2},
+        \frac{l^3 + {r'}^3}{3(r_2 + h)^2} - l \right]
+        \Biggr|_{r'=r_1}^{r'=r_2},
     \label{eq:shellgz}
@@
 \begin{equation}
-    g_{xx}(h) = g_{yy}(h) = -\dfrac{1}{2} g_{zz}(h),
+    g_{xx}(h) = g_{yy}(h) = -\frac{1}{2} g_{zz}(h),
 \end{equation}
@@
     g_{zz}(h) = 2\pi G \rho \left[
-        2\dfrac{l^3 + {r'}^3}{3(r_2 + h)^3}
-        - \dfrac{l}{r_2 + h} + \dfrac{r_2 + h}{l}
+        2\frac{l^3 + {r'}^3}{3(r_2 + h)^3}
+        - \frac{l}{r_2 + h} + \frac{r_2 + h}{l}
         \right]
-        \Biggr \rvert_{r'=r_1}^{r'=r_2},
+        \Biggr|_{r'=r_1}^{r'=r_2},
     \label{eq:shellgzz}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figs/error-20deg}
-    \caption{This is a figure caption.}
-    \label{fig:error20}
-\end{figure}
-
-This is a reference to Figure \ref{fig:error20}.
+Meh.
 
@@
 
-\bibliographystyle{seg}
+\bibliographystyle{gji}
 \bibliography{references}

### 2014-08-21 manuscript.tex
Updated split fig to split in each dimension
Better figure for tessXshell
Font size is still off and need a better color palette. It looks pretty
ugly as is.
Tweaks to tessXshell figure
Added a matplotlibrc file with the nice config for the plot.
More text in the methods
Moved some things to the introduction. Stopped before the division
algorithm  description.

@@
 \citep{Braitenberg2011}.
+
+The above integrals
+have to be solved numerically
+\citep{Wild-Pfeiffer2008}
+and different approaches have been proposed in the literature.
+\citet{Heck2007}
+used a Taylor series expansion
+of the integral kernels
+around the geometrical center of the tesseroid.
+\citet{Asgharzadeh2007}
+used the Gauss-Legendre Quadrature (GLQ)
+for the numerical integration.
+\citet{Wild-Pfeiffer2008} investigated
+the use of both methods,
+along with alternative mass elements,
+and found the GLQ to be the most efficient method.
+We have opted to use
+the 3D GLQ integration in our computations.
+
+The accuracy of the numerical integration
+using the GLQ
+was investigated by \citet{Ku1977}
+for the case of
+the gravitational attraction of
+a right-rectangular prism
+in Cartesian coordinates.
+\citet{Ku1977} states that the accuracy
+depends on the ratio between
+the average distance between GLQ nodes
+and the distance to the observation point P.
+As a rule-of-thumb,
+\citet{Ku1977} suggests that
+the distance between nodes
+should not be greater than
+the distance to point P.
+Extrapolating this rule-of-thumb
+to tesseroids,
+\citet{Li2011}
+devised a scheme that
+recursively divides a tesseroid
+into smaller ones when the rule is violated.
+This procedure
+effectively decreases the distance between nodes
+and, theoretically, guarantees the accuracy of the GLQ integration.
 
@@
 $\alpha,\beta \in \{x,y,z\}$
-(the local coordinate system of point P),
+are the local coordinate system of point P,
 and
@@
         View of a tesseroid,
-        the integration point $Q$,
+        the integration point $Q$ inside the tesseroid,
         a geocentric coordinate system $(X, Y, Z)$,
@@
 
-The above integrals
-have to be solved numerically
-\citep{Wild-Pfeiffer2008}
-and different approaches have been proposed.
-\citet{Heck2007}
-used a Taylor series expansion,
-while \citet{Asgharzadeh2007}
-used the Gauss-Legendre Quadrature (GLQ).
-\citet{Wild-Pfeiffer2008} investigated
-the use of both methods,
-along with alternative mass elements,
-and found the GLQ to be the most efficient method.
-Therefore, we have opted to use
-the 3D GLQ in our computations.
-
-The Gauss-Legendre Quadrature
+We have opted to use
+the 3D Gauss-Legendre Quadrature
+to numerically integrate
+\ref{eq:tesspot},
+\ref{eq:tessgrav},
+and
+\ref{eq:tesstensor}.
+The GLQ
 approximates the integral by
@@
 \noindent
-where the discretization points (nodes) $x_i$
+where
+$W_i$ are weights and
+the discretization points (nodes) $x_i$
 are the roots of the $N$th order Legendre polynomial $P_N$
-and the weights $W_i$ are \citep{Hildebrand1987}
+scaled to the integration limits.
+The roots of a second-order Legendre polynomial are
+$x_1=-0.577350269$ and $x_2=0.577350269$.
+For an arbitrary order $N$,
+we have used the root-finder algorithm
+of \citet{Barrera-Figueroa2006}
+to calculate the roots of $P_N$.
+The weights $W_i$ are \citep{Hildebrand1987}
 
 \begin{equation}
-    W_i = \frac{2}{(1 - x_i^2)(P'_N(x_i))^2}.
+    W_i = \frac{2}{(1 - x_i^2)(P'_N(x_i))^2},
 \end{equation}
+
+\noindent
+where $P'_N$ is the first derivative of $P_N$.
 
@@
 
-The accuracy of the numerical integration
-using the GLQ
-was investigated by \citet{Ku1977}
-for the case of the right-rectangular prism.
-\citet{Ku1977} found that the accuracy
-depends on the ratio between
-the average distance between GLQ nodes
-and the distance to the observation point P.
-As a rule-of-thumb,
-\citet{Ku1977} suggests that
-the distance between nodes
-should not be greater than
-the distance to point P.
-\citet{Li2011} devised a scheme that
-recursively subdivides the tesseroids
-into smaller ones,
-effectively decreasing the distance between nodes.
-This scheme guarantees that
-the rule-of-thumb is not violated.
-Hence,
-optimal accuracy of the numerical integration
-is achieved automatically.
-
-We use a modified version
+We have developed a modified version
 of the recursive scheme of \citet{Li2011}.
 We fix the number of nodes in the GLQ
-$N^\lambda=N^\phi=N^r=2$.
-In this case, the maximum distance between nodes
-is approximately the largest dimension of the tesseroid ($L$)
-\citep{Wild-Pfeiffer2008}.
-For a computation point $P$
-at a distance $d$ from
-the center of the top face of the tesseroid
-(Fig.~\ref{fig:ratio}a),
-the tesseroid is divided in eight parts if
+$N^\lambda=N^\phi=N^r=2$
+so that the distance between nodes
+in the radial, longitudinal, and latitudinal directions
+are approximately the dimensions of the tesseroid.
+We define the distance $d$ as
+the Cartesian distance between
+the computation point $P$
+and the center of the top face of the tesseroid
+(Fig.~\ref{fig:ratio}a).
+Thus, the rule-of-thumb of \citet{Ku1977}
+can be written as
 
 \begin{equation}
-    \frac{d}{L} < R,
+    \frac{d}{L} < r,
     \label{eq:ratio}
@@
 \noindent
-where $R$ is the minimum distance/size ratio desired.
+where
+$L$ is one of the dimensions of the tesseroid
+and
+$r$ is the distance/size ratio.
+For \citet{Ku1977} this ratio should be equal to one.
+The dimensions of the tesseroid are
+
+\begin{eqnarray*}
+    L_r &=& r_2 - r_1,\\
+    L_\phi &=& R(\phi_2 - \phi_1),\\
+    L_\lambda &=& R(\lambda_2 - \lambda_1),
+\end{eqnarray*}
+
+\noindent
+where $R$ is the mean Earth radius
+and the longitudes and latitudes are given in radians.
+
 In case the tesseroid is divided,
@@
 of the smaller tesseroids.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
+    \caption{
+        Adaptive discretization
+        of the tesseroid shown in (a)
+        for a computation point P
+        using $R$ equal to
+        (b) 1, (c) 2, and (d) 6.
+        $d$ is the distance between
+        the center of the top face of the tesseroid
+        and $P$.
+        $L$ is the largest dimension of the tesseroid.
+    }
+    \label{fig:ratio}
+\end{figure}
 
@@
 and 2472 tesseroids for  $R=2$ (Fig.~\ref{fig:ratio}d).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
-    \caption{
-        Adaptive discretization
-        of the tesseroid shown in (a)
-        for a computation point P
-        using $R$ equal to
-        (b) 0.5, (c) 1, and (d) 2.
-        $d$ is the distance between
-        the center of the top face of the tesseroid
-        and $P$.
-        $L$ is the largest dimension of the tesseroid.
-    }
-    \label{fig:ratio}
-\end{figure}
 
@@
 written in the C programming language.
-Tesseroids is free and open-source software
+It is freely available
+online\footnote{www.leouieda.com/tesseroids}
 under the terms of
-the 3-clause BSD license.
-
-Each program performs a single type of calculation.
-Programs receive input from
-the standard input stream (\texttt{stdin})
-and command-line arguments.
-Output is made in plain text
-through the standard output stream (\texttt{stdout}).
-
-The main programs in Tesseroids are
-\texttt{tesspot},
-\texttt{tessgx},
-\texttt{tessgy},
-\texttt{tessgz},
-\texttt{tessgxx},
-\ldots ,
-\texttt{tessgzz}
-(henceforth referred to collectively as \texttt{tess*} programs).
-These programs
-calculate the gravitational potential,
-gravitational attraction,
-and Marussi tensor
-caused by a tesseroid model.
-The computation points are passed
-through the standard input stream
-and the tesseroid model
-is passed through
-a simple model file.
-All programs receive
-optional command-line arguments
-that control the GLQ order (default is $N^\lambda=N^\phi=N^r=2$)
-and distance/size ratio $R$
-for the recursive division scheme.
-
-The complete list of programs provided in Tesseroids is:
-
-\begin{itemize}
-    \item \texttt{tesspot}, \texttt{tessgx}, \texttt{tessgy}, \texttt{tessgz},
-        \texttt{tessgxx}, \texttt{tessgxy}, \texttt{tessgxz},
-        \texttt{tessgyy}, \texttt{tessgyz}, \texttt{tessgzz}:
-        Calculate the gravitational potential, gravitational attraction,
-        and Marussi tensor caused by a tesseroid model.
-    \item \texttt{prismpot},
-        \texttt{prismgx}, \texttt{prismgy}, \texttt{prismgz},
-        \texttt{prismgxx}, \texttt{prismgxy}, \texttt{prismgxz},
-        \texttt{prismgyy}, \texttt{prismgyz}, \texttt{prismgzz}:
-        Calculate the gravitational potential, gravitational attraction,
-        and Marussi tensor caused by a right-rectangular prism model
-        in Cartesian coordinates
-        using the formula of \citet{Nagy2000}.
-    \item \texttt{prismpots}, \texttt{prismgs}, \texttt{prismggts}:
-        Calculate the gravitational potential, gravitational attraction,
-        and Marussi tensor caused by a right-rectangular prism model
-        in spherical coordinates
-        using the formula of \citet{Wild-Pfeiffer2008}.
-    \item \texttt{tess2prism}: Converts a tesseroid model into
-        right-rectangular prisms of equal mass.
-    \item \texttt{tessgrd}: Generate regular grids.
-    \item \texttt{tessdefaults}: Output the default values
-        and values of physical constants used.
-    \item \texttt{tessmodgen}: Generate a tesseroid model
-        from a gridded relief (e.g., a digital elevation model).
-    \item \texttt{tesslayers}: Generate a tesseroid model of
-        a series of stacked layers.
-    \item \texttt{tessmass}: Calculate the total mass of a tesseroid model.
-\end{itemize}
-
-Tesseroids is freely available
-online\footnote{www.leouieda.com/tesseroids}.
+the open-source BSD 3-clause license.
 \textbf{(cite archive on figshare)} provides a persistent
@@
 Meh.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figs/tesseroid-x-shell}
+    \caption{
+        Tesseroid vs spherical shell.
+    }
+    \label{fig:tesseroid-x-shell}
+\end{figure*}
 

### 2014-08-22 manuscript.tex
Will separate the division scheme in methods

@@
 
+THE EFFECTS OF THE NUMBER OF NODES AND DISTANCE ON THE COMPUTATIONS.
+
+\subsection{Recursive division algorithm}
+
 We have developed a modified version
-of the recursive scheme of \citet{Li2011}.
+of the recursive division scheme of \citet{Li2011}.
 We fix the number of nodes in the GLQ
@@
 in the radial, longitudinal, and latitudinal directions
-are approximately the dimensions of the tesseroid.
+are approximately the dimensions of the tesseroid
+\citep{Wild-Pfeiffer2008}.
 We define the distance $d$ as
@@
 (Fig.~\ref{fig:ratio}a).
+
+DESCRIBE ALGORITHM FIRST.
+
 Thus, the rule-of-thumb of \citet{Ku1977}
-can be written as
+can be interpreted as
 
 \begin{equation}
-    \frac{d}{L} < r,
+    \frac{d}{L_\zeta} < r,
     \label{eq:ratio}
@@
 where
-$L$ is one of the dimensions of the tesseroid
+$L_\zeta$, $\zeta \in \{r, \phi, \lambda\}$,
+is one of the dimensions of the tesseroid
 and
-$r$ is the distance/size ratio.
-For \citet{Ku1977} this ratio should be equal to one.
+$r$ is the maximum allowed ratio between distance and size.
+For \citet{Ku1977}, this ratio should be equal to one.
 The dimensions of the tesseroid are
@@
 
-\subsection{Comparison with a spherical half-shell}
+\subsection{Determination of $r$ by comparison with a spherical shell}
 

### 2014-08-25 manuscript.tex
Added figure to show when fails
Calculate the effect at a low height and show effect. The use higher
order.
Need better plot. Maybe use Cartopy.
Caption for glq fail example fig

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/vary-height-and-order}
+    \caption{
+        Example of the effect of varying
+        the computation height and the Gauss-Legendre Quadrature order
+        (number of nodes in each dimension).
+        The $g_{xy}$ component caused by a
+        $7^\circ \times 7^\circ \times 20\ km$ tesseroid
+        with $2.67\ g.cm^{-3}$ density
+        and top at $z=0$
+        was calculated on a regular grid
+        at different heights and using different orders.
+        The GLQ order used was the same for all three dimensions.
+        a) At $400\ km$ height and order 2, the effect is similar to that
+        presented by \citet{Asgharzadeh2007}. b) At $150\ km$ height and order
+        2, the effect resembles that of four point sources located at the GLQ
+        nodes. c) At $150\ km$ but with a higher order of 30, the effect is
+        again similar to the maps in \citet{Asgharzadeh2007}.
+    }
+    \label{fig:sample}
+\end{figure}
+
 \subsection{Recursive division algorithm}

### 2014-08-25 manuscript.tex
Made glq fail fig horizontal

@@
 \documentclass[extra]{gji}
+
 \usepackage{timet}
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics[width=\columnwidth]{figs/vary-height-and-order}
+    \includegraphics[width=\textwidth]{figs/vary-height-and-order}
     \caption{
@@
     \label{fig:sample}
-\end{figure}
+\end{figure*}
 
@@
 with that of a spherical half-shell
-(Fig.~\ref{fig:shell}),
 for which exist analytical expressions.
@@
 caused by the homogeneous spherical half-shell
-shown in Fig.~\ref{fig:shell}
 at a computation point
@@
 respectively.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figs/spherical-shell}
-    \caption{A spherical half-shell
-        spanning the ranges $0^{\circ} \le \lambda \le 180^\circ$,
-        $-90^\circ \le \phi \le 90^\circ$,
-        and $r_1 \le r \le r_2$,
-        the corresponding tesseroid model,
-        the computation point $P$
-        at $(\lambda=90^\circ,\phi=0^\circ,r=r_2+h)$,
-        and its local coordinate system $(x, y, z)$.
-    }
-    \label{fig:shell}
-\end{figure}
 

### 2014-08-25 manuscript.tex
Better figure letters in split and caption
Added part describing errors in glq

@@
-%\documentclass[extra,mreferee]{gji}
+\documentclass[extra,mreferee]{gji}
 % Use the option below to make a nicely formatted manuscript
-\documentclass[extra]{gji}
+%\documentclass[extra]{gji}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid}
+    \includegraphics{figs/tesseroid}
     \caption{
@@
 
-THE EFFECTS OF THE NUMBER OF NODES AND DISTANCE ON THE COMPUTATIONS.
+The accuracy of the integration
+depends on the number of nodes, or GLQ order,
+used in the discretization.
+However, \citet{Ku1977} showed
+that it also depends on the ratio between
+the distance to the computation point
+and the distance between adjacent nodes.
+Fig.~\ref{fig:sample}
+illustrates this effect.
+To increase the accuracy of the integration,
+one could increase the number of nodes,
+as in Fig.~\ref{fig:sample}c.
+Likewise,
+one could keep the number of nodes fixed
+and instead divide the tesseroid into smaller ones.
+This would increase the total number of nodes
+and effectively decrease
+the distance between adjacent nodes.
+The latter is the approach
+taken by \citet{Li2011}.
+Here, we will follow the same approach
+and make some improvements to the algorithm.
+We will also attempt to quantify
+the error committed in the numerical integration.
 
@@
 We have developed a modified version
-of the recursive division scheme of \citet{Li2011}.
+of the recursive division algorithm of \citet{Li2011}.
 We fix the number of nodes in the GLQ
@@
     \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
+    \includegraphics{figs/tesseroid-split}
     \caption{
@@
         and $P$.
-        $L$ is the largest dimension of the tesseroid.
+        $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the tesseroid.
     }

### 2014-08-25 manuscript.tex
Added a bit about the tesseroid dimensions

@@
 \end{equation}
+
 \begin{equation}
@@
 in which
+$G$ is the gravitational constant,
 $(\phi, \lambda, r)$ are
@@
 We have developed a modified version
-of the recursive division algorithm of \citet{Li2011}.
-We fix the number of nodes in the GLQ
-$N^\lambda=N^\phi=N^r=2$
-so that the distance between nodes
-in the radial, longitudinal, and latitudinal directions
+of the recursive division algorithm proposed by \citet{Li2011}.
+One of the modifications made is that
+we fix the number of GLQ nodes as
+$N^\lambda=N^\phi=N^r=2$.
+In this way, the distances between adjacent nodes
 are approximately the dimensions of the tesseroid
-\citep{Wild-Pfeiffer2008}.
+\citep[see][]{Wild-Pfeiffer2008}.
+Thus, we may use the tesseroid dimensions
+as proxies for the distance between nodes.
+We define the dimensions of a tesseroid
+in the radial, latitudinal, longitudinal directions, respectively,
+as
+
+\begin{eqnarray}
+    L_r &=& r_2 - r_1,\\
+    L_\phi &=& R(\phi_2 - \phi_1),\\
+    L_\lambda &=& R(\lambda_2 - \lambda_1),
+\end{eqnarray}
+
+\noindent
+where $R=6378137\ m$ is the mean Earth radius
+and the longitudes and latitudes are given in radians.
+
 We define the distance $d$ as
@@
 The dimensions of the tesseroid are
-
-\begin{eqnarray*}
-    L_r &=& r_2 - r_1,\\
-    L_\phi &=& R(\phi_2 - \phi_1),\\
-    L_\lambda &=& R(\lambda_2 - \lambda_1),
-\end{eqnarray*}
-
-\noindent
-where $R$ is the mean Earth radius
-and the longitudes and latitudes are given in radians.
 

### 2014-08-27 body.tex
Using new templates to build the paper

--- 2014-08-25 manuscript.tex
@@
-\documentclass[extra,mreferee]{gji}
-% Use the option below to make a nicely formatted manuscript
-%\documentclass[extra]{gji}
-
-\usepackage{timet}
-\usepackage{graphicx}
-
-\title[Optimal forward modeling using tesseroids]{
-    Optimal forward modeling of gravitational fields
-    using tesseroids with adaptive discretization
-}
-\author[Uieda et al.]{
-    Leonardo Uieda$^{1,2}$,
-    Vanderlei C. Oliveira Jr$^2$,
-    Val\'eria C. F. Barbosa$^2$
-    \\
-    $^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil.
-    $^2$Observat\'orio Nacional, Rio de Janeiro, Brazil.
-}
-
-\begin{document}
-
-\label{firstpage}
-\maketitle
-
-\begin{summary}
+\begin{abstract}
 Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
@@
 pellentesque, quis molestie sem vestibulum.
-\end{summary}
+\end{abstract}
 
@@
 pellentesque, quis molestie sem vestibulum.
-
-\bibliographystyle{gji}
-\bibliography{references}
-
-\end{document}

### 2014-08-27 templates.py

@@ -1,0 +1,178 @@
+"""
+Templates to build Latex files for specific journals.
+"""
+
+
+SEGABS_HEAD = r"""
+\documentclass{segabs}
+\newcommand{{\mitbf}}[1]{{\mathbf{{#1}}}}
+\usepackage{{graphicx}}
+
+\begin{{document}}
+
+\title{{{title}}}
+\lefthead{{{heads[0]}}}
+\righthead{{{heads[1]}}}
+
+{authors}
+"""
+
+GEOPHYISICS_HEAD = r"""
+\documentclass[{options}]{{geophysics}}
+{endfloat}
+\usepackage{{graphicx}}
+\newcommand{{\mitbf}}[1]{{\mathbf{{#1}}}}
+
+\begin{{document}}
+
+\title{{{title}}}
+\lefthead{{{heads[0]}}}
+\righthead{{{heads[1]}}}
+% manuscript number
+\ms{{}}
+
+{authors}
+"""
+
+GJI_HEAD = r"""
+\documentclass[{options}]{{gji}}
+\usepackage{{timet}}
+\usepackage{{graphicx}}
+
+\title[{heads[1]}]{{
+{title}
+}}
+
+\author[{heads[0]}]{{
+{authors}
+}}
+
+\begin{{document}}
+
+\label{{firstpage}}
+\maketitle
+"""
+
+FOOTER = r"""
+\bibliographystyle{{{bst}}}
+\bibliography{{{bibfile}}}
+\end{{document}}
+"""
+
+
+class Manuscript(object):
+    """
+    Gather metadata for a manuscript and make the main .tex file for a journal.
+
+    Use 'Manuscript.build' in a SCons Command to make the tex file from the
+    body of the text.
+
+    Example:
+
+    Putting this in a file named 'SConstruct':
+
+        from templates import Manuscript
+        builder = Manuscript(
+            title='My title',
+            author=['First Author',
+                    'Second Author'],
+            affil=['First auth affiliation.',
+                   'First and second auth affiliation.'],
+            # Indexes in 'affil' corresponding to each author
+            author_affil=[[0, 1], [1]],
+            journal='gji',
+            heads=['Author et al.', 'Short title'],
+            bibfile='references.bib')
+        env = Environment()
+        ms = env.Command('manuscript.tex', 'body.tex', builder.build)
+        pdf = env.PDF(target='manuscript.pdf', source=ms)
+
+    Then running 'scons' will produce 'manuscript.tex' and compile it to
[…]

### 2014-08-28 body.tex
Small fix
Tweaks to caption of vary height/order fig

@@
 Fig.~\ref{fig:sample}
-illustrates this effect.
+illustrates this effect
+on the $g_{xy}$ gravity gradient component.
 To increase the accuracy of the integration,
@@
         Example of the effect of varying
-        the computation height and the Gauss-Legendre Quadrature order
-        (number of nodes in each dimension).
-        The $g_{xy}$ component caused by a
+        the computation height
+        and the number of nodes in the Gauss-Legendre Quadrature
+        (i.e., the GLQ order).
+        The $g_{xy}$ component was produced by a
         $7^\circ \times 7^\circ \times 20\ km$ tesseroid
         with $2.67\ g.cm^{-3}$ density
-        and top at $z=0$
-        was calculated on a regular grid
-        at different heights and using different orders.
-        The GLQ order used was the same for all three dimensions.
-        a) At $400\ km$ height and order 2, the effect is similar to that
-        presented by \citet{Asgharzadeh2007}. b) At $150\ km$ height and order
-        2, the effect resembles that of four point sources located at the GLQ
-        nodes. c) At $150\ km$ but with a higher order of 30, the effect is
-        again similar to the maps in \citet{Asgharzadeh2007}.
+        and top at $z=0\ km$.
+        The maps were calculated on a regular grid
+        with $100\times100$ points.
+        The GLQ order used was the same
+        in the radial, longitudinal, and latitudial dimensions.
+        a) Field calculated at $400\ km$ height using GLQ order 2.
+        The effect is similar to that presented by \citet{Asgharzadeh2007}.
+        b) At $150\ km$ height and order 2,
+        the field resembles that of
+        four point sources located at the GLQ nodes.
+        This effect was shown by \citet{Ku1977}
+        for the case of the $g_z$ component of a right rectangular prism.
+        c) At $150\ km$ but with a higher GLQ order of 30,
+        the field is similiar to that expected for a single mass source.
     }

### 2014-08-28 templates.py
Update template files

@@ -84,19 +84,16 @@ class Manuscript(object):
             heads=['Author et al.', 'Short title'],
             bibfile='references.bib')
         env = Environment()
-        ms = env.Command('manuscript.tex', 'body.tex', builder.build)
+        ms = env.Command('manuscript.tex', 'body.tex', builder.build_ms)
         pdf = env.PDF(target='manuscript.pdf', source=ms)
 
     Then running 'scons' will produce 'manuscript.tex' and compile it to
     'manuscript.pdf' using the Geophysical Journal International template.
-
-    Pass 'pretty=True' to 'Manuscript' to produce a 'final paper'-looking
-    output.
     """
 
 
     def __init__(self, title, author, affil, author_affil, journal, heads,
-                 bibfile, pretty=False):
+                 bibfile):
         self.title = title
         self.author = author
         self.affil = affil
@@ -104,15 +101,15 @@ class Manuscript(object):
         self.journal = journal
         self.heads = heads
         self.bibfile = bibfile
-        self.pretty = pretty
 
 
-    def build(self, target, source, env):
+    def build_ms(self, target, source, env):
         """
         To be used in a SCons Command as:
 
             builder = Manuscript(title=...)
-            ms = env.Command('manuscrit.tex', 'ms_body.tex', builder.build)
+            latexfile = env.Command('paper.tex', 'ms_body.tex',
+                                    builder.build_ms)
 
         This will take the body of the text (just the sections and text, no
         documentclass, title, bibiliography, etc) and make the final tex file
@@ -125,7 +122,26 @@ class Manuscript(object):
             f.write(text)
 
 
-    def geophysics(self, body):
+    def build_paper(self, target, source, env):
+        """
+        To be used in a SCons Command as:
+
+            builder = Manuscript(title=...)
+            latexfile = env.Command('paper.tex', 'ms_body.tex',
+                                    builder.build_paper)
+
+        Same as 'build_ms' but will make a latex file for a pretty
+        (published-looking) version of the paper (if the document class has
+        options for this).
+        """
+        with open(str(source[0])) as f:
+            body = f.read()
+        text = getattr(self, self.journal)(body, pretty=True)
+        with open(str(target[0]), 'w') as f:
+            f.write(text)
+
+
+    def geophysics(self, body, pretty=False):
         """
         Builds manuscripts for Geophysics articles.
         """
@@ -141,7 +157,7 @@ class Manuscript(object):
[…]

### 2014-09-01 body.tex
Tweaks to methods

@@
 $\alpha,\beta \in \{x,y,z\}$
-are the local coordinate system of point P,
+are directions in the local coordinate system of point P,
 and
@@
 
-We have opted to use
-the 3D Gauss-Legendre Quadrature
-to numerically integrate
+Equations
 \ref{eq:tesspot},
@@
 and
-\ref{eq:tesstensor}.
+\ref{eq:tesstensor}
+can be numerically integrated using
+the 3D Gauss-Legendre Quadrature (GLQ)
+\citep{Asgharzadeh2007}.
 The GLQ
@@
 the discretization points (nodes) $x_i$
-are the roots of the $N$th order Legendre polynomial $P_N$
-scaled to the integration limits.
+are the roots of the $N$th order Legendre polynomial $P_N$.
+Note that the roots have to be scaled to the integration limits.
 The roots of a second-order Legendre polynomial are

### 2014-09-01 body.tex
Improvements to methods and finished algorithm

@@
 and make some improvements to the algorithm.
-We will also attempt to quantify
-the error committed in the numerical integration.
 
@@
 
-We have developed a modified version
-of the recursive division algorithm proposed by \citet{Li2011}.
-One of the modifications made is that
-we fix the number of GLQ nodes as
+We start by fixing the number of GLQ nodes to
 $N^\lambda=N^\phi=N^r=2$.
@@
 \noindent
-where $R=6378137\ m$ is the mean Earth radius
-and the longitudes and latitudes are given in radians.
+where $R=6,378,137\ m$ is the mean Earth radius.
+The longitudes and latitudes are given in radians.
 
-We define the distance $d$ as
-the Cartesian distance between
-the computation point $P$
+We write the condition for an accurate GLQ integration
+described by \citet{Ku1977}
+as a function of
+the average of the dimensions of the tesseroid, $L$,
+
+\begin{equation}
+    \frac{d}{L} > 1,
+\end{equation}
+
+\noindent
+where $d$ is the Cartesian distance
+between the computation point $P$
 and the center of the top face of the tesseroid
 (Fig.~\ref{fig:ratio}a).
-
-DESCRIBE ALGORITHM FIRST.
-
-Thus, the rule-of-thumb of \citet{Ku1977}
-can be interpreted as
+In a more general form, the condition can be expressed as
 
 \begin{equation}
-    \frac{d}{L_\zeta} < r,
+    \frac{d}{L} > r,
     \label{eq:ratio}
@@
 \noindent
-where
-$L_\zeta$, $\zeta \in \{r, \phi, \lambda\}$,
-is one of the dimensions of the tesseroid
-and
-$r$ is the maximum allowed ratio between distance and size.
-For \citet{Ku1977}, this ratio should be equal to one.
-The dimensions of the tesseroid are
+for a positive non-zero scalar $r$,
+henceforth referred to as
+the ``distance-size ratio''.
 
-In case the tesseroid is divided,
-the inequality in equation \ref{eq:ratio}
-is checked for each of the eight parts.
-This process is repeated recursively
-until no divisions are required,
-ensuring that equation \ref{eq:ratio}
-holds for all tesseroids created.
-The gravitational effect
-of the original tesseroid
-is then calculated as
-the sum of the effect
-of the smaller tesseroids.
+To guarantee that equation~\ref{eq:ratio} is satisfied,
+one could divide the tesseroid in half for each of the three dimensions.
+The division would be repeated recursively
+for each of the eight smaller tesseroids
+until equation~\ref{eq:ratio} is satisfied for all smaller tesseroids.
+Then, the effect of the original tesseroid can be calculated
+as the sum of the
+effects of the smaller tesseroids,
+each of which is calculated using a second-order GLQ.
+
+However, this would result in unnecessary divisions,
+specially if one dimension of the tesseroid is much smaller than the others.
+A more efficient approach is to check if the condition is met for each of the
+three dimensions of the tesseroid ($L_r$, $L_\phi$, and $L_\lambda$).
+We only divide the tesseroid in half along dimensions for which the condition
+is not met.
+
+This algorithm can be implemented
+in the form of a recursive function
+as follows:
+
+\begin{verbatim}
+function gravity(P, model):
+  result = 0
+  for each tesseroid in model:
+    d = distance from P to the top of tesseroid
+    if d < r*L_lon:
+      divide_lon = true
+    if d < r*L_lat:
+      divide_lat = true
+    if d < r*L_r:
+      divide_r = true
+    if divide_lon or divide_lat or divide_r:
+      newmodel = (divide tesseroid in the
+                  required dimensions)
+      result = result + gravity(P, newmodel)
+    else
+      result = result + effect of tesseroid
+  return result
+\end{verbatim}
+
+The distance-size ratio $r$ controls
+how far away $P$ has to be
+compared to the dimensions of the tesseroid.
+A ratio $r=2$ means that
+the recursive division algorithm will ensure that
+tesseroids are smaller than
+half the distance to $P$.
+Thus, the ratio determines how many divisions are required.
+Fig.~\ref{fig:ratio}
+illustrates the recursive division of a tesseroid
+(Fig.~\ref{fig:ratio}a)
+using increasing values of $r$.
+The tesseroid is divided into
+$28$ tesseroids for $r=1$ (Fig.~\ref{fig:ratio}b),
+$146$ tesseroids for $r=2$ (Fig.~\ref{fig:ratio}c),
+and $3106$ tesseroids for $r=6$ (Fig.~\ref{fig:ratio}d).
+
+Because the number of tesseroids increases with $r$,
+so does the computation time.
+Hence, there is need to determine
+an optimal value of the distance-size ratio $r$
+that yields an acceptable trade-off between
+accuracy and computation time.
+In the following section,
+we describe a methodology
+to quantify the error of the GLQ integration
+as a function of the distance-size ratio $r$.
+These results can be used
+to determine optimal values of $r$
+for the gravitational potential,
+gravitational attraction,
+and gravity gradient tensor.
 
@@
         for a computation point P
-        using $R$ equal to
+        using the distance-size ratio $r$ equal to
         (b) 1, (c) 2, and (d) 6.
-        $d$ is the distance between
+        $d$ is the Cartesian distance between
         the center of the top face of the tesseroid
@@
         $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the tesseroid.
+        Note that increasing $r$
+        results in a fine division of the tesseroid
+        close the computation point
+        and a coarser division further away.
     }
@@
 
-The rule-of-thumb of \citet{Ku1977}
-used by \citet{Li2011}
-is equivalent to setting $R=1$.
-Fig.~\ref{fig:ratio}
-illustrates the subdivision of a tesseroid
-(Fig.~\ref{fig:ratio}a)
-for computation point $P$
-using increasing values of $R$.
-Note that increasing $R$
-results is a finer division of the tesseroid
-around the computation point
-and thus a more accurate GLQ integration.
-However, higher values of $R$
-also result in
-larger computation times.
-For example,
-the tesseroid in Fig.~\ref{fig:ratio}a
-is replaced by
-50 tesseroids for $R=0.5$ (Fig.~\ref{fig:ratio}b),
-344 tesseroids for $R=1$ (Fig.~\ref{fig:ratio}c),
-and 2472 tesseroids for  $R=2$ (Fig.~\ref{fig:ratio}d).
-
-Clearly,
-there is need to determine
-an optimal value of $R$
-that yields the desired accuracy
-in the GLQ integration.
-In the following section,
-we compare the computed gravitational effects
-of a tesseroid model
-with that of a spherical half-shell
-for which exist analytical expressions.
-
-\subsection{Determination of $r$ by comparison with a spherical shell}
+\subsection{Comparison with a spherical shell}
 

### 2014-09-01 body.tex
Copied shell equations to text

@@
 Marussi (gravity gradient) tensor
-caused by the homogeneous spherical half-shell
-at a computation point
-$P = (\lambda=90^\circ, \phi=0^\circ, r=r_2+h)$
-are
+of a homogeneous spherical shell
+are \citep{Grombein2013}
 
 \begin{equation}
-    V(h) = 2\pi G \rho \left[
-        \frac{l^3 + {r'}^3}{3(r_2 + h)} - 0.5 {r'}^2 \right]
-         \Biggr|_{r'=r_1}^{r'=r_2},
-    \label{eq:shellpot}
+V(r) = \frac{4}{3}\pi G \rho \frac{r_2^3 - r_1^3}{r},
 \end{equation}
+
 \begin{equation}
-    g_z(h) = 2\pi G \rho \left[
-        \frac{l^3 + {r'}^3}{3(r_2 + h)^2} - l \right]
-        \Biggr|_{r'=r_1}^{r'=r_2},
-    \label{eq:shellgz}
+g_z(r) = \frac{V(r)}{r},
 \end{equation}
+
 \begin{equation}
-    g_{xx}(h) = g_{yy}(h) = -\frac{1}{2} g_{zz}(h),
+g_{zz}(r) = 2\frac{V(r)}{r^2},
 \end{equation}
+
 \begin{equation}
-    g_{xy}(h) = g_{xz}(h) = g_{yz}(h) = 0,
+g_{xx}(r) = g_{yy}(r) = -\frac{g_{zz}}{2},
 \end{equation}
@@
 \begin{equation}
-    g_{zz}(h) = 2\pi G \rho \left[
-        2\frac{l^3 + {r'}^3}{3(r_2 + h)^3}
-        - \frac{l}{r_2 + h} + \frac{r_2 + h}{l}
-        \right]
-        \Biggr|_{r'=r_1}^{r'=r_2},
-    \label{eq:shellgzz}
+g_x(r) = g_y(r) = g_{xy}(r) = g_{xz}(r) = g_{yz}(r) = 0,
 \end{equation}
@@
 \noindent
-where $\rho$ is the density,
-$l = \sqrt{(r_2 + h)^2 + {r'}^2}$,
-and
-$r_1$ and $r_2$ are radial coordinates of
-the bottom and top of the spherical shell,
-respectively.
-
-For our comparison,
-we used a spherical half-shell with
-$\rho=2.7\ g.cm^{-3}$,
-$r_1=6378.137\ km$ (mean Earth radius),
-and
-$r_2 = r_1 + 50\ km$.
-We discretized the half-shell
-into tesseroids of equal size
-and calculated the gravitational effects of
-the half-shell
-(using equations \ref{eq:shellpot}-\ref{eq:shellgzz})
-and the corresponding tesseroid model
-at various heights.
-These calculations were repeated
-for varying values of
-the distance/size ratio $R$.
-This analysis
-allows us to establish
-the optimal value of $R$
-that achieves a desired error level.
-
-Data and source code to produce the results of this comparison
-are provided by \textbf{(cite notebook on figshare)}.
+where $r$ is the radial coordinate of the computation point and
+$r_1$ and $r_2$ are radial coordinates
+of the bottom and top of the spherical shell, respectively.
 
@@
 
-The methods described above
-were implemented the software package
-Tesseroids version 1.1.1.
+The recursive division algorithm described above
+is implemented in the software package
+Tesseroids.
 The package consists
-of a range of command-line programs
+of command-line programs
 written in the C programming language.
-It is freely available
-online\footnote{www.leouieda.com/tesseroids}
-under the terms of
-the open-source BSD 3-clause license.
-\textbf{(cite archive on figshare)} provides a persistent
-archive for the source code and compiled binaries of Tesseroids.
+Tesseroids is open-source software
+and is freely available
+on the internet\footnote{www.leouieda.com/tesseroids}
+\textbf{CITE ARCHIVE} provides a persistent
+archive for the source code.
+
+The code used to produce the results and figures presented here
+is available at \textbf{DOI LINK AND CITATION}.
 
@@
 \section{Acknowledgments}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
-et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
-tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
-Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
-eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
-fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
-non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
-consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
-convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
-nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
-pellentesque, quis molestie sem vestibulum.

### 2014-09-02 body.tex
Corr to methods text and plot GLQ nodes in sample
Added some text to the part with the tesseroid equations. Added the
location of the GLQ nodes to the sample figure to make the case for
using the tesseroid dimensions instead node spacing.

@@
 
-Citing articles in text \citet{Asgharzadeh2007} and with parenthesis
-\citep{Braitenberg2011}.
+Using optimized Cartesian integration kernels by \citet{Grombein2013}.
+Their advantage over the spherical integration kernels
+\citep{Wild-Pfeiffer2008} is that they do not suffer from a singularity at the
+poles.
 
@@
         \int\limits_{r_1}^{r_2}
-        \frac{\Delta_{\alpha}}{\ell^3} \kappa dr' d\phi' d\lambda'
-        \ \ \ \alpha \in \{x,y,z\},
+        \frac{\Delta\alpha}{\ell^3} \kappa dr' d\phi' d\lambda',
     \label{eq:tessgrav}
@@
         \left(
-            \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
+            \frac{3\Delta\alpha \Delta\beta}{\ell^5} -
             \frac{\delta_{\alpha\beta}}{\ell^3}
         \right) \kappa, \\
-    \Delta_x &=& r' K_{\phi} , \\
-    \Delta_y &=& r' \cos \phi' \sin(\lambda' - \lambda) , \\
-    \Delta_z &=& r' \cos \psi - r, \\
+    \Delta x &=& r' K_{\phi} , \\
+    \Delta y &=& r' \cos \phi' \sin(\lambda' - \lambda) , \\
+    \Delta z &=& r' \cos \psi - r, \\
     \ell &=& \sqrt{r'^2 + r^2 - 2 r' r \cos \psi} , \\
@@
 \ref{eq:tesstensor}
-can be numerically integrated using
-the 3D Gauss-Legendre Quadrature (GLQ)
-\citep{Asgharzadeh2007}.
+must be numerically integrated
+\citep{Grombein2013}.
+We follow \citet{Asgharzadeh2007}
+and use the 3D Gauss-Legendre Quadrature (GLQ) integration.
 The GLQ
@@
         The GLQ order used was the same
-        in the radial, longitudinal, and latitudial dimensions.
+        in the radial, longitudinal, and latitudinal dimensions.
+        Black circles represent the horizontal location of the GLQ nodes
+        (not shown for GLQ order 30).
         a) Field calculated at $400\ km$ height using GLQ order 2.
@@
         c) At $150\ km$ but with a higher GLQ order of 30,
-        the field is similiar to that expected for a single mass source.
+        the field is similar to that expected for a single mass source.
     }
@@
 In this way, the distances between adjacent nodes
-are approximately the dimensions of the tesseroid
-\citep[see][]{Wild-Pfeiffer2008}.
+can be roughly approximated by the dimensions of the tesseroid
+(Fig.~\ref{fig:sample}a-b).
 Thus, we may use the tesseroid dimensions

### 2014-09-02 body.tex
Cite eq when checking the distance-size condition

@@
 specially if one dimension of the tesseroid is much smaller than the others.
-A more efficient approach is to check if the condition is met for each of the
-three dimensions of the tesseroid ($L_r$, $L_\phi$, and $L_\lambda$).
-We only divide the tesseroid in half along dimensions for which the condition
-is not met.
+A more efficient approach is
+to check if equation~\ref{eq:ratio} is met
+for each of the three dimensions of the tesseroid
+($L_r$, $L_\phi$, and $L_\lambda$).
+We only divide the tesseroid in half
+along dimensions for which the condition is not met.
 

### 2014-09-02 body.tex
Table with the parameters of the shell

@@
 
-The gravitational potential,
-gravitational attraction,
-and
-Marussi (gravity gradient) tensor
-of a homogeneous spherical shell
-are \citep{Grombein2013}
+We must establish reference values to evaluate the error of the numerical
+integration.
+
+The homogeneous spherical shell is a convenient geometry with analytical
+solutions for
+the gravitational potential, attraction, and gravity gradient tensor \citep{Grombein2013}
 
@@
 of the bottom and top of the spherical shell, respectively.
+
+The computation point can be considered at any longitude or latitude because of
+the symmetry of the spherical shell.
+
+The spherical shell had dimensions of $r_1 = R$ ($R$ is the mean Earth radius) and $r_2 = R +
+1\ km$ and density $\rho = 2.67\ g.cm^{-3}$.
+
+The largest errors due are concentrated directly on top of the tesseroid
+(Fig.~\ref{fig:sample}b).
+
+Therefore, we perform the computations on an $11 \times 11$ point regular grid
+in an area directly on top of the tesseroid with corner at $\lambda = 0^\circ$
+$\phi = 0^\circ$.
+
+The computation grid was located at a constant height above the spherical shell
+$h = r_2 + 1\ km$.
+
+Table~\ref{tab:shell} contains the parameters defining the spherical shell.
+
+\begin{table}
+    \caption{Parameters of the spherical shell.}
+    \label{tab:shell}
+    \begin{tabular}{lr}
+        $\rho$                    & $2.76\ g.cm^{-3}$ \\
+        $\Delta r$                & $1\ km$ \\
+        $r_1 = R$                 & $6,378.137\ km$ \\
+        $r_2 = r_1 + \Delta r$    & $6,379.137\ km$ \\
+    \end{tabular}
+\end{table}
 

### 2014-09-02 body.tex
Slight mods to shell tex

@@
 
-We must establish reference values to evaluate the error of the numerical
-integration.
-
-The homogeneous spherical shell is a convenient geometry with analytical
-solutions for
-the gravitational potential, attraction, and gravity gradient tensor \citep{Grombein2013}
+We must establish reference values
+to evaluate the error of the numerical integration.
+For this purpose, we have used a homogeneous spherical shell
+with its bottom at a sphere with radius $R$ (the mean Earth radius),
+thickness $\Delta r$, and density $\rho$.
+It is a convenient geometry
+with analytical solutions for
+the gravitational potential, attraction, and gravity gradient tensor
+\citep{Grombein2013}
 
@@
 where $r$ is the radial coordinate of the computation point and
-$r_1$ and $r_2$ are radial coordinates
+$r_1 = R$ and $r_2 = r_1 + \Delta r$ are radial coordinates
 of the bottom and top of the spherical shell, respectively.
+Table~\ref{tab:shell} summarises the parameters defining the spherical shell.
 
@@
 
-Table~\ref{tab:shell} contains the parameters defining the spherical shell.
-
 \begin{table}
@@
     \begin{tabular}{lr}
-        $\rho$                    & $2.76\ g.cm^{-3}$ \\
+        $\rho$                    & $2.67\ g.cm^{-3}$ \\
         $\Delta r$                & $1\ km$ \\

### 2014-09-10 body.tex
Put the condition in pseudocode like equations

@@
     d = distance from P to the top of tesseroid
-    if d < r*L_lon:
+    if d/L_lon < r:
       divide_lon = true
-    if d < r*L_lat:
+    if d/L_lat < r:
       divide_lat = true
-    if d < r*L_r:
+    if d/L_r < r:
       divide_r = true

### 2014-09-10 body.tex
Dist/size ratio should be D, r is radius already

@@
 \begin{equation}
-    \frac{d}{L} > r,
+    \frac{d}{L} > D,
     \label{eq:ratio}
@@
 \noindent
-for a positive non-zero scalar $r$,
+for a positive non-zero scalar $D$,
 henceforth referred to as
@@
     d = distance from P to the top of tesseroid
-    if d/L_lon < r:
+    if d/L_lon < D:
       divide_lon = true
-    if d/L_lat < r:
+    if d/L_lat < D:
       divide_lat = true
-    if d/L_r < r:
+    if d/L_r < D:
       divide_r = true
@@
 
-The distance-size ratio $r$ controls
+The distance-size ratio $D$ controls
 how far away $P$ has to be
 compared to the dimensions of the tesseroid.
-A ratio $r=2$ means that
+A ratio $D=2$ means that
 the recursive division algorithm will ensure that
@@
 (Fig.~\ref{fig:ratio}a)
-using increasing values of $r$.
+using increasing values of $D$.
 The tesseroid is divided into
-$28$ tesseroids for $r=1$ (Fig.~\ref{fig:ratio}b),
-$146$ tesseroids for $r=2$ (Fig.~\ref{fig:ratio}c),
-and $3106$ tesseroids for $r=6$ (Fig.~\ref{fig:ratio}d).
+$28$ tesseroids for $D=1$ (Fig.~\ref{fig:ratio}b),
+$146$ tesseroids for $D=2$ (Fig.~\ref{fig:ratio}c),
+and $3106$ tesseroids for $D=6$ (Fig.~\ref{fig:ratio}d).
 
-Because the number of tesseroids increases with $r$,
+Because the number of tesseroids increases with $D$,
 so does the computation time.
 Hence, there is need to determine
-an optimal value of the distance-size ratio $r$
+an optimal value of the distance-size ratio $D$
 that yields an acceptable trade-off between
@@
 to quantify the error of the GLQ integration
-as a function of the distance-size ratio $r$.
+as a function of the distance-size ratio $D$.
 These results can be used
-to determine optimal values of $r$
+to determine optimal values of $D$
 for the gravitational potential,
@@
         for a computation point P
-        using the distance-size ratio $r$ equal to
+        using the distance-size ratio $D$ equal to
         (b) 1, (c) 2, and (d) 6.
@@
         $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the tesseroid.
-        Note that increasing $r$
+        Note that increasing $D$
         results in a fine division of the tesseroid
@@
 where $r$ is the radial coordinate of the computation point and
-$r_1 = R$ and $r_2 = r_1 + \Delta r$ are radial coordinates
+$r_1$ and $r_2$ are radial coordinates
 of the bottom and top of the spherical shell, respectively.
-Table~\ref{tab:shell} summarises the parameters defining the spherical shell.
+Table~\ref{tab:shell} summarizes the parameters defining the spherical shell.
 
-The computation point can be considered at any longitude or latitude because of
+The shell was discretized into tesseroids with $S \times S \times 1\ km$ size.
+
+We then compared the result of the spherical shell with that calculated for the
+tesseroid model.
+
+The tesseroid calculations were performed with the recursive division
+algorithm.
+
+The tesseroid calculations were repeated for values of the distance/size ratio
+$D$ ranging from zero (i.e., no divisions) to eight with an interval of 0.5.
+
+The computation point can be considered at any longitude and latitude because of
 the symmetry of the spherical shell.
 
-The spherical shell had dimensions of $r_1 = R$ ($R$ is the mean Earth radius) and $r_2 = R +
-1\ km$ and density $\rho = 2.67\ g.cm^{-3}$.
+Fig.~\ref{fig:sample}b shows that the largest errors are concentrated directly
+on top of the tesseroid.
 
-The largest errors due are concentrated directly on top of the tesseroid
-(Fig.~\ref{fig:sample}b).
+Making the comparison between the tesseroid
 
-Therefore, we perform the computations on an $11 \times 11$ point regular grid
-in an area directly on top of the tesseroid with corner at $\lambda = 0^\circ$
-$\phi = 0^\circ$.
+We perform the computations on an $11 \times 11$ point regular grid.
 
-The computation grid was located at a constant height above the spherical shell
-$h = r_2 + 1\ km$.
+The southwestern corner of the grid is $(\lambda, \phi) = (0, 0)$
+and the northeastern corner is $(S, S)$, i.e., directly on top of a tesseroid.
+
+The grid is located at a constant height $h = 1\ km$ above the spherical shell.
+
+The computations where repeated for discretization into tesseroids of sizes $S = 1^\circ$, $2^\circ$,
+$5^\circ$, and $30^\circ$.
 

### 2014-09-10 body.tex
Added table with shell data

@@
 
-The computation point can be considered at any longitude and latitude because of
-the symmetry of the spherical shell.
+Fig.~\ref{fig:sample}b shows that the largest errors are concentrated directly
+on top of the tesseroid and near the quadrature nodes.
 
-Fig.~\ref{fig:sample}b shows that the largest errors are concentrated directly
-on top of the tesseroid.
+Because of that, comparing the shell and tesseroid effects at a single point
+would not be representative of the actual error committed.
 
-Making the comparison between the tesseroid
+The gravitational fields of the spherical shell are the same for any point at a
+radial distance $r>r_2$ because of the symmetry of the shell.
 
-We perform the computations on an $11 \times 11$ point regular grid.
+We perform the computations for the tesseroid model on an $11 \times 11$ point regular grid.
 
@@
     \label{tab:shell}
-    \begin{tabular}{lr}
+    \begin{tabular}{ll}
         $\rho$                    & $2.67\ g.cm^{-3}$ \\
-        $\Delta r$                & $1\ km$ \\
+        $\Delta r$                & $1\ km$\\
         $r_1 = R$                 & $6,378.137\ km$ \\
         $r_2 = r_1 + \Delta r$    & $6,379.137\ km$ \\
+    \end{tabular}
+\end{table}
+
+\begin{table}
+    \caption{Reference values of the gravitational potential, attraction, and
+    gravity gradient tensor of the spherical shell.}
+    \label{tab:shell-data}
+    \begin{tabular}{ll}
+        $V$                    &    $14278.021191\ m^2.s^{-2}$ \\
+        $g_z$                  &    $223.788630$ mGal \\
+        $g_{xx} = g_{yy}$      &    $-0.350758$ E\"otv\"os\\
+        $g_{zz}$               &    $0.701517$ E\"otv\"os\\
     \end{tabular}

### 2014-09-10 body.tex
Few more sentences in shell section

@@
 Table~\ref{tab:shell} summarizes the parameters defining the spherical shell.
+Table~\ref{tab:shell-data} contains the values of the gravitational potential,
+attraction, and gravity gradient tensor for such a shell computed at a height
+$h = 1\ km$ above the shell.
 
@@
 Because of that, comparing the shell and tesseroid effects at a single point
-would not be representative of the actual error committed.
+would not be representative of the actual error.
 
@@
 
-We perform the computations for the tesseroid model on an $11 \times 11$ point regular grid.
+We performed the computations for the tesseroid model on an $11 \times 11$ point regular grid.
 
@@
 
-The grid is located at a constant height $h = 1\ km$ above the spherical shell.
+The grid was located at a constant height $h = 1\ km$ above the spherical shell.
 
-The computations where repeated for discretization into tesseroids of sizes $S = 1^\circ$, $2^\circ$,
-$5^\circ$, and $30^\circ$.
+We can take the error in the computation as the maximum absolute difference between the
+value calculated for the spherical shell and the corresponding value for the
+tesseroid model.
+
+The above computations were repeated for tesseroids of sizes $S = 1^\circ$, $2^\circ$,
+$5^\circ$, and $30^\circ$ to verify if the result depends on the discretization.
 
@@
         $g_z$                                       &  $223.788630$ mGal \\
-        $g_{xx} = g_{yy}$      &    $-0.350758$ E\"otv\"os\\
+        $g_{xx}$, $g_{yy}$                          &  $-0.350758$ E\"otv\"os\\
         $g_{zz}$                                    &  $0.701517$ E\"otv\"os\\
+        $g_x$, $g_y$, $g_{xy}$, $g_{xz}$, $g_{yz}$  &  $0$\\
     \end{tabular}
@@
 and is freely available
-on the internet\footnote{www.leouieda.com/tesseroids}
+on the Internet\footnote{leouieda.com/tesseroids}
 \textbf{CITE ARCHIVE} provides a persistent
@@
 \section{Results}
-
-Meh.
 

### 2014-11-07 templates.py
Templates don't really work
Revert back to a simple makefile and GJI template. It will be easier to
copy this to writeLatex or something when co-authors need to review

@@ -1,194 +1,0 @@
-"""
-Templates to build Latex files for specific journals.
-"""
-
-
-SEGABS_HEAD = r"""
-\documentclass{segabs}
-\newcommand{{\mitbf}}[1]{{\mathbf{{#1}}}}
-\usepackage{{graphicx}}
-
-\begin{{document}}
-
-\title{{{title}}}
-\lefthead{{{heads[0]}}}
-\righthead{{{heads[1]}}}
-
-{authors}
-"""
-
-GEOPHYISICS_HEAD = r"""
-\documentclass[{options}]{{geophysics}}
-{endfloat}
-\usepackage{{graphicx}}
-\newcommand{{\mitbf}}[1]{{\mathbf{{#1}}}}
-
-\begin{{document}}
-
-\title{{{title}}}
-\lefthead{{{heads[0]}}}
-\righthead{{{heads[1]}}}
-% manuscript number
-\ms{{}}
-
-{authors}
-"""
-
-GJI_HEAD = r"""
-\documentclass[{options}]{{gji}}
-\usepackage{{timet}}
-\usepackage{{graphicx}}
-
-\title[{heads[1]}]{{
-{title}
-}}
-
-\author[{heads[0]}]{{
-{authors}
-}}
-
-\begin{{document}}
-
-\label{{firstpage}}
-\maketitle
-"""
-
-FOOTER = r"""
-\bibliographystyle{{{bst}}}
-\bibliography{{{bibfile}}}
-\end{{document}}
-"""
-
-
-class Manuscript(object):
-    """
-    Gather metadata for a manuscript and make the main .tex file for a journal.
-
-    Use 'Manuscript.build' in a SCons Command to make the tex file from the
-    body of the text.
-
-    Example:
-
-    Putting this in a file named 'SConstruct':
-
-        from templates import Manuscript
-        builder = Manuscript(
-            title='My title',
-            author=['First Author',
-                    'Second Author'],
-            affil=['First auth affiliation.',
-                   'First and second auth affiliation.'],
-            # Indexes in 'affil' corresponding to each author
-            author_affil=[[0, 1], [1]],
-            journal='gji',
-            heads=['Author et al.', 'Short title'],
-            bibfile='references.bib')
-        env = Environment()
-        ms = env.Command('manuscript.tex', 'body.tex', builder.build_ms)
-        pdf = env.PDF(target='manuscript.pdf', source=ms)
-
-    Then running 'scons' will produce 'manuscript.tex' and compile it to
[…]

### 2014-11-09 manuscript-old.tex
Started rewriting the text
If got to the point where I had no idea what I was doing anymore.
Started a clean MS and threw some loose sentences in there for the
introduction. Need something about the uses of forward modeling
(inversion and terrain correction) and why use tesseroids (GOCE, large
areas, unprecedented accuracy, bla bla bla).

--- 2014-09-10 body.tex
@@
+\documentclass[extra,paper]{gji}
+\usepackage{timet}
+\usepackage{graphicx}
+
+\title[Optimal forward modeling using tesseroids]{
+Optimal forward modeling of gravitational fields using tesseroids with adaptive discretization
+}
+
+\author[Uieda et al.]{
+Leonardo Uieda$^{1,2}$,
+Vanderlei C. Oliveira Jr$^{2}$,
+Val\'eria C. F. Barbosa$^{2}$\\
+$^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil.
+$^2$Observat\'orio Nacional, Rio de Janeiro, Brazil.
+}
+
+\begin{document}
+
+\label{firstpage}
+\maketitle
+
 \begin{abstract}
@@
 \section{Acknowledgments}
+
+\bibliographystyle{gji}
+\bibliography{references.bib}
+\end{document}

### 2014-11-09 manuscript.tex
Started rewriting the text
If got to the point where I had no idea what I was doing anymore.
Started a clean MS and threw some loose sentences in there for the
introduction. Need something about the uses of forward modeling
(inversion and terrain correction) and why use tesseroids (GOCE, large
areas, unprecedented accuracy, bla bla bla).

--- 2014-08-25 manuscript.tex
@@
-%\documentclass[extra,mreferee]{gji}
-% Use the option below to make a nicely formatted manuscript
-\documentclass[extra]{gji}
-
+\documentclass[extra,mreferee]{gji}
+%\documentclass[extra,paper]{gji}
 \usepackage{timet}
@@
 
-\title[Optimal forward modeling using tesseroids]{
-    Optimal forward modeling of gravitational fields
-    using tesseroids with adaptive discretization
+\title[Tesseroid forward modeling with adaptive discretization]{
+Forward modeling of gravitational fields using tesseroids with adaptive
+discretization
 }
+
 \author[Uieda et al.]{
 Leonardo Uieda$^{1,2}$,
-    Vanderlei C. Oliveira Jr$^2$,
-    Val\'eria C. F. Barbosa$^2$
+Val\'eria C. F. Barbosa$^{2}$,
+Vanderlei C. Oliveira Jr$^{2}$
 \\
@@
 
-\begin{summary}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
-et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
-tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
-Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
-eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
-fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
-non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
-consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
-convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
-nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
-pellentesque, quis molestie sem vestibulum.
-\end{summary}
+\begin{abstract}
+\end{abstract}
 
@@
 
-Citing articles in text \citet{Asgharzadeh2007} and with parenthesis
-\citep{Braitenberg2011}.
+We use the optimized formula of \citet{Grombein2013} with Cartesian integral
+kernels.
 
-The above integrals
-have to be solved numerically
-\citep{Wild-Pfeiffer2008}
-and different approaches have been proposed in the literature.
-\citet{Heck2007}
-used a Taylor series expansion
-of the integral kernels
-around the geometrical center of the tesseroid.
-\citet{Asgharzadeh2007}
-used the Gauss-Legendre Quadrature (GLQ)
-for the numerical integration.
-\citet{Wild-Pfeiffer2008} investigated
-the use of both methods,
-along with alternative mass elements,
-and found the GLQ to be the most efficient method.
-We have opted to use
-the 3D GLQ integration in our computations.
+The Cartesian kernels are faster to compute.
 
-The accuracy of the numerical integration
-using the GLQ
-was investigated by \citet{Ku1977}
-for the case of
-the gravitational attraction of
-a right-rectangular prism
-in Cartesian coordinates.
-\citet{Ku1977} states that the accuracy
-depends on the ratio between
-the average distance between GLQ nodes
-and the distance to the observation point P.
-As a rule-of-thumb,
-\citet{Ku1977} suggests that
-the distance between nodes
-should not be greater than
-the distance to point P.
-Extrapolating this rule-of-thumb
-to tesseroids,
-\citet{Li2011}
-devised a scheme that
-recursively divides a tesseroid
-into smaller ones when the rule is violated.
-This procedure
-effectively decreases the distance between nodes
-and, theoretically, guarantees the accuracy of the GLQ integration.
+The Cartesian kernels don't have singularities in the poles.
+
+The traditional spherical integral kernels suffer from a singularity at the
+poles \citep{Heck2007, Wild-Pfeiffer2008}.
+
+The Cartesian formula are numerically integrated using a Taylor series
+expansion as proposed by \citet{Heck2007}.
+
+The Taylor series expansion approach produces accurate results at low
+latitudes.
+
+The Taylor series expansion presents a decrease in accuracy towards the polar
+regions.
+
+The tesseroids have an approximately rectangular surface area at low latitudes
+but collapse into an approximately triangular shape at the poles.
+
+\citet{Grombein2013} use a near-zone separation to mitigate the increased error
+in high latitudes.
+
+In the so called "near-zone" of the computation point they use a finner
+discretization with smaller tesseroids.
+
+This is accomplished by dividing the tesseroids along the horizontal
+dimensions.
+
+The determination of an optimal size of the near-zone remains an open question
+\citep{Grombein2013}.
+
+We use the Gauss-Legendre Quadrature (GLQ) to numerically integrate the Cartesian
+kernels instead of the Taylor series expansion \citep{Asgharzadeh2007}.
+
+The GLQ integration consists of approximating the volume integral by a weighted sum of
+the effect of point masses.
+
+The point masses are distributed according to the roots of Legendre polynomials
+\citep{Hildebrand1987}.
+
+An advantage of the GLQ approach is that the accuracy of integration can be
+controlled by the number of point masses used.
+
+A disadvantage of the GLQ is the increased computation time as the number of
+point masses increases.
+
+There is a trade-off between accuracy and computation time.
+
+\citet{Ku1977} suggests that the accuracy of the GLQ integration depends on
+the ratio between distance to the computation point and the distance between
+adjacent point masses.
+
+\citet{Ku1977} proposes an empirical criteria that the distance between adjacent
+point masses should be less than the distance to the computation point.
+
+This empirical criteria was used by \citet{Asgharzadeh2007} as a way to
+choose the number of point masses used in the GLQ integration.
+
+They suggested using this criteria for gravitational attraction as well as
+the gravity gradient tensor (or Marussi tensor) of a tesseroid.
+
+\citet{Ku1977} presented results for the vertical component of the
+gravitational acceleration ($g_z$) caused by a right rectangular prism.
+
+To our knowledge, there has been no investigation if the empirical criteria of
+\citet{Ku1977} is valid for the second derivatives of the
+gravitational potential.
+
+There has also been no attempt to quantify the error committed in the GLQ
+integration of gravity gradients when applying the criteria of \citet{Ku1977}.
+
+\citet{Li2011} devised an automatic algorithm to enforce the criteria of
+\citet{Ku1977}.
+
+Their algorithm keeps the number of points masses per tesseroid fixed.
+
+They then check the ratio between the minimum distance to the computation point
+and the largest dimension of the tesseroid.
+
+If the ratio below a specified threshold value, the tesseroid is divided into
+smaller tesseroids.
+
+This division is repeated recursively until all tesseroids obey the criteria.
+
+The GLQ integration is then performed for each of the smaller tesseroids
+using the fixed number of point masses.
+
+The algorithm of \citet{Li2011} is similar to the near-zone separation used by
+\citet{Grombein2013}.
+
+The advantage of the adaptive discretization of \citet{Li2011} over simply
+increasing the number of points masses in the GLQ is that the
+amount of point masses will be greater only close to the computation point.
+
+This makes the adaptive discretization more computationally efficient.
+
+We propose an improvement over the adaptive discretization algorithm of
+\citet{Li2011} to further increase the computational efficiency.
+
+We will also quantify the error committed for the gravity gradients.
+
+This is done in an attempt to establish an optimal minimum threshold for the
+ratio between the distance to the computation point and the dimensions of the
+tesseroid.
 
@@
 
-The gravitational potential,
-gravitational acceleration,
-and Marussi (gravity gradient) tensor
-caused by a homogeneous tesseroid
-with density $\rho$
-are \citep{Grombein2013}
-
-\begin{equation}
-    V(r,\phi,\lambda) = G \rho
-        \int\limits_{\lambda_1}^{\lambda_2}
-        \int\limits_{\phi_1}^{\phi_2}
-        \int\limits_{r_1}^{r_2}
-        \frac{1}{\ell} \kappa  dr' d\phi' d\lambda',
-    \label{eq:tesspot}
-\end{equation}
-\begin{equation}
-    g_{\alpha}(r,\phi,\lambda) = G \rho
-        \int\limits_{\lambda_1}^{\lambda_2}
-        \int\limits_{\phi_1}^{\phi_2}
-        \int\limits_{r_1}^{r_2}
-        \frac{\Delta_{\alpha}}{\ell^3} \kappa dr' d\phi' d\lambda'
-        \ \ \ \alpha \in \{x,y,z\},
-    \label{eq:tessgrav}
-\end{equation}
-
-\noindent
-and
-
-\begin{equation}
-    g_{\alpha\beta}(r,\phi,\lambda) = G \rho
-        \int\limits_{\lambda_1}^{\lambda_2}
-        \int\limits_{\phi_1}^{\phi_2}
-        \int\limits_{r_1}^{r_2}
-        I_{\alpha\beta}
-        dr' d\phi' d\lambda',
-    \label{eq:tesstensor}
-\end{equation}
-
-\noindent
-in which
-$(\phi, \lambda, r)$ are
-the latitude, longitude, and radius
-coordinates of computation point P (Fig.~\ref{fig:tesseroid}),
-$\alpha,\beta \in \{x,y,z\}$
-are the local coordinate system of point P,
-and
-
-\begin{eqnarray}
-    I_{\alpha\beta} &=&
-        \left(
-            \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
-            \frac{\delta_{\alpha\beta}}{\ell^3}
-        \right) \kappa, \\
-    \Delta_x &=& r' K_{\phi} , \\
-    \Delta_y &=& r' \cos \phi' \sin(\lambda' - \lambda) , \\
-    \Delta_z &=& r' \cos \psi - r, \\
-    \ell &=& \sqrt{r'^2 + r^2 - 2 r' r \cos \psi} , \\
-    \cos\psi &=& \sin\phi\sin\phi' + \cos\phi\cos\phi'
-                 \cos(\lambda' - \lambda) , \\
-    K_{\phi} &=& \cos\phi\sin\phi' - \sin\phi\cos\phi'
-                 \cos(\lambda' - \lambda), \\
-    \kappa &=& {r'}^2 \cos \phi'.
-\end{eqnarray}
-
-\noindent
-$\delta_{\alpha\beta}$ is the Kronecker delta,
-i.e. $\delta_{\alpha\beta}=1$ if $\alpha=\beta$
-and $\delta_{\alpha\beta}=0$ otherwise.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid}
-    \caption{
-        View of a tesseroid,
-        the integration point $Q$ inside the tesseroid,
-        a geocentric coordinate system $(X, Y, Z)$,
-        the computation $P$ and it's local coordinate system $(x, y, z)$.
-        $r$, $\phi$, $\lambda$ are
-        the radius, latitude, and longitude, respectively, of point $P$,
-        and $\ell$ is the Cartesian distance between $P$ and $Q$.
-    }
-    \label{fig:tesseroid}
-\end{figure}
-
-We have opted to use
-the 3D Gauss-Legendre Quadrature
-to numerically integrate
-\ref{eq:tesspot},
-\ref{eq:tessgrav},
-and
-\ref{eq:tesstensor}.
-The GLQ
-approximates the integral by
-a weighted sum
-\citep{Hildebrand1987},
-e.g.
-
-\begin{equation}
-    \int\limits_a^b f(x) dx \approx
-    \frac{b-a}{2}\sum\limits_{i=1}^N W_i f(x_i),
-\end{equation}
-
-\noindent
-where
-$W_i$ are weights and
-the discretization points (nodes) $x_i$
-are the roots of the $N$th order Legendre polynomial $P_N$
-scaled to the integration limits.
-The roots of a second-order Legendre polynomial are
-$x_1=-0.577350269$ and $x_2=0.577350269$.
-For an arbitrary order $N$,
-we have used the root-finder algorithm
-of \citet{Barrera-Figueroa2006}
-to calculate the roots of $P_N$.
-The weights $W_i$ are \citep{Hildebrand1987}
-
-\begin{equation}
-    W_i = \frac{2}{(1 - x_i^2)(P'_N(x_i))^2},
-\end{equation}
-
-\noindent
-where $P'_N$ is the first derivative of $P_N$.
-
-The triple integrals in equations
-\ref{eq:tesspot},
-\ref{eq:tessgrav},
-and
-\ref{eq:tesstensor}
-become
-
-\begin{equation}
-    V(r,\phi,\lambda) \approx
-        A
-        \sum\limits_{k=1}^{N^{\lambda}}
-        \sum\limits_{j=1}^{N^{\phi}}
-        \sum\limits_{i=1}^{N^r}
-        W^r_i W^{\phi}_j W^{\lambda}_k
-        \frac{1}{\ell} \kappa,
-\end{equation}
-\begin{equation}
-    g_{\alpha}(r,\phi,\lambda) \approx
-        A
-        \sum\limits_{k=1}^{N^{\lambda}}
-        \sum\limits_{j=1}^{N^{\phi}}
-        \sum\limits_{i=1}^{N^r}
-        W^r_i W^{\phi}_j W^{\lambda}_k
-        \frac{\Delta_{\alpha}}{\ell^3} \kappa,
-\end{equation}
-
-\noindent
-and
-
-\begin{equation}
-    g_{\alpha\beta}(r,\phi,\lambda) \approx
-        A
-        \sum\limits_{k=1}^{N^{\lambda}}
-        \sum\limits_{j=1}^{N^{\phi}}
-        \sum\limits_{i=1}^{N^r}
-        W^r_i W^{\phi}_j W^{\lambda}_k
-        I_{\alpha\beta},
-\end{equation}
-
-\noindent
-where
-
-\begin{equation}
-    A = G \rho
-    \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8},
-\end{equation}
-
-\noindent
-$W_i^r$, $W_j^{\phi}$, and $W_k^{\lambda}$
-are the weights
-and $N^r$, $N^{\phi}$, and $N^{\lambda}$
-are the number of nodes
-for the radial, latitudinal, and longitudinal dimensions, respectively.
-
-THE EFFECTS OF THE NUMBER OF NODES AND DISTANCE ON THE COMPUTATIONS.
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figs/vary-height-and-order}
-    \caption{
-        Example of the effect of varying
-        the computation height and the Gauss-Legendre Quadrature order
-        (number of nodes in each dimension).
-        The $g_{xy}$ component caused by a
-        $7^\circ \times 7^\circ \times 20\ km$ tesseroid
-        with $2.67\ g.cm^{-3}$ density
-        and top at $z=0$
-        was calculated on a regular grid
-        at different heights and using different orders.
-        The GLQ order used was the same for all three dimensions.
-        a) At $400\ km$ height and order 2, the effect is similar to that
-        presented by \citet{Asgharzadeh2007}. b) At $150\ km$ height and order
-        2, the effect resembles that of four point sources located at the GLQ
-        nodes. c) At $150\ km$ but with a higher order of 30, the effect is
-        again similar to the maps in \citet{Asgharzadeh2007}.
-    }
-    \label{fig:sample}
-\end{figure*}
-
-\subsection{Recursive division algorithm}
-
-We have developed a modified version
-of the recursive division scheme of \citet{Li2011}.
-We fix the number of nodes in the GLQ
-$N^\lambda=N^\phi=N^r=2$
-so that the distance between nodes
-in the radial, longitudinal, and latitudinal directions
-are approximately the dimensions of the tesseroid
-\citep{Wild-Pfeiffer2008}.
-We define the distance $d$ as
-the Cartesian distance between
-the computation point $P$
-and the center of the top face of the tesseroid
-(Fig.~\ref{fig:ratio}a).
-
-DESCRIBE ALGORITHM FIRST.
-
-Thus, the rule-of-thumb of \citet{Ku1977}
-can be interpreted as
-
-\begin{equation}
-    \frac{d}{L_\zeta} < r,
-    \label{eq:ratio}
-\end{equation}
-
-\noindent
-where
-$L_\zeta$, $\zeta \in \{r, \phi, \lambda\}$,
-is one of the dimensions of the tesseroid
-and
-$r$ is the maximum allowed ratio between distance and size.
-For \citet{Ku1977}, this ratio should be equal to one.
-The dimensions of the tesseroid are
-
-\begin{eqnarray*}
-    L_r &=& r_2 - r_1,\\
-    L_\phi &=& R(\phi_2 - \phi_1),\\
-    L_\lambda &=& R(\lambda_2 - \lambda_1),
-\end{eqnarray*}
-
-\noindent
-where $R$ is the mean Earth radius
-and the longitudes and latitudes are given in radians.
-
-In case the tesseroid is divided,
-the inequality in equation \ref{eq:ratio}
-is checked for each of the eight parts.
-This process is repeated recursively
-until no divisions are required,
-ensuring that equation \ref{eq:ratio}
-holds for all tesseroids created.
-The gravitational effect
-of the original tesseroid
-is then calculated as
-the sum of the effect
-of the smaller tesseroids.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
-    \caption{
-        Adaptive discretization
-        of the tesseroid shown in (a)
-        for a computation point P
-        using $R$ equal to
-        (b) 1, (c) 2, and (d) 6.
-        $d$ is the distance between
-        the center of the top face of the tesseroid
-        and $P$.
-        $L$ is the largest dimension of the tesseroid.
-    }
-    \label{fig:ratio}
-\end{figure}
-
-The rule-of-thumb of \citet{Ku1977}
-used by \citet{Li2011}
-is equivalent to setting $R=1$.
-Fig.~\ref{fig:ratio}
-illustrates the subdivision of a tesseroid
-(Fig.~\ref{fig:ratio}a)
-for computation point $P$
-using increasing values of $R$.
-Note that increasing $R$
-results is a finer division of the tesseroid
-around the computation point
-and thus a more accurate GLQ integration.
-However, higher values of $R$
-also result in
-larger computation times.
-For example,
-the tesseroid in Fig.~\ref{fig:ratio}a
-is replaced by
-50 tesseroids for $R=0.5$ (Fig.~\ref{fig:ratio}b),
-344 tesseroids for $R=1$ (Fig.~\ref{fig:ratio}c),
-and 2472 tesseroids for  $R=2$ (Fig.~\ref{fig:ratio}d).
-
-Clearly,
-there is need to determine
-an optimal value of $R$
-that yields the desired accuracy
-in the GLQ integration.
-In the following section,
-we compare the computed gravitational effects
-of a tesseroid model
-with that of a spherical half-shell
-for which exist analytical expressions.
-
-\subsection{Determination of $r$ by comparison with a spherical shell}
-
-The gravitational potential,
-gravitational attraction,
-and
-Marussi (gravity gradient) tensor
-caused by the homogeneous spherical half-shell
-at a computation point
-$P = (\lambda=90^\circ, \phi=0^\circ, r=r_2+h)$
-are
-
-\begin{equation}
-    V(h) = 2\pi G \rho \left[
-        \frac{l^3 + {r'}^3}{3(r_2 + h)} - 0.5 {r'}^2 \right]
-         \Biggr|_{r'=r_1}^{r'=r_2},
-    \label{eq:shellpot}
-\end{equation}
-\begin{equation}
-    g_z(h) = 2\pi G \rho \left[
-        \frac{l^3 + {r'}^3}{3(r_2 + h)^2} - l \right]
-        \Biggr|_{r'=r_1}^{r'=r_2},
-    \label{eq:shellgz}
-\end{equation}
-\begin{equation}
-    g_{xx}(h) = g_{yy}(h) = -\frac{1}{2} g_{zz}(h),
-\end{equation}
-\begin{equation}
-    g_{xy}(h) = g_{xz}(h) = g_{yz}(h) = 0,
-\end{equation}
-
-\noindent
-and
-
-\begin{equation}
-    g_{zz}(h) = 2\pi G \rho \left[
-        2\frac{l^3 + {r'}^3}{3(r_2 + h)^3}
-        - \frac{l}{r_2 + h} + \frac{r_2 + h}{l}
-        \right]
-        \Biggr|_{r'=r_1}^{r'=r_2},
-    \label{eq:shellgzz}
-\end{equation}
-
-\noindent
-where $\rho$ is the density,
-$l = \sqrt{(r_2 + h)^2 + {r'}^2}$,
-and
-$r_1$ and $r_2$ are radial coordinates of
-the bottom and top of the spherical shell,
-respectively.
-
-For our comparison,
-we used a spherical half-shell with
-$\rho=2.7\ g.cm^{-3}$,
-$r_1=6378.137\ km$ (mean Earth radius),
-and
-$r_2 = r_1 + 50\ km$.
-We discretized the half-shell
-into tesseroids of equal size
-and calculated the gravitational effects of
-the half-shell
-(using equations \ref{eq:shellpot}-\ref{eq:shellgzz})
-and the corresponding tesseroid model
-at various heights.
-These calculations were repeated
-for varying values of
-the distance/size ratio $R$.
-This analysis
-allows us to establish
-the optimal value of $R$
-that achieves a desired error level.
-
-Data and source code to produce the results of this comparison
-are provided by \textbf{(cite notebook on figshare)}.
-
-\subsection{Software implementation}
-
-The methods described above
-were implemented the software package
-Tesseroids version 1.1.1.
-The package consists
-of a range of command-line programs
-written in the C programming language.
-It is freely available
-online\footnote{www.leouieda.com/tesseroids}
-under the terms of
-the open-source BSD 3-clause license.
-\textbf{(cite archive on figshare)} provides a persistent
-archive for the source code and compiled binaries of Tesseroids.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
-
-Meh.
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figs/tesseroid-x-shell}
-    \caption{
-        Tesseroid vs spherical shell.
-    }
-    \label{fig:tesseroid-x-shell}
-\end{figure*}
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
-egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
-et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
-tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
-Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
-eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
-fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
-non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
-consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
-convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
-nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
-pellentesque, quis molestie sem vestibulum.
-
 \bibliographystyle{gji}
-\bibliography{references}
-
+\bibliography{references.bib}
 \end{document}

### 2014-11-10 manuscript.tex
A bit of progress on methods
Need to better describe the adaptive discretization algorithm.
Duped by previous authors in dividing only when necessary. It was hidden
and obscured in the pseudo-code of the appendix. Focus on numerical
results.

@@
 
-\title[Tesseroid forward modeling with adaptive discretization]{
-Forward modeling of gravitational fields using tesseroids with adaptive
-discretization
+\title[Quantifying the error in tesseroid forward modeling]{
+Quantifying the error in forward modeling gravitational fields using tesseroids
+with adaptive discretization
 }
@@
 Val\'eria C. F. Barbosa$^{2}$,
-Vanderlei C. Oliveira Jr$^{2}$
+Vanderlei C. Oliveira Jr$^{2}$,
+Carla Braitenberg$^{3}$
 \\
-$^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil.
-$^2$Observat\'orio Nacional, Rio de Janeiro, Brazil.
+$^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil. \\
+$^2$Observat\'orio Nacional, Rio de Janeiro, Brazil. \\
+$^3$Department of Mathematics and Geosciences, University of Trieste, Trieste, Italy.
 }
@@
 
-We propose an improvement over the adaptive discretization algorithm of
-\citet{Li2011} to further increase the computational efficiency.
+We will attempt to quantify the error committed in the computation of the
+gravity gradients using the GLQ with adaptive discretization.
 
-We will also quantify the error committed for the gravity gradients.
+We will establish an empirical relationship between the integration error and
+the ratio between the distance to the computation point and the dimensions of
+the tesseroid.
 
-This is done in an attempt to establish an optimal minimum threshold for the
-ratio between the distance to the computation point and the dimensions of the
-tesseroid.
+This will allow one to choose the appropriate amount of discretization to
+achieve the desired accuracy.
 
@@
 \section{Methodology}
+
+\citet{Grombein2013} provides Cartesian kernels for the tesseroid volume
+internals that define the gravitational potential, gravitational attraction,
+and Marussi tensor (EQUATIONS).
+\begin{equation}
+    V(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        \frac{1}{\ell} \kappa  dr' d\phi' d\lambda',
+    \label{eq:tesspot}
+\end{equation}
+\begin{equation}
+    g_{\alpha}(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        \frac{\Delta\alpha}{\ell^3} \kappa dr' d\phi' d\lambda',
+    \label{eq:tessgrav}
+\end{equation}
+\noindent
+and
+\begin{equation}
+    g_{\alpha\beta}(r,\phi,\lambda) = G \rho
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        I_{\alpha\beta}
+        dr' d\phi' d\lambda',
+    \label{eq:tesstensor}
+\end{equation}
+
+\citet{Grombein2013} follows \citet{Heck2007} and solves the integrals by a
+Taylor expansion around the geometric center of the tesseroid.
+
+We will follow \citet{Asgharzadeh2007} and perform the numerical integration
+using the Gauss-Legendre Quadrature (GLQ) instead.
+
+The GLQ consists of approximating the integral by a weighted sum of the
+integration kernel \citep{Hildebrand1987},
+\begin{equation}
+    \int\limits_a^b f(x) dx \approx
+    \frac{b-a}{2}\sum\limits_{i=1}^N W_i f(x_i),
+    \label{eq:glq1d}
+\end{equation}
+
+$N$ is the number of points used in the GLQ, i.e. the order of the quadrature.
+
+The points $x_i$ where the kernel is evaluated are called the quadrature nodes.
+
+The points are roots of the $N^{th}$ order Legendre polynomial $P_N(x)$.
+
+The roots are $x = \pm 0.577350269$ for a second order polynomial ($P_2(x)$).
+
+A root finder algorithm can be used to determine the roots for larger order
+polynomials.
+
+We refer the reader to \citet{Barrera-Figueroa2006} for an algorithm tailored
+for Legendre polynomials in particular.
+
+The second order roots, as well as the ones determined by a root finder
+algorithm, will be within the range $[-1, 1]$.
+
+Thus, the roots must be scaled to the integration limits $[a, b]$ by
+\begin{equation}
+    x^{scaled}_i = \frac{b - a}{2} x_i + \frac{b + a}{2}.
+\end{equation}
+
+The weights of the GLQ are given by \citep{Hildebrand1987},
+\begin{equation}
+    W_i = \frac{2}{(1 - x_i^2)(P^\prime_N(x_i))^2},
+\end{equation}
+
+The values of the $P_N(x)$ and its first derivative $P^\prime_N(x)$ can be
+calculated with recursive relations (PUT FORMULA HERE?).
+
+The Gauss-Legendre Quadrature for volume integrals, like equations
+\ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
+becomes \citep{Asgharzadeh2007} (EQUATION).
+
+Applying equation (GLQ 3D) to equations
+\ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
+we see that $f(r_i, \phi_j, \lambda_k)$ is the effect of a point
+mass located on the quadrature nodes.
+
+It can be said that the GLQ integration approximates the volume integrals  by a
+weighted sum of point mass effects.
+
+\section{Adaptive discretization algorithm}
+
+The quantity of, and therefore distance between, point masses regulates the
+accuracy of the numerical integration \citep{Ku1977}.
+
+\citet{Ku1977} determined an empirical criterion that the distance between
+point masses (quadrature nodes) should be smaller than the minimum distance to
+the computation point.
+
+Thus, if a computation point is too close to the tesseroid one would have to
+decrease the distance between the point masses in order to obtain an accurate
+result.
+
+One way to do this would be increase the order of the quadrature $N$ in all
+three dimensions.
+
+An alternative way would be to fix the quadrature order (e.g., $N=2$) and
+instead divide the tesseroid in half along each dimension.
+
+The gravitational effect of the larger tesseroid would be calculated as the sum
+of the effects of the eight smaller tesseroids.
+
+This division would effectively decrease the distance between nodes because of
+the smaller tesseroids.
+
+\citet{Li2011} developed an algorithm based on the tesseroid division approach.
+
+In their algorithm, a tesseroid is divided in half along one of its dimensions
+if the minimum distance to the computation point is smaller than a multiple of
+the size of the tesseroid along that dimension.
+
+The criterion check and (possible) subsequent division is repeated recursively
+(i.e., to each of the smaller tesseroids and so on).
+
+The recursion stops when none of the smaller tesseroids fail the criterion.
+
+The advantage of this adaptive discretization is that the number of point
+masses is only increased in regions of the tesseroid that are closer to the
+computation point.
+
+Alternatively, simply increasing the order of the GLQ would increase the number
+of point masses evenly throughout the whole tesseroid.
+
+The key controlling point of their algorithm is the ratio between
+the smallest distance to the computation point and
+the largest side length of the tesseroid.
+
+The value of this ratio is used to decide whether a tesseroid will be divided
+or not.
+
+Thus, the value chosen for the minimum allowed value of this ratio determines
+how many divisions are made.
+
+Consequently, the value for the minimum distance-size ratio controls both the
+accuracy of the integration and the computation time.
+
+\citet{Li2011} provide a few examples of values for this ratio in the case of
+the vertical component of gravitational attraction.
+
+However, there has been no further investigation into
+the relationship between values of the minimum distance-size
+ratio and the computation error.
+
+It is also not clear whether the same relationship holds for the gravitational
+attraction as well as for the gradient tensor components.
 

### 2014-11-14 manuscript.tex
Moved to Geophysics again. Will focus on code
Moving the focus of the paper to the algorithm and implementation. This
will fit nicely on the Software section of Geophysics.

@@
-\documentclass[extra,mreferee]{gji}
-%\documentclass[extra,paper]{gji}
-\usepackage{timet}
+\documentclass[paper,twocolumn]{geophysics}
+%\documentclass[manuscript]{geophysics}
+%\documentclass[manuscript,endfloat]{geophysics}
 \usepackage{graphicx}
-
-\title[Quantifying the error in tesseroid forward modeling]{
-Quantifying the error in forward modeling gravitational fields using tesseroids
-with adaptive discretization
-}
-
-\author[Uieda et al.]{
-Leonardo Uieda$^{1,2}$,
-Val\'eria C. F. Barbosa$^{2}$,
-Vanderlei C. Oliveira Jr$^{2}$,
-Carla Braitenberg$^{3}$
-\\
-$^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil. \\
-$^2$Observat\'orio Nacional, Rio de Janeiro, Brazil. \\
-$^3$Department of Mathematics and Geosciences, University of Trieste, Trieste, Italy.
-}
+\newcommand{\mitbf}[1]{\mathbf{{#1}}}
 
@@
 
-\label{firstpage}
+\title{Title}
+\lefthead{Meh}
+\righthead{Meh}
+% manuscript number
+\ms{Submission}
+
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+\address{
+\footnotemark[1] Universidade do Estado do Rio de Janeiro,
+Rio de Janeiro, Brazil.
+\\
+\footnotemark[2] Observat\'orio Nacional,
+Rio de Janeiro, Brazil.
+\\
+\footnotemark[3] Department of Mathematics and Geosciences,
+University of Trieste, Trieste, Italy.
+}
+
+\author{
+    Leonardo Uieda\footnotemark[1]\footnotemark[2]
+    Val\'eria C. F. Barbosa\footnotemark[2],
+    Vanderlei C. Oliveira Jr\footnotemark[2],
+    and
+    Carla Braitenberg\footnotemark[3]
+}
+
 \maketitle
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methodology}
+\section{Theory}
 
@@
 
-\citet{Grombein2013} follows \citet{Heck2007} and solves the integrals by a
-Taylor expansion around the geometric center of the tesseroid.
-
 We will follow \citet{Asgharzadeh2007} and perform the numerical integration
-using the Gauss-Legendre Quadrature (GLQ) instead.
+using the Gauss-Legendre Quadrature (GLQ).
 
@@
 
-$N$ is the number of points used in the GLQ, i.e. the order of the quadrature.
+$N$ is the order of the quadrature, i.e. the number of points used in the GLQ.
 
@@
 
-\section{Adaptive discretization algorithm}
+The accuracy of the integration
+depends on the number of point masses used in the summation.
 
-The quantity of, and therefore distance between, point masses regulates the
-accuracy of the numerical integration \citep{Ku1977}.
+\citet{Ku1977} showed that it also depends on the ratio between
+the distance to the computation point and the distance between adjacent nodes.
+
+Figure~\ref{fig:glqerrorsample}
+illustrates this effect on the $g_{xy}$ gravity gradient component.
+
+The $g_{xy}$ component was produced by a
+$7^\circ \times 7^\circ \times 20\ km$ tesseroid
+with $2.67\ g.cm^{-3}$ density
+and top at $z=0\ km$.
+
+The maps were calculated on a regular grid
+with $100\times100$ points.
+
+The GLQ order used was the same
+in the radial, longitudinal, and latitudinal dimensions.
+
+Figure~\ref{fig:glqerrorsample}a shows the $g_{xy}$ component
+calculated at 400 km height using
+GLQ with order two (two point masses in the longitude, latitude, and vertical
+directions, respectively).
+
+Figure~\ref{fig:glqerrorsample}b shows $g_{xy}$ computed with order two
+GLQ as well but at 150 km height.
+
+Notice that the computed effect is concentrated around each point mass (black
+dots) and does not resemble the effect of a tesseroid.
 
@@
 
-One way to do this would be increase the order of the quadrature $N$ in all
-three dimensions.
+One way to accomplish this would be increase the order of the quadrature
+$N$ in all three directions.
 
-An alternative way would be to fix the quadrature order (e.g., $N=2$) and
-instead divide the tesseroid in half along each dimension.
+Figure~\ref{fig:glqerrorsample}c shows the $g_{xy}$ component calculated at
+150km height but with a GLQ order of 30.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figs/vary-height-and-order}
+    \caption{
+        Example of the effect of varying
+        the computation height
+        and the number of point masses in the Gauss-Legendre Quadrature
+        (i.e., the GLQ order).
+        Black circles represent the horizontal location of the GLQ nodes
+        (not shown for GLQ order 30).
+        a) Field calculated at $400\ km$ height using GLQ order 2.
+        The effect is similar to that presented by \citet{Asgharzadeh2007}.
+        b) At $150\ km$ height and order 2,
+        the field resembles that of
+        four point sources located at the GLQ nodes.
+        This effect was shown by \citet{Ku1977}
+        for the case of the $g_z$ component of a right rectangular prism.
+        c) At $150\ km$ but with a higher GLQ order of 30,
+        the field is similar to that expected for a single mass source.
+    }
+    \label{fig:glqerrorsample}
+\end{figure}
+
+\subsection{Adaptive discretization}
+
+\citet{Li2011} proposed an alternative way to decrease the distance between
+point masses and achieve an accurate integration.
+
+Instead of increasing the GLQ order,
+they divide the tesseroid into smaller tesseroids while keeping the number of
+point masses per tesseroid fixed.
 
@@
 
-\citet{Li2011} developed an algorithm based on the tesseroid division approach.
+The criterion for dividing a tesseroid is that
+the distance to the computation point must be smaller than
+a constant times the size of the tesseroid.
 
-In their algorithm, a tesseroid is divided in half along one of its dimensions
-if the minimum distance to the computation point is smaller than a multiple of
-the size of the tesseroid along that dimension.
+The size of the tesseroid is used as proxy for the distance between point
+masses.
 
-The criterion check and (possible) subsequent division is repeated recursively
-(i.e., to each of the smaller tesseroids and so on).
-
-The recursion stops when none of the smaller tesseroids fail the criterion.
+This procedure is repeated recursively
+(i.e., to each of the smaller tesseroids)
+until all tesseroids are within the acceptable distance-size ratio
+or a minimum size is achieved.
 
@@
 of point masses evenly throughout the whole tesseroid.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Numerical evaluation of the accuracy}
 
@@
 
-\bibliographystyle{gji}
+\bibliographystyle{seg}
 \bibliography{references.bib}

### 2014-11-14 manuscript.tex
Added some figures to the text

@@
 
+A tesseroid is a spherical mass element defined in geocentric coordinates
+(Figure~\ref{fig:tesseroid}).
+It is bounded by two meridians, two parallels, and two concentric circles.
+The gravitational fields of a tesseroid at a point $P = (r,\phi,\lambda)$
+are determined with respect to the local North-oriented coordinate system at
+P (x, y, z in Figure~\ref{fig:tesseroid}).
+Notice that this is a left-handed coordinate system.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{figs/tesseroid}
+    \caption{
+        View of a tesseroid,
+        the integration point $Q$ inside the tesseroid,
+        a geocentric coordinate system $(X, Y, Z)$,
+        the computation $P$ and it's local coordinate system $(x, y, z)$.
+        $r$, $\phi$, $\lambda$ are
+        the radius, latitude, and longitude, respectively, of point $P$,
+        and $\ell$ is the Cartesian distance between $P$ and $Q$.
+    }
+    \label{fig:tesseroid}
+\end{figure}
+
 \citet{Grombein2013} provides Cartesian kernels for the tesseroid volume
 internals that define the gravitational potential, gravitational attraction,
-and Marussi tensor (EQUATIONS).
+and Marussi tensor, respectively,
+
 \begin{equation}
@@
         I_{\alpha\beta}
-        dr' d\phi' d\lambda',
+        dr' d\phi' d\lambda'.
     \label{eq:tesstensor}
@@
     \centering
-    \includegraphics[width=\textwidth]{figs/vary-height-and-order}
+    \includegraphics[width=0.4\textwidth]{figs/vary-height-and-order}
     \caption{
@@
 
+\begin{figure}
+    \centering
+    \includegraphics{figs/tesseroid-split}
+    \caption{
+        Adaptive discretization
+        of the tesseroid shown in (a)
+        for a computation point P
+        using the distance-size ratio $D$ equal to
+        (b) 1, (c) 2, and (d) 6.
+        $d$ is the Cartesian distance between
+        the center of the top face of the tesseroid
+        and $P$.
+        $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the tesseroid.
+        Note that increasing $D$
+        results in a fine division of the tesseroid
+        close the computation point
+        and a coarser division further away.
+    }
+    \label{fig:ratio}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Numerical evaluation of the accuracy}
+\section{Evaluation of the accuracy}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}

### 2014-11-14 manuscript.tex
Fixed some typos

@@
 
-Notice that the computed effect is concentrated around each point mass (black
-dots) and does not resemble the effect of a tesseroid.
+Notice that the computed effect is concentrated around each point mass
+of the GLQ (black dots) and does not resemble the effect of a tesseroid.
 
@@
 150km height but with a GLQ order of 30.
+
+The computed $g_{xy}$ component more closely resembles the results from
+\citet{Asgharzadeh2007}.
 
@@
 
-The size of the tesseroid is used as proxy for the distance between point
+The size of the tesseroid is used as a proxy for the distance between point
 masses.

### 2014-11-14 manuscript.tex
Introduction to the implementation section

@@
 
-\renewcommand{\thefootnote}{\fnsymbol{footnote}}
-\address{
-\footnotemark[1] Universidade do Estado do Rio de Janeiro,
-Rio de Janeiro, Brazil.
-\\
-\footnotemark[2] Observat\'orio Nacional,
-Rio de Janeiro, Brazil.
-\\
-\footnotemark[3] Department of Mathematics and Geosciences,
-University of Trieste, Trieste, Italy.
-}
-
 \author{
@@
     Carla Braitenberg\footnotemark[3]
+}
+
+\address{
+\footnotemark[1]Universidade do Estado do Rio de Janeiro,
+Rio de Janeiro, Brazil.
+email: leouieda@gmail.com
+\\
+\footnotemark[2]Observat\'orio Nacional,
+Rio de Janeiro, Brazil.
+\\
+\footnotemark[3]Department of Mathematics and Geosciences,
+University of Trieste, Trieste, Italy.
 }
@@
 
+The calculation of the tesseroid gravitational fields with adaptive
+discretization is implemented in the open-source package \textit{Tesseroids}.
+
+Tesseroids is freely available under the BSD 3-clause free-software license.
+
+An archived version of the source code is available along with this article.
+
+The latest version can be downloaded from the official
+website\footnote{http://www.leouieda.com/tesseroids}.
+
+Ongoing development takes place through the source code hosting site
+Github\footnote{https://github.com/leouieda/tesseroids}.
+
+\textit{Tesseroids} is written in the C programming language.
+
+Tesseroids consists of command-line programs.
+
+Tesseroids tries to follow the Unix principals of
+(1) making each program do one thing well,
+(2) using plain text for input and output,
+(3) making each program a ``filter'' that can be used in conjunction with other
+programs.
+
+The gravitational potential, attraction, and Marussi tensor components
+are each computed by a separate program.
+
+Programs for model and computation grid generation are also included.
+
+The package also includes programs for computing the gravitational fields of
+right rectangular prisms in Cartesian and spherical coordinates.
+
+Tesseroids programs receive input through command-line arguments and the
+standard input channel (STDIN) and output the results through
+the standard output channel (STDOUT).
+
+The command-line interface of each program was designed so that they
+can be called in sequence using a Unix pipe.
+
+For example, the command to
+generate a regular grid with $NLON \times NLAT$ points,
+calculate on it the $g_z$ and $g_{zz}$ caused by the tesseroids in file
+MODELFILE,
+and save the results to a file called OUTPUT
+is:
+
+\begin{verbatim}
+tessgrd -rW/E/S/N -bNLON/NLAT -zHEIGHT | \
+    tessgz MODELFILE | \
+    tessgzz MODELFILE > OUTPUT
+\end{verbatim}
+
+The \texttt{src} folder of the distribution archive
+contains the \texttt{.c} files that build the command-line programs.
+
+The \texttt{src/lib} folder contains the source files that implement the
+numerical computations.
+
+We will not describe here the implementation of the input/output parsing and
+other miscellanea.
+
+Instead, we will focus on the details of the Gauss-Legendre Quadrature
+integration of equations~\ref{eq:tesspot},~\ref{eq:tessgrav},
+and~\ref{eq:tesstensor}
+and the adaptive discretization of tesseroids.
+
+\subsection{Numerical integration}
+
+The first step in the numerical integration with the GLQ is to compute the
+locations of the point masses (i.e., the discretization points).
+
+These points are roots of Legendre polynomials.
+
+We use the
+
+\subsection{Implementation of adaptive discretization}
+
 \begin{figure}

### 2014-11-14 manuscript.tex
Numerical integration section draft

@@
     x^{scaled}_i = \frac{b - a}{2} x_i + \frac{b + a}{2}.
+    \label{eq:glq_scaling}
 \end{equation}
@@
     W_i = \frac{2}{(1 - x_i^2)(P^\prime_N(x_i))^2},
+    \label{eq:glq_weights}
 \end{equation}
@@
 The values of the $P_N(x)$ and its first derivative $P^\prime_N(x)$ can be
-calculated with recursive relations (PUT FORMULA HERE?).
+calculated with recursive relations.
 
@@
 
-The latest version can be downloaded from the official
-website\footnote{http://www.leouieda.com/tesseroids}.
+The official website\footnote{http://www.leouieda.com/tesseroids}
+hosts the software documentation and
+archives of the latest binary executables and source code.
 
@@
 The \texttt{src} folder of the distribution archive
-contains the \texttt{.c} files that build the command-line programs.
+contains the C files that build the command-line programs
+(e.g., \texttt{tessgz.c}).
 
@@
 
-The first step in the numerical integration with the GLQ is to compute the
+The source file \texttt{src/lib/glq.c} contains the code necessary to perform a
+Gauss-Legendre Quadrature integration.
+
+The first step in the GLQ is to compute the
 locations of the point masses (i.e., the discretization points).
@@
 
-We use the
+Precomputed values for these roots are available low order polynomials,
+usually up to order 5.
+
+For flexibility, we use the multiple root-finder algorithm of
+\citet{Barrera-Figueroa2006} to compute the roots for polynomials of any order.
+
+The additional computational load is minimal since the root-finder algorithm
+must be run only once per program execution.
+
+The root-finder algorithm is implemented in functions \texttt{glq\_nodes} and
+\texttt{glq\_next\_root}.
+
+The computed roots will be in the range $[-1, 1]$
+and must be scaled to the integration limits (the limits of the tesseroid)
+using function \texttt{glq\_set\_limits} (see equation~\ref{eq:glq_scaling}).
+
+The GLQ weights are computed using equation~\ref{eq:glq_weights} by function
+\texttt{glq\_weights}.
+
+Both the computed roots and weights are stored in a data structure (a C
+\texttt{struct}) called \texttt{GLQ}.
+
+Function \texttt{glq\_new} allocates the memory necessary,
+calculates the roots and weights,
+and returns the complete \texttt{GLQ} structure.
+
+The numerical integration is performed by the functions in module
+\texttt{src/lib/grav\_tess.c}.
+
+The functions \texttt{tess\_pot}, \texttt{tess\_gx}, \texttt{tess\_gy},
+and so on, compute the gravitational fields of a single tesseroid on a single
+computation point using the GLQ.
+
+These functions require three \texttt{GLQ} structures with roots scaled to the
+integration limits $[\lambda_1, \lambda_2], [\phi_1, \phi_2], [r_1, r_2]$.
+
+The integration consists of three \texttt{for} loops: one for longitude, one
+for latitude, and one for radius.
+
+The evaluation of the trigonometric functions in
+equations~\ref{eq:tesspot}-\ref{eq:tesstensor} inside the inner loops
+are the biggest bottlenecks for the numerical integration.
+
+Better performance is achieved by moving some trigonometric function
+evaluations to the outer loops
+and precomputing the sine and cossine of latitudes.
 

### 2014-11-15 manuscript.tex
Algorithm description done
Need to tie it with the calc_tess_model_adapt function

@@
 \usepackage{graphicx}
+\usepackage{amsmath}
 \newcommand{\mitbf}[1]{\mathbf{{#1}}}
@@
 
-\title{Title}
-\lefthead{Meh}
-\righthead{Meh}
+\title{
+Tesseroids: forward modeling gravitational fields in spherical coordinates
+}
+\righthead{Tesseroids}
 % manuscript number
@@
 \author{
-    Leonardo Uieda\footnotemark[1]\footnotemark[2]
+    Leonardo Uieda\footnotemark[1]\footnotemark[2],
     Val\'eria C. F. Barbosa\footnotemark[2],
-    Vanderlei C. Oliveira Jr\footnotemark[2],
+    Carla Braitenberg\footnotemark[3],
     and
-    Carla Braitenberg\footnotemark[3]
+    Vanderlei C. Oliveira Jr\footnotemark[2]
 }
@@
 These functions require three \texttt{GLQ} structures with roots scaled to the
-integration limits $[\lambda_1, \lambda_2], [\phi_1, \phi_2], [r_1, r_2]$.
+integration limits $[\lambda_1, \lambda_2], [\phi_1, \phi_2], [r_1, r_2]$
+(see equations~\ref{eq:tesspot}-\ref{eq:tesstensor}).
 
@@
 
-Better performance is achieved by moving some trigonometric function
-evaluations to the outer loops
-and precomputing the sine and cossine of latitudes.
+Better performance is achieved by
+precomputing the sine and cossine of latitudes
+and moving some trigonometric function
+evaluations to the outer loops.
 
 \subsection{Implementation of adaptive discretization}
+
+The adaptive discretization algorithm of \citet{Li2011} recursively
+divides a tesseroid.
+
+A tesseroid will be divided when the smallest distance between it and the
+computation point is smaller than a constant times the largest dimension of the
+tesseroid.
+
+In our implementation,
+we measure the distance between
+the computation point
+and the center of the top face tesseroid
+(see Figure~\ref{fig:division}a).
+
+The distance $d$ is
+
+\begin{equation}
+    d = \left[
+        r^2 + r_2^2 - 2 r r_2
+        \cos(\sin\phi\sin\Delta\phi + \cos\phi\cos\Delta\phi
+            \cos(\lambda - \Delta\lambda))
+        \right],
+    \label{eq:distance}
+\end{equation}
+
+where
+$(r, \lambda, \phi)$ are the coordinates of the computation point,
+$\Delta\lambda = 0.5(\lambda_2 + \lambda_1)$,
+and
+$\Delta\phi = 0.5(\phi_2 + \phi_1)$.
+
+We define the dimensions of the tesseroid
+along longitude, latitude, and radius, respectively, as
+
+\begin{equation}
+    L_\lambda = r_2 \arccos(\sin^2\Delta\phi +
+        \cos^2\Delta\phi\cos(\lambda_2 - \lambda_1)),
+    \label{eq:sizelon}
+\end{equation}
+\begin{equation}
+    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1),
+\end{equation}
+\begin{equation}
+    L_r = r_2 - r_2.
+    \label{eq:sizer}
+\end{equation}
+
+We fix the GLQ order $N=2$ for all three dimensions, resulting in eight point
+masses in total.
+
+The criterion we use to determine if a tesseroid must be divided is
+
+\begin{equation}
+    \frac{d}{L_i} < D,
+    \label{eq:condition}
+\end{equation}
+\noindent
+where $i \in (\lambda, \phi, r)$
+and $D$ is a positive scalar called the ``distance-size ratio''.
+$D$ determines how close the
+computation point can be before we must divide the tesseroid.
+
+We used a stack-based implementation of the algorithm instead of using
+recursive function calls.
+
+Stacks are array-like data structures with a particular way of inserting and
+removing elements from it.
+
+In a stack, one can only insert elements on the top (the last empty position).
+
+Likewise,
+one can only remove the last element that was placed on the stack
+(commonly referred to as ``popping'' the stack).
+
+Because of this stacks are also known as ``Last-in-first-out'', or LIFO,
+structures.
+
+LIFOs are often described as being analogous to a stack of plates.
+
+One can only place a new plate on the top of the stack and
+can only remove a plate safely from the top of the stack.
+
+The algorithm is described for a single tesseroid and computation point.
+
+The recursive division algorithm implemented with a stack is as follows.
+
+Start by initializing the stack with the tesseroid.
+
+Thus, so far, the stack has only one element.
+
+Once the stack is initialized, the steps of the algorithm are:
+
+\begin{enumerate}
+    \item ``Pop'' the stack (i.e., take the last element from it).
+    \item Compute the distance $d$ between the tesseroid and
+        the computation point (equation~\ref{eq:distance}).
+    \item Compute the dimensions of the tesseroid $L_\lambda$, $L_\phi$,
+        and $L_r$ (equations~\ref{eq:sizelon}-\ref{eq:sizer}).
+    \item Check the condition in equation~\ref{eq:condition} for each
+        dimension of the tesseroid.
+    \item If none of the dimensions fail the above condition,
+        compute the gravitational effect of the tesseroid using the
+        Gauss-Legendre Quadrature with order two
+        (equation~\ref{eq:glq3d}).
+    \item If any of the dimensions fail the condition,
+        divide the tesseroid in half along each dimension that failed the
+        condition.
+        Then, place the smaller tesseroids into the stack.
+    \item Repeat the above steps until the stack is empty.
+\end{enumerate}
 
@@
     }
-    \label{fig:ratio}
+    \label{fig:division}
 \end{figure}

### 2014-11-18 manuscript.tex
Cite IPython. Connect algorithm with code
Mention the code parts in the algorithm description. Mention notebooks
for reproducing results and cite IPython.

@@
 
+All figures and numerical experiments in this article were produced by the
+IPython notebook\footnote{http://ipython.org}
+files \citep{Perez2007} in the supplementary material.
+
+These files combine the analysis source code
+with explanatory text, equations, and the figures generated by the code.
+
+This supplementary material can be run and executed locally or viewed
+online\footnote{https://github.com/pinga-lab/paper-tesseroids}
+(in a non-executable form).
+
 \subsection{Numerical integration}
@@
 and $D$ is a positive scalar called the ``distance-size ratio''.
+
 $D$ determines how close the
@@
 
+Thus, the value of $D$ is indirectly responsible for
+the accuracy of the solution and the computation time.
+
+The modeling programs provide default values of $D$.
+
+It can also be controlled through command-line arguments passed to the
+programs.
+
 We used a stack-based implementation of the algorithm instead of using
-recursive function calls.
+recursive function calls, as  originally proposed by \citet{Li2011}.
 
@@
 
-The algorithm is described for a single tesseroid and computation point.
+We will describe the stack-based algorithm for a single tesseroid and
+computation point.
 
-The recursive division algorithm implemented with a stack is as follows.
+The algorithm is implemented in function \texttt{calc\_tess\_model\_adapt}
+of the file \texttt{src/lib/grav\_tess.c}.
 
-Start by initializing the stack with the tesseroid.
+Start by creating an empty stack of tesseroids.
 
-Thus, so far, the stack has only one element.
+The  stack is represented by the \texttt{stack} variable,
+an array of \texttt{TESSEROID} structures.
+
+We must define a maximum size for the stack.
+
+We use the integer \texttt{stktop} to keep track of the index of the last
+element in the stack (the top of the stack).
+
+The stack is initialized with the input tesseroid.
+
+The initialization is done by copying the tesseroid into the stack and
+setting \texttt{stktop} to zero (the first element).
 
@@
     \item ``Pop'' the stack (i.e., take the last element from it).
+        This will cause \texttt{stktop} to be reduced by one.
     \item Compute the distance $d$ between the tesseroid and
@@
         (equation~\ref{eq:glq3d}).
-    \item If any of the dimensions fail the condition,
-        divide the tesseroid in half along each dimension that failed the
-        condition.
-        Then, place the smaller tesseroids into the stack.
-    \item Repeat the above steps until the stack is empty.
+    \item If any of the dimensions fail the condition:
+    \begin{enumerate}
+        \item Divide the tesseroid in half along each dimension that failed
+             the condition.
+        \item Check if the number new (smaller) tesseroids fit in the stack.
+             In other words, the number of new elements plus \texttt{stktop}
+             must be smaller than the maximum stack size.
+             If the stack is full, warn the user of a ``stack overflow''
+             and compute the effect of the larger tesseroid (step 5).
+        \item If the stack is not full, place the smaller tesseroids into
+             the stack.
     \end{enumerate}
+    \item Repeat the above steps until the stack is empty
+        (\texttt{stktop = -1}).
+\end{enumerate}
+
+This stack-based implementation has some advantages over the original recursive
+implementation suggested by \citet{Li2011}.
+
+(1) It will be faster because it bypasses the overhead of function calls.
+
+The speedup will be specially significant if the implementation is done
+in high level interpreted languages, like Matlab and Python.
+
+(2) It gives the developer more control over the recursion step.
+
+In recursive implementations,
+the developer has no control over the maximum number of consecutive
+recursive calls (the ``recursion depth'').
+
+This limit may vary with programming language, compiler, and operating system.
+
+Overflowing the maximum recursion depth usually results in program
+crashes with cryptic or inexistent error messages.
+
+In the stack-based implementation, the developer has complete control.
+
+Overflowing of the stack can be handled gracefully with an error message or
+even performing a suitable approximation of the result.
 

### 2014-11-20 manuscript.tex
Reminder to include a section on unit tests
Reminder to add subsect: gzz x height and time x D
Describe the division figure in the text

@@
-\documentclass[paper,twocolumn]{geophysics}
-%\documentclass[manuscript]{geophysics}
+%\documentclass[paper,twocolumn]{geophysics}
+\documentclass[manuscript]{geophysics}
 %\documentclass[manuscript,endfloat]{geophysics}
@@
             \cos(\lambda - \Delta\lambda))
-        \right],
+        \right]^{\frac{1}{2}},
     \label{eq:distance}
@@
 \end{equation}
+
+USE THE MAX OF SOUTH AND NORTH FOR DLON? MIGHT MAKE IT BETTER IN THE POLES.
 
@@
 
-Because of this stacks are also known as ``Last-in-first-out'', or LIFO,
-structures.
+Stacks are also known as ``Last-in-first-out'', or LIFO, structures.
 
@@
 
-We must define a maximum size for the stack.
+We must define a maximum size for the stack to allocate memory for it.
+
+Defining a maximum size allows us to avoid an infinite loop and large
+consumption of computer memory in case the computation point is
+on (or sufficiently close to) the surface of the tesseroid.
 
@@
              the condition.
-        \item Check if the number new (smaller) tesseroids fit in the stack.
-             In other words, the number of new elements plus \texttt{stktop}
-             must be smaller than the maximum stack size.
-             If the stack is full, warn the user of a ``stack overflow''
+        \item Check if the number of new elements plus \texttt{stktop}
+             is smaller than the maximum stack size.
+             If it is not (i.e., the stack is full),
+             warn the user of a ``stack overflow''
              and compute the effect of the larger tesseroid (step 5).
@@
     \item Repeat the above steps until the stack is empty
-        (\texttt{stktop = -1}).
+        (\texttt{stktop} is equal to -1).
 \end{enumerate}
@@
 
+Figure~\ref{fig:division} shows the effect of the distance-size ration $D$
+in the adaptive discretization of a tesseroid for the computation at a point P.
+
+Computing the gravitational effect of the tesseroid in
+Figure~\ref{fig:division}a using the algorithm described above
+is equivalent to computing the effect of the tesseroids in
+Figures~\ref{fig:division}b-d.
+
+Note that increasing the value of $D$ results in a finer mesh of tesseroids.
+
+For $D = 1$ (Figure~\ref{fig:division}b) the number of tesseroids in the
+adaptive discretization is 12,
+for $D = 2$ (Figure~\ref{fig:division}c) it is 52,
+and for $D = 6$ (Figure~\ref{fig:division}d) it is 1092.
+
 \begin{figure}
@@
         and a coarser division further away.
+        The number of tesseroids in the discretization is (b) 12, (c) 52,
+        and (d) 1092.
     }
@@
 
+\subsection{$g_{zz}$ variability with computation height}
+
+\subsection{Computational performance as a function of $D$}
+
+For close P, vary $D$ and graph computation time.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-11-21 manuscript.tex
Divide figure using dist to center of tesseroid
Matches the latest change in Tesseroids.

@@
-%\documentclass[paper,twocolumn]{geophysics}
-\documentclass[manuscript]{geophysics}
+\documentclass[paper,twocolumn]{geophysics}
+%\documentclass[manuscript]{geophysics}
 %\documentclass[manuscript,endfloat]{geophysics}
@@
 
-These functions require three \texttt{GLQ} structures with roots scaled to the
+These functions require three \texttt{GLQ} structures that contain the roots
+and weights for GLQ integration in the three dimensions.
+
+The roots must be scaled to the
 integration limits $[\lambda_1, \lambda_2], [\phi_1, \phi_2], [r_1, r_2]$
@@
 
-The integration consists of three \texttt{for} loops: one for longitude, one
-for latitude, and one for radius.
+The integration consists of three \texttt{for} loops that sum the weighted
+kernel functions evaluated at each GLQ point mass (the scaled roots).
 
@@
 
-The adaptive discretization algorithm of \citet{Li2011} recursively
-divides a tesseroid.
+The adaptive discretization algorithm of \citet{Li2011}
+divides a tesseroid recursively.
 
@@
 \end{equation}
+
+UPDATE THE EQUATION TO THE CENTER OF THE TESSEROID + NUMBER OF DIVISION IN
+EXAMPLE FIGURE.
 

### 2014-11-26 manuscript.tex
Correct distance and size formula
They now reflect the distance to the center of the tesseroid. Fixed the
number of divisions for the division example.

@@
 
-In our implementation,
-we measure the distance between
-the computation point
-and the center of the top face tesseroid
-(see Figure~\ref{fig:division}a).
-
-The distance $d$ is
+Instead of the smallest distance,
+we use in our implementation
+the distance $d$ between
+the computation point $(r, \lambda, \phi)$
+and the geometric center of the tesseroid
+$(r_t, \lambda_t, \phi_t)$
 
@@
     d = \left[
-        r^2 + r_2^2 - 2 r r_2
-        \cos(\sin\phi\sin\Delta\phi + \cos\phi\cos\Delta\phi
-            \cos(\lambda - \Delta\lambda))
-        \right]^{\frac{1}{2}},
+        r^2 + r_t^2 - 2 r r_t
+        \cos(\sin\phi\sin\phi_t + \cos\phi\cos\phi_t
+            \cos(\lambda - \lambda_t))
+        \right]^{\frac{1}{2}}.
     \label{eq:distance}
@@
 
-UPDATE THE EQUATION TO THE CENTER OF THE TESSEROID + NUMBER OF DIVISION IN
-EXAMPLE FIGURE.
-
-where
-$(r, \lambda, \phi)$ are the coordinates of the computation point,
-$\Delta\lambda = 0.5(\lambda_2 + \lambda_1)$,
-and
-$\Delta\phi = 0.5(\phi_2 + \phi_1)$.
-
 We define the dimensions of the tesseroid
 along longitude, latitude, and radius, respectively, as
+(see Figure~\ref{fig:division}a)
 
 \begin{equation}
-    L_\lambda = r_2 \arccos(\sin^2\Delta\phi +
-        \cos^2\Delta\phi\cos(\lambda_2 - \lambda_1)),
+    L_\lambda = r_2 \arccos(\sin^2\phi_t +
+        \cos^2\phi_t\cos(\lambda_2 - \lambda_1)),
     \label{eq:sizelon}
@@
 \end{equation}
-
-USE THE MAX OF SOUTH AND NORTH FOR DLON? MIGHT MAKE IT BETTER IN THE POLES.
 
@@
 For $D = 1$ (Figure~\ref{fig:division}b) the number of tesseroids in the
-adaptive discretization is 12,
-for $D = 2$ (Figure~\ref{fig:division}c) it is 52,
-and for $D = 6$ (Figure~\ref{fig:division}d) it is 1092.
+adaptive discretization is 4,
+for $D = 2$ (Figure~\ref{fig:division}c) it is 38,
+and for $D = 6$ (Figure~\ref{fig:division}d) it is 936.
 
@@
         and a coarser division further away.
-        The number of tesseroids in the discretization is (b) 12, (c) 52,
-        and (d) 1092.
     }
@@
 
-\subsection{Computational performance as a function of $D$}
-
-For close P, vary $D$ and graph computation time.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-11-26 manuscript.tex
Filled in missing equations from the methods

@@
 
-A tesseroid is a spherical mass element defined in geocentric coordinates
-(Figure~\ref{fig:tesseroid}).
-It is bounded by two meridians, two parallels, and two concentric circles.
-The gravitational fields of a tesseroid at a point $P = (r,\phi,\lambda)$
-are determined with respect to the local North-oriented coordinate system at
-P (x, y, z in Figure~\ref{fig:tesseroid}).
-Notice that this is a left-handed coordinate system.
-
 \begin{figure}
@@
 
-\citet{Grombein2013} provides Cartesian kernels for the tesseroid volume
-internals that define the gravitational potential, gravitational attraction,
+A tesseroid is a spherical mass element defined in geocentric spherical
+coordinates
+(Figure~\ref{fig:tesseroid}).
+It is bounded by two meridians, two parallels, and two concentric circles.
+The gravitational fields of a tesseroid at a point $P = (r,\phi,\lambda)$
+are determined with respect to the local North-oriented coordinate system at
+P (x, y, z in Figure~\ref{fig:tesseroid}).
+\citet{Grombein2013} formulated Cartesian kernels for the volume integrals
+that define the tesseroid gravitational potential, gravitational attraction,
 and Marussi tensor, respectively,
@@
         \int\limits_{r_1}^{r_2}
-        \frac{1}{\ell} \kappa  dr' d\phi' d\lambda',
+        \frac{1}{\ell}
+        \kappa\  dr^\prime d\phi^\prime d\lambda^\prime,
     \label{eq:tesspot}
@@
         \int\limits_{r_1}^{r_2}
-        \frac{\Delta\alpha}{\ell^3} \kappa dr' d\phi' d\lambda',
+        \frac{\Delta_\alpha}{\ell^3}
+        \kappa\ dr^\prime d\phi^\prime d\lambda^\prime,
     \label{eq:tessgrav}
@@
         \int\limits_{r_1}^{r_2}
-        I_{\alpha\beta}
-        dr' d\phi' d\lambda'.
+        \left(
+            \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
+            \frac{\delta_{\alpha\beta}}{\ell^3}
+        \right)
+        \kappa\ dr^\prime d\phi^\prime d\lambda^\prime.
     \label{eq:tesstensor}
 \end{equation}
+
+where $\alpha, \beta \in \{x, y, z\}$,
+$\rho$ is the density,
+$G = 0.00000000006673\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
+$\kappa = {r^\prime}^2 \cos \phi^\prime$,
+and
+
+\begin{eqnarray}
+    \Delta_x &=& r^\prime\cos\phi\sin\phi^\prime - \sin\phi\cos\phi^\prime
+                 \cos(\lambda^\prime - \lambda), \\
+    \Delta_y &=& r^\prime \cos \phi^\prime \sin(\lambda^\prime - \lambda), \\
+    \Delta_z &=& r^\prime \cos \psi - r,\\
+    \ell &=& \sqrt{{r^\prime}^2 + r^2 - 2 r^\prime r \cos \psi}, \\
+    \cos\psi &=& \sin\phi\sin\phi^\prime + \cos\phi\cos\phi^\prime
+                 \cos(\lambda^\prime - \lambda).
+\end{eqnarray}
 
@@
 
-We refer the reader to \citet{Barrera-Figueroa2006} for an algorithm tailored
-for Legendre polynomials in particular.
-
 The second order roots, as well as the ones determined by a root finder
@@
 
-Thus, the roots must be scaled to the integration limits $[a, b]$ by
+Before being used for GLQ integration,
+the roots must be scaled to the integration limits $[a, b]$ using
+
 \begin{equation}
@@
 The weights of the GLQ are given by \citep{Hildebrand1987},
+
 \begin{equation}
@@
 \ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
-becomes \citep{Asgharzadeh2007} (EQUATION).
+becomes \citep{Asgharzadeh2007}
 
-Applying equation (GLQ 3D) to equations
+\begin{equation}
+    \iiint\limits_{\Omega}
+    f(r^\prime, \lambda^\prime, \phi^\prime)
+    dr^\prime d\lambda^\prime d\phi^\prime
+    \approx
+    A
+    \sum\limits_{i=1}^{N^r}
+    \sum\limits_{j=1}^{N^\phi}
+    \sum\limits_{k=1}^{N^\lambda}
+    f(r_i, \phi_j, \lambda_k),
+    \label{eq:glq3d}
+\end{equation}
+
+where
+
+\begin{equation}
+    A = \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8}.
+\end{equation}
+
+Applying equation~\ref{eq:glq3d} to equations
 \ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
@@
 
-\citet{Li2011} proposed an alternative way to decrease the distance between
-point masses and achieve an accurate integration.
+Instead of increasing the GLQ order,
+\citet{Li2011} proposed an alternative method
+for decreasing the distance between point masses
+and achieve an accurate integration.
 

### 2014-11-26 manuscript.tex
Joined text for the theory part with equations

@@
 
+\noindent
 where $\alpha, \beta \in \{x, y, z\}$,
@@
 using the Gauss-Legendre Quadrature (GLQ).
-
 The GLQ consists of approximating the integral by a weighted sum of the
@@
 
-$N$ is the order of the quadrature, i.e. the number of points used in the GLQ.
-
-The points $x_i$ where the kernel is evaluated are called the quadrature nodes.
-
-The points are roots of the $N^{th}$ order Legendre polynomial $P_N(x)$.
-
-The roots are $x = \pm 0.577350269$ for a second order polynomial ($P_2(x)$).
-
-A root finder algorithm can be used to determine the roots for larger order
-polynomials.
-
-The second order roots, as well as the ones determined by a root finder
-algorithm, will be within the range $[-1, 1]$.
-
+\noindent
+in which $N$ is the order of the quadrature,
+i.e. the number of points used in the GLQ.
+The points $x_i$ are called the quadrature nodes.
+They are the roots of the $N^{th}$ order Legendre polynomial $P_N(x)$.
+For a second order polynomial ($P_2(x)$),
+the roots are $x = \pm 0.577350269$.
+Roots for larger order polynomials
+can be determined by a root finder algorithm.
+Roots of Legendre polynomials
+will be within the range $[-1, 1]$.
 Before being used for GLQ integration,
@@
 \begin{equation}
-    W_i = \frac{2}{(1 - x_i^2)(P^\prime_N(x_i))^2},
+    W_i = \frac{2}{(1 - x_i^2)(P^\prime_N(x_i))^2}.
     \label{eq:glq_weights}
@@
 
-The values of the $P_N(x)$ and its first derivative $P^\prime_N(x)$ can be
-calculated with recursive relations.
+\noindent
+The values of $P_N(x)$ and its first derivative $P^\prime_N(x)$
+can be calculated with recursive relations.
 
-The Gauss-Legendre Quadrature for volume integrals, like equations
+The Gauss-Legendre Quadrature for three-dimensional volume integrals,
+like equations
 \ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
@@
 
+\noindent
 where
@@
 mass located on the quadrature nodes.
-
-It can be said that the GLQ integration approximates the volume integrals  by a
+Thus, it can be said that the GLQ integration
+approximates the volume integrals  by a
 weighted sum of point mass effects.
@@
 depends on the number of point masses used in the summation.
-
 \citet{Ku1977} showed that it also depends on the ratio between
 the distance to the computation point and the distance between adjacent nodes.
-
 Figure~\ref{fig:glqerrorsample}
 illustrates this effect on the $g_{xy}$ gravity gradient component.
-
 The $g_{xy}$ component was produced by a
@@
 and top at $z=0\ km$.
-
 The maps were calculated on a regular grid
 with $100\times100$ points.
-
-The GLQ order used was the same
-in the radial, longitudinal, and latitudinal dimensions.
-
 Figure~\ref{fig:glqerrorsample}a shows the $g_{xy}$ component
@@
 directions, respectively).
-
 Figure~\ref{fig:glqerrorsample}b shows $g_{xy}$ computed with order two
 GLQ as well but at 150 km height.
-
 Notice that the computed effect is concentrated around each point mass
 of the GLQ (black dots) and does not resemble the effect of a tesseroid.
-
 \citet{Ku1977} determined an empirical criterion that the distance between
@@
 the computation point.
-
 Thus, if a computation point is too close to the tesseroid one would have to
@@
 result.
-
 One way to accomplish this would be increase the order of the quadrature
 $N$ in all three directions.
-
 Figure~\ref{fig:glqerrorsample}c shows the $g_{xy}$ component calculated at
 150km height but with a GLQ order of 30.
-
 The computed $g_{xy}$ component more closely resembles the results from

### 2014-11-28 manuscript.tex
Reordered and joined sentences for methods

@@
     Leonardo Uieda\footnotemark[1]\footnotemark[2],
-    Val\'eria C. F. Barbosa\footnotemark[2],
-    Carla Braitenberg\footnotemark[3],
-    and
-    Vanderlei C. Oliveira Jr\footnotemark[2]
+    et al.
 }
@@
     email: leouieda@gmail.com
-    \\
-    \footnotemark[2]Observat\'orio Nacional,
-    Rio de Janeiro, Brazil.
-    \\
-    \footnotemark[3]Department of Mathematics and Geosciences,
-    University of Trieste, Trieste, Italy.
 }
@@
 
-A tesseroid is a spherical mass element defined in geocentric spherical
+A tesseroid is a mass element defined in geocentric spherical
 coordinates
@@
 $\rho$ is the density,
-$G = 0.00000000006673\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
-$\kappa = {r^\prime}^2 \cos \phi^\prime$,
+$G = 6.673\times10^{-11}\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
 and
@@
     \Delta_z &=& r^\prime \cos \psi - r,\\
+    \kappa &=& {r^\prime}^2 \cos \phi^\prime, \\
     \ell &=& \sqrt{{r^\prime}^2 + r^2 - 2 r^\prime r \cos \psi}, \\
@@
 The Gauss-Legendre Quadrature for three-dimensional volume integrals,
-like equations
-\ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
+like equations~\ref{eq:tesspot}-\ref{eq:tesstensor},
 becomes \citep{Asgharzadeh2007}
@@
 
-Applying equation~\ref{eq:glq3d} to equations
-\ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
+Comparing equation~\ref{eq:glq3d} with
+equations~\ref{eq:tesspot}-\ref{eq:tesstensor},
 we see that $f(r_i, \phi_j, \lambda_k)$ is the effect of a point
@@
 calculated at 400 km height using
-GLQ with order two (two point masses in the longitude, latitude, and vertical
-directions, respectively).
+GLQ with order two ($2 \times 2 \times 2 = 8$ point masses).
 Figure~\ref{fig:glqerrorsample}b shows $g_{xy}$ computed with order two
@@
 Figure~\ref{fig:glqerrorsample}c shows the $g_{xy}$ component calculated at
-150km height but with a GLQ order of 30.
-The computed $g_{xy}$ component more closely resembles the results from
-\citet{Asgharzadeh2007}.
+150km height but with a GLQ order of 30
+($30 \times 30 \times 30 = 27,000$ point masses).
+The computed $g_{xy}$ component more closely resembles
+the expect results for a single tesseroid \citep{Asgharzadeh2007}.
 
@@
         the computation height
-        and the number of point masses in the Gauss-Legendre Quadrature
-        (i.e., the GLQ order).
-        Black circles represent the horizontal location of the GLQ nodes
-        (not shown for GLQ order 30).
-        a) Field calculated at $400\ km$ height using GLQ order 2.
-        The effect is similar to that presented by \citet{Asgharzadeh2007}.
-        b) At $150\ km$ height and order 2,
-        the field resembles that of
-        four point sources located at the GLQ nodes.
-        This effect was shown by \citet{Ku1977}
-        for the case of the $g_z$ component of a right rectangular prism.
+        and the number of point masses in the Gauss-Legendre Quadrature.
+        Black circles represent the horizontal location of the point masses.
+        a) $g_{xy}$ calculated at $400\ km$ height using GLQ order 2
+        ($2 \times 2 \times 2 = 8$ point masses).
+        b) At $150\ km$ height and GLQ order 2,
+        the result resembles that of
+        four point mases instead of a single tesseroid.
+        This effect was shown by \citet{Ku1977}.
         c) At $150\ km$ but with a higher GLQ order of 30,
-        the field is similar to that expected for a single mass source.
+        the result is similar to that expected for a single mass source.
     }
@@
 
-Instead of increasing the GLQ order,
 \citet{Li2011} proposed an alternative method
@@
 and achieve an accurate integration.
+Instead of increasing the GLQ order,
+they keep it fixed to a given number
+and divide the tesseroid into smaller parts.
+The sum of the effects of the smaller tesseroids
+is equal to the gravitational effect of the larger tesseroid.
+This division effectively decreases
+the distance between nodes because
+of the smaller size of the tesseroids.
+The criterion for dividing a tesseroid or not
+is that the distance to the computation point
+must be smaller than a constant times the size of the tesseroid.
+This is analogous to the criterion of \citet{Ku1977}
+because the size of the tesseroid serves as a proxy
+for the distance between point masses.
+This procedure is repeated recursively
+(i.e., to each of the smaller tesseroids and so on)
+until all tesseroids are within the acceptable
+ratio of distance and size or a minimum size is achieved.
 
-Instead of increasing the GLQ order,
-they divide the tesseroid into smaller tesseroids while keeping the number of
-point masses per tesseroid fixed.
-
-The gravitational effect of the larger tesseroid would be calculated as the sum
-of the effects of the eight smaller tesseroids.
-
-This division would effectively decrease the distance between nodes because of
-the smaller tesseroids.
-
-The criterion for dividing a tesseroid is that
-the distance to the computation point must be smaller than
-a constant times the size of the tesseroid.
-
-The size of the tesseroid is used as a proxy for the distance between point
-masses.
-
-This procedure is repeated recursively
-(i.e., to each of the smaller tesseroids)
-until all tesseroids are within the acceptable distance-size ratio
-or a minimum size is achieved.
-
-The advantage of this adaptive discretization is that the number of point
-masses is only increased in regions of the tesseroid that are closer to the
-computation point.
-
-Alternatively, simply increasing the order of the GLQ would increase the number
-of point masses evenly throughout the whole tesseroid.
+The advantage of this adaptive discretization is
+that the number of point masses is only increased
+in regions of the tesseroid that are
+closer to the computation point.
+Alternatively,
+simply increasing the order of the GLQ
+would increase the number of point masses
+evenly throughout the whole tesseroid.
 
@@
 
-The calculation of the tesseroid gravitational fields with adaptive
-discretization is implemented in the open-source package \textit{Tesseroids}.
+We have implemented the calculation of
+the tesseroid gravitational fields
+with adaptive discretization
+in the open-source package \textit{Tesseroids}.
+The software is freely available
+on the Internet\footnote{http://www.leouieda.com/tesseroids}
+under the BSD 3-clause license.
+An archived version of the source code
+is available as part of this article.
 
-Tesseroids is freely available under the BSD 3-clause free-software license.
-
-An archived version of the source code is available along with this article.
-
-The official website\footnote{http://www.leouieda.com/tesseroids}
-hosts the software documentation and
-archives of the latest binary executables and source code.
-
-Ongoing development takes place through the source code hosting site
-Github\footnote{https://github.com/leouieda/tesseroids}.
-
-\textit{Tesseroids} is written in the C programming language.
-
-Tesseroids consists of command-line programs.
-
-Tesseroids tries to follow the Unix principals of
+\textit{Tesseroids}  consists of command-line programs
+written in the C programming language.
+The command-line interface follows the Unix principals of
 (1) making each program do one thing well,
@@
 programs.
-
-The gravitational potential, attraction, and Marussi tensor components
-are each computed by a separate program.
-
-Programs for model and computation grid generation are also included.
-
-The package also includes programs for computing the gravitational fields of
-right rectangular prisms in Cartesian and spherical coordinates.
-
-Tesseroids programs receive input through command-line arguments and the
-standard input channel (STDIN) and output the results through
-the standard output channel (STDOUT).
-
-The command-line interface of each program was designed so that they
-can be called in sequence using a Unix pipe.
-
-For example, the command to
-generate a regular grid with $NLON \times NLAT$ points,
-calculate on it the $g_z$ and $g_{zz}$ caused by the tesseroids in file
-MODELFILE,
-and save the results to a file called OUTPUT
+The package includes programs to calculate
+the gravitational fields of tesseroids and
+rectangular prisms (in both Cartesian and spherical coordinates),
+as well as model and computation grid generation.
+All programs receive input through
+command-line arguments and the standard input channel (STDIN)
+and output the results through the standard output channel (STDOUT).
+For example,
+the command to generate a regular grid with $NLON \times NLAT$ points,
+calculate on it $g_z$ and $g_{zz}$ caused by
+the tesseroids in a file ``MODELFILE'',
+and save the results to a file called ``OUTPUT''
 is:
@@
 
-The \texttt{src} folder of the distribution archive
+The \texttt{src} folder of the source code archive
 contains the C files that build the command-line programs
 (e.g., \texttt{tessgz.c}).
-
-The \texttt{src/lib} folder contains the source files that implement the
-numerical computations.
-
+The \texttt{src/lib} folder contains
+the source files that implement the numerical computations.
 We will not describe here the implementation of the input/output parsing and
 other miscellanea.
-
 Instead, we will focus on the details of the Gauss-Legendre Quadrature
-integration of equations~\ref{eq:tesspot},~\ref{eq:tessgrav},
-and~\ref{eq:tesstensor}
+integration of equations~\ref{eq:tesspot}-\ref{eq:tesstensor}
 and the adaptive discretization of tesseroids.
 
-All figures and numerical experiments in this article were produced by the
-IPython notebook\footnote{http://ipython.org}
-files \citep{Perez2007} in the supplementary material.
-
-These files combine the analysis source code
-with explanatory text, equations, and the figures generated by the code.
-
-This supplementary material can be run and executed locally or viewed
-online\footnote{https://github.com/pinga-lab/paper-tesseroids}
-(in a non-executable form).
+All figures and numerical experiments in this article
+were produced by IPython notebooks\footnote{http://ipython.org}
+\citep{Perez2007}.
+These files combine the analysis and figure generation source code
+with text, equations, and the figures generated by the code.
+These files,
+as well as instructions for installing and running the analysis,
+are distributed along with the source code archive
+that accompanies this article.
+Alternatively,
+the notebooks and can also be downloaded
+or viewed in a non-executable form from
+an online repository\footnote{https://github.com/pinga-lab/paper-tesseroids}.
 
@@
 
-The source file \texttt{src/lib/glq.c} contains the code necessary to perform a
-Gauss-Legendre Quadrature integration.
-
+The source file \texttt{src/lib/glq.c}
+contains the code necessary to perform
+a Gauss-Legendre Quadrature integration.
 The first step in the GLQ is to compute the
-locations of the point masses (i.e., the discretization points).
-
+locations of the discretization points (i.e., the point masses).
 These points are roots of Legendre polynomials.
-
-Precomputed values for these roots are available low order polynomials,
-usually up to order 5.
-
-For flexibility, we use the multiple root-finder algorithm of
-\citet{Barrera-Figueroa2006} to compute the roots for polynomials of any order.
-
-The additional computational load is minimal since the root-finder algorithm
+Precomputed values are available low order polynomials,
+typically up to order five.
+For flexibility and to compute higher order roots,
+we use the multiple root-finder algorithm of
+\citet{Barrera-Figueroa2006}.
+The additional computational load is minimal
+because the root-finder algorithm
 must be run only once per program execution.
-
-The root-finder algorithm is implemented in functions \texttt{glq\_nodes} and
-\texttt{glq\_next\_root}.
-
+The root-finder is implemented in functions
+\texttt{glq\_nodes} and \texttt{glq\_next\_root}.
 The computed roots will be in the range $[-1, 1]$
-and must be scaled to the integration limits (the limits of the tesseroid)
+and must be scaled to the integration limits
+(the physical boundaries of the tesseroid)
 using function \texttt{glq\_set\_limits} (see equation~\ref{eq:glq_scaling}).
 
-The GLQ weights are computed using equation~\ref{eq:glq_weights} by function
-\texttt{glq\_weights}.
-
-Both the computed roots and weights are stored in a data structure (a C
-\texttt{struct}) called \texttt{GLQ}.
-
-Function \texttt{glq\_new} allocates the memory necessary,
+The GLQ weights (equation~\ref{eq:glq_weights})
+are computed by function \texttt{glq\_weights}.
+Both the computed roots and weights are stored in a data structure
+(a C \texttt{struct}) called \texttt{GLQ}.
+Function \texttt{glq\_new}
+handles memory allocation,
 calculates the roots and weights,
@@
 
-The numerical integration is performed by the functions in module
-\texttt{src/lib/grav\_tess.c}.
+The numerical integration of the tesseroid gravitational fields
+is performed by the functions in module \texttt{src/lib/grav\_tess.c}.
+Functions \texttt{tess\_pot}, \texttt{tess\_gx}, \texttt{tess\_gy}, and so on,
+compute the gravitational fields of a single tesseroid
+on a single computation point.
+These functions require three \texttt{GLQ} structures,
+each containing the roots and weights
+for GLQ integration in the three dimensions.
+The roots must be scaled to the
+integration limits
+$[\lambda_1, \lambda_2], [\phi_1, \phi_2], [r_1, r_2]$
+(see equations~\ref{eq:tesspot}-\ref{eq:tesstensor}).
+The integration consists of three loops
+that sum the weighted kernel functions
+evaluated at each GLQ point mass (the scaled roots).
 
-The functions \texttt{tess\_pot}, \texttt{tess\_gx}, \texttt{tess\_gy},
-and so on, compute the gravitational fields of a single tesseroid on a single
-computation point using the GLQ.
-
-These functions require three \texttt{GLQ} structures that contain the roots
-and weights for GLQ integration in the three dimensions.
-
-The roots must be scaled to the
-integration limits $[\lambda_1, \lambda_2], [\phi_1, \phi_2], [r_1, r_2]$
-(see equations~\ref{eq:tesspot}-\ref{eq:tesstensor}).
-
-The integration consists of three \texttt{for} loops that sum the weighted
-kernel functions evaluated at each GLQ point mass (the scaled roots).
-
-The evaluation of the trigonometric functions in
-equations~\ref{eq:tesspot}-\ref{eq:tesstensor} inside the inner loops
-are the biggest bottlenecks for the numerical integration.
-
-Better performance is achieved by
-precomputing the sine and cossine of latitudes
-and moving some trigonometric function
-evaluations to the outer loops.
+The biggest bottlenecks for the numerical integration are
+the number of point masses used
+and he evaluation of the trigonometric functions in
+equations~\ref{eq:tesspot}-\ref{eq:tesstensor} inside the inner loops.
+Better performance is achieved
+by pre-computing the sine and cosine of latitudes
+and moving some trigonometric function evaluations
+to the outer loops.
 
@@
 
-The adaptive discretization algorithm of \citet{Li2011}
-divides a tesseroid recursively.
-
-A tesseroid will be divided when the smallest distance between it and the
-computation point is smaller than a constant times the largest dimension of the
-tesseroid.
-
+Our implementation of the adaptive discretization algorithm
+differs in a few ways from the original.
+In \citet{Li2011},
+a tesseroid will be divided when
+the smallest distance between it and the computation point
+is smaller than a constant times
+the largest dimension of the tesseroid.
 Instead of the smallest distance,
-we use in our implementation
-the distance $d$ between
+we use the easier to calculate
+distance between
 the computation point $(r, \lambda, \phi)$
@@
 
-We define the dimensions of the tesseroid
-along longitude, latitude, and radius, respectively, as
-(see Figure~\ref{fig:division}a)
+Our definition of the dimensions of the tesseroid
+(the ``side lengths'' of \citet{Li2011})
+along longitude, latitude, and radius, respectively, are
+(Figure~\ref{fig:division}a)
 
@@
 
-We fix the GLQ order $N=2$ for all three dimensions, resulting in eight point
-masses in total.
-
-The criterion we use to determine if a tesseroid must be divided is
+To determine if a tesseroid must be divided,
+we check if
 
@@
 \end{equation}
+
 \noindent
-where $i \in (\lambda, \phi, r)$
-and $D$ is a positive scalar called the ``distance-size ratio''.
+for each $i \in (\lambda, \phi, r)$.
+$D$ is a positive scalar
+hereafter referred to as the ``distance-size ratio''.
+If the inequality holds for all three dimensions,
+the tesseroid is not divided.
+Thus, the distance-size ratio determines
+how close the computation point can be
+before we must divide the tesseroid.
+The value of $D$ is indirectly responsible for
+the accuracy of the solution and the computation time.
+We will explore the relationship with the accuracy in the following section.
 
-$D$ determines how close the
-computation point can be before we must divide the tesseroid.
-
-Thus, the value of $D$ is indirectly responsible for
-the accuracy of the solution and the computation time.
-
-The modeling programs provide default values of $D$.
-
-It can also be controlled through command-line arguments passed to the
-programs.
-
-We used a stack-based implementation of the algorithm instead of using
-recursive function calls, as originally proposed by \citet{Li2011}.
-
-Stacks are array-like data structures with a particular way of inserting and
-removing elements from it.
-
-In a stack, one can only insert elements on the top (the last empty position).
-
-Likewise,
-one can only remove the last element that was placed on the stack
-(commonly referred to as ``popping'' the stack).
-
-Stacks are also known as ``Last-in-first-out'', or LIFO, structures.
-
-LIFOs are often described as being analogous to a stack of plates.
-
-One can only place a new plate on the top of the stack and
-can only remove a plate safely from the top of the stack.
-
-We will describe the stack-based algorithm for a single tesseroid and
-computation point.
-
-The algorithm is implemented in function \texttt{calc\_tess\_model\_adapt}
-of the file \texttt{src/lib/grav\_tess.c}.
-
-Start by creating an empty stack of tesseroids.
-
-The  stack is represented by the \texttt{stack} variable,
-an array of \texttt{TESSEROID} structures.
-
-We must define a maximum size for the stack to allocate memory for it.
-
-Defining a maximum size allows us to avoid an infinite loop and large
-consumption of computer memory in case the computation point is
-on (or sufficiently close to) the surface of the tesseroid.
-
-We use the integer \texttt{stktop} to keep track of the index of the last
-element in the stack (the top of the stack).
-
-The stack is initialized with the input tesseroid.
-
-The initialization is done by copying the tesseroid into the stack and
-setting \texttt{stktop} to zero (the first element).
-
-Once the stack is initialized, the steps of the algorithm are:
-
-\begin{enumerate}
-    \item ``Pop'' the stack (i.e., take the last element from it).
-        This will cause \texttt{stktop} to be reduced by one.
-    \item Compute the distance $d$ between the tesseroid and
-        the computation point (equation~\ref{eq:distance}).
-    \item Compute the dimensions of the tesseroid $L_\lambda$, $L_\phi$,
-        and $L_r$ (equations~\ref{eq:sizelon}-\ref{eq:sizer}).
-    \item Check the condition in equation~\ref{eq:condition} for each
-        dimension of the tesseroid.
-    \item If none of the dimensions fail the above condition,
-        compute the gravitational effect of the tesseroid using the
-        Gauss-Legendre Quadrature with order two
-        (equation~\ref{eq:glq3d}).
-    \item If any of the dimensions fail the condition:
-    \begin{enumerate}
-        \item Divide the tesseroid in half along each dimension that failed
-             the condition.
-        \item Check if the number of new elements plus \texttt{stktop}
-             is smaller than the maximum stack size.
-             If it is not (i.e., the stack is full),
-             warn the user of a ``stack overflow''
-             and compute the effect of the larger tesseroid (step 5).
-        \item If the stack is not full, place the smaller tesseroids into
-             the stack.
-    \end{enumerate}
-    \item Repeat the above steps until the stack is empty
-        (\texttt{stktop} is equal to -1).
-\end{enumerate}
-
-This stack-based implementation has some advantages over the original recursive
-implementation suggested by \citet{Li2011}.
-
-(1) It will be faster because it bypasses the overhead of function calls.
-
-The speedup will be specially significant if the implementation is done
-in high level interpreted languages, like Matlab and Python.
-
-(2) It gives the developer more control over the recursion step.
-
-In recursive implementations,
-the developer has no control over the maximum number of consecutive
-recursive calls (the ``recursion depth'').
-
-This limit may vary with programming language, compiler, and operating system.
-
-Overflowing the maximum recursion depth usually results in program
-crashes with cryptic or inexistent error messages.
-
-In the stack-based implementation, the developer has complete control.
-
-Overflowing of the stack can be handled gracefully with an error message or
-even performing a suitable approximation of the result.
-
-Figure~\ref{fig:division} shows the effect of the distance-size ration $D$
-in the adaptive discretization of a tesseroid for the computation at a point P.
-
-Computing the gravitational effect of the tesseroid in
-Figure~\ref{fig:division}a using the algorithm described above
-is equivalent to computing the effect of the tesseroids in
-Figures~\ref{fig:division}b-d.
-
-Note that increasing the value of $D$ results in a finer mesh of tesseroids.
-
-For $D = 1$ (Figure~\ref{fig:division}b) the number of tesseroids in the
-adaptive discretization is 4,
-for $D = 2$ (Figure~\ref{fig:division}c) it is 38,
-and for $D = 6$ (Figure~\ref{fig:division}d) it is 936.
+Figure~\ref{fig:division} shows examples of
+the resulting tesseroid models after adaptive discretization.
+Figure~\ref{fig:division}a shows
+the initial tesseroid and computation point P.
+Figures~\ref{fig:division}b-d are
+the result of adaptive discretization using
+different values of the distance-size ratio $D$,
+respectively,
+$D=1$, $D=2$, and $D=6$.
+The number of tesseroids in the resulting discretization is, respectively,
+4, 38, and 936.
+Code to perform the adaptive discretization
+and produce Figure~\ref{fig:division} is available in
+the IPython notebook \texttt{methods\_figures.ipynb}.
 
@@
 \end{figure}
+
+Instead of using recursive function calls,
+as originally proposed by \citet{Li2011},
+we use a stack-based implementation of the algorithm.
+Stacks are array-like data structures
+with a particular way of inserting and removing elements from it.
+In a stack,
+one can only insert elements to the top of the stack
+(the last empty position).
+Likewise,
+one can only remove the last element of the stack
+(commonly referred to as ``popping'' the stack).
+Stacks are also known as ``Last-in-first-out'', or LIFO, structures.
+
+We will describe the stack-based algorithm
+for computing the effect of a single tesseroid
+on a single computation point.
+The algorithm is implemented in function \texttt{calc\_tess\_model\_adapt}
+of the file \texttt{src/lib/grav\_tess.c}.
+The stack of tesseroids is represented by
+the \texttt{stack} variable,
+an array of \texttt{TESSEROID} structures.
+We must define a maximum size for the stack to allocate memory for it.
+Defining a maximum size allows us to
+avoid an infinite loop
+in case the computation point is on
+(or sufficiently close to) the surface of the tesseroid.
+We use the integer \texttt{stktop}
+to keep track of the index of
+the last element in the stack (the top of the stack).
+
+The algorithm starts by creating an empty stack of tesseroids.
+Then, the stack is initialized with the input tesseroid.
+The initialization is done by copying the tesseroid into the stack
+and setting \texttt{stktop} to zero (the first element).
+Once the stack is initialized, the steps of the algorithm are:
+
+\begin{enumerate}
+    \item ``Pop'' the stack (i.e., take the last tesseroid from it).
+        This will cause \texttt{stktop} to be reduced by one.
+        This tesseroid is the one that will be evaluated in the following
+        steps.
+    \item Compute the distance $d$ between
+        the geometric center of the tesseroid and
+        the computation point using equation~\ref{eq:distance}.
+    \item Compute the dimensions of the tesseroid $L_\lambda$, $L_\phi$,
+        and $L_r$ using equations~\ref{eq:sizelon}-\ref{eq:sizer}.
+    \item Check the condition in equation~\ref{eq:condition} for each
+        dimension of the tesseroid.
+    \item If none of the dimensions fail the above condition,
+        compute the gravitational effect of the tesseroid using the
+        Gauss-Legendre Quadrature
+        (equations~\ref{eq:tesspot}-\ref{eq:tesstensor} and~\ref{eq:glq3d}).
+        We use a GLQ order of two for all three dimensions
+        ($2 \times 2 \times 2 = 8$ point masses)
+        by default.
+        This value can be changed using
+        a command-line argument of the modeling programs.
+    \item If any of the dimensions fail the condition:
+    \begin{enumerate}
+        \item Divide the tesseroid in half along each dimension that failed
+             the condition.
+        \item Check if the number of new elements plus \texttt{stktop}
+             is smaller than the maximum stack size.
+             If it is not (i.e., the stack is full),
+             warn the user of a ``stack overflow''
+             and compute the effect of the larger tesseroid (step 5).
+        \item If the stack is not full, place the smaller tesseroids into
+             the stack.
+    \end{enumerate}
+    \item Repeat the above steps until the stack is empty
+        (\texttt{stktop} is equal to -1).
+\end{enumerate}
+
+This stack-based implementation
+has some advantages over the original recursive implementation,
+namely:
+(1) It will be faster because it bypasses the overhead of function calls.
+(2) It gives the developer more control over the recursion step.
+In recursive implementations,
+the developer has no control over
+the maximum number of consecutive recursive calls
+(i.e., the ``recursion depth'').
+This limit may vary with programming language,
+compiler, and operating system.
+Overflowing the maximum recursion depth
+may result in program crashes,
+typically with cryptic or inexistent error messages.
+In the stack-based implementation,
+the developer has complete control.
+Overflowing of the stack can be handled gracefully
+with an error message
+or even performing a suitable approximation of the result.
 

### 2014-12-03 manuscript.tex
Added a little bit of text to accuracy section

@@
 
-The key controlling point of their algorithm is the ratio between
-the smallest distance to the computation point and
-the largest side length of the tesseroid.
+The key controlling point of the adaptive discretization algorithm
+is the distance-size ratio $D$ (equation~\ref{eq:condition}).
 
-The value of this ratio is used to decide whether a tesseroid will be divided
-or not.
+The value chosen for this ratio determines how many divisions are made.
 
-Thus, the value chosen for the minimum allowed value of this ratio determines
-how many divisions are made.
-
-Consequently, the value for the minimum distance-size ratio controls both the
-accuracy of the integration and the computation time.
+$D$ controls both accuracy of the integration and the computation time.
 
@@
 
-However, there has been no further investigation into
-the relationship between values of the minimum distance-size
-ratio and the computation error.
+There has been no further investigation into
+the relationship between values of the distance-size ratio and
+the computation error.
 
-It is also not clear whether the same relationship holds for the gravitational
-attraction as well as for the gradient tensor components.
+It is also not clear whether the same value of $D$ used for the gravitational
+attraction can be used for the gradient tensor components.
 
-\subsection{$g_{zz}$ variability with computation height}
+We need a reference value against which to compare the computed tesseroid effects.
+
+The spherical shell is a convenient model with an analytical solution along
+the polar axis.
+
+\citep{Grombein2013} provides the equations for
+the gravitational potential, attraction, and gravity gradient components
+of a homogeneous spherical shell.
 

### 2014-12-08 manuscript.tex
Draft of accuracy section

@@
 
-$D$ controls both accuracy of the integration and the computation time.
+$D$ controls both the accuracy of the integration and the computation time.
 
 \citet{Li2011} provide a few examples of values for this ratio in the case of
-the vertical component of gravitational attraction.
+the vertical component of the gravitational attraction.
 
@@
 
+The computation point can be considered at any point with equal radial
+coordinate located outside the shell because of the symmetry of the shell.
+
+We used a spherical shell with a thickness of 1 km.
+
+The density of the shell is $2670\ kg.m^{-3}$.
+
+The bottom of the shell is at height 0 km above the reference sphere.
+
+The top of the shell is at 1 km height.
+
+The largest errors in the GLQ integration occur directly on top of the
+tesseroid but not exactly at it's center, as can be seen in
+Figure~\ref{fig:glqerrorsample}.
+
+We discretized the spherical shell into tesseroids
+with dimensions of $1^\circ \times 1^\circ \times 1\ km$.
+
+We calculate the effect of a tesseroid model of the spherical shell on a
+regular grid.
+
+The grid is at a constant height of 2 km (1 km above the shell).
+
+The grid is located on top of a tesseroid located at the North pole.
+
+We repeated the tesseroid calculations with values of the distance-size ratio
+varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
+
+Figure~\ref{fig:tess_vs_shell} shows the maximum difference between the
+shell and tesseroid fields as a function of the distance-size ratio.
+
+Differences are given as a percentage of the shell value.
+
+The curves allow us to choose a value of the distance-size ratio that
+produces an acceptable error.
+
+We chose 0.1\% as the maximum allowed error.
+
+We performed the computations again for three different scenarios: at 260 km
+height (the mean orbit height of the GOCE satellite), on top of a tesseroid at
+the equator, and for a $30^\circ \times 30^\circ \times 1\ km$ tesseroid model
+of the shell.
+
+We performed these computations to investigate if the GLQ error decays the same
+way under different scenarios.
+
+The error curves for the gravitational potential show increased errors at the
+equator and for a $30^\circ$ tesseroid.
+
+All errors will be smaller than 0.1\% if one uses a distance-size ratio of 1.
+
+Even in the extreme case of a $30^\circ$ tesseroid with computation points at a
+height of 1 km above the tesseroid.
+
+A distance-size ratio of 1.5 is sufficient to achieve errors below 0.1\% in
+$g_z$.
+
+A distance-size ratio of 8 is required to achieve errors below 0.1\% in
+the gravity gradient components.
+
+The error curve for the computation grid at 260 km height shows that
+the error is smaller at higher altitudes.
+
+Computation at higher altitudes would not change the chosen values of the
+distance-size ratio for the potential and $g_z$.
+
+The computation height changes the value of the distance-size ratio required
+for the gravity gradient components.
+
+We computed the error curves for $g_{zz}$ only at different heights
+(Figure~\ref{fig:gzz_error_height}).
+
+The distance-size ratio can be reduced to 4 if the computation is done at 50 km
+height.
+
+One can use a distance-size ratio of 2.5 for computations at 260 km height.
+
+These reductions in the distance-size ratio can reduce the computation time
+required to achieve the desired.
+
+The software \textit{Tesseroids} includes default values of the distance-size
+ratio.
+
+We used the values required to achieve 0.1\% accuracy: 1 for the potential,
+1.5 for the gravitational attraction, and 8 for the gravity gradient
+components.
+
+The value of $D=8$ for the gravity gradients is conservative.
+
+We chose it to ensure the accuracy of the computation in most cases.
+
+The experienced user of the software can set smaller values using command-line
+options.
+
+The smaller values result in less computation time.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-12-08 manuscript.tex
Cite who used Tesseroids before
Revising accuracy section
Added some more sentences to explain the problem setup and reordered.

@@
 This makes the adaptive discretization more computationally efficient.
+
+This article describes the software \textit{Tesseroids} and the implementation
+of the adaptive discretization algorithm.
+
+Previous versions of \textit{Tesseroids} have been used by, e.g.,
+\citet{Alvarez2012, Bouman2013, Bouman2013a, Mariani2013, Braitenberg2014,
+Braitenberg2011, Fullea2014}.
 
@@
 
-The value chosen for this ratio determines how many divisions are made.
+$D$ controls both the accuracy of the integration and the computation time.
 
-$D$ controls both the accuracy of the integration and the computation time.
+The value chosen for $D$ determines how many divisions are made.
 
@@
 
-There has been no further investigation into
-the relationship between values of the distance-size ratio and
+To our knowledge,
+there has been no further investigation
+into the relationship between
+values of the distance-size ratio and the computation error.
+
+It is also not clear whether
+the same value of $D$ yields compatible error levels
+for the gravitational attraction
+and the gradient tensor components.
+
+We must determine the relationship between the distance-size ratio $D$ and
 the computation error.
 
-It is also not clear whether the same value of $D$ used for the gravitational
-attraction can be used for the gradient tensor components.
+We need a reference against which to compare the computed tesseroid fields.
 
-We need a reference value against which to compare the computed tesseroid effects.
+The spherical shell is a convenient reference model.
 
-The spherical shell is a convenient model with an analytical solution along
-the polar axis.
+The spherical shell has a analytical solutions for the gravitational potential,
+attraction, and gradient tensor along the polar axis.
 
@@
 of a homogeneous spherical shell.
-
-The computation point can be considered at any point with equal radial
-coordinate located outside the shell because of the symmetry of the shell.
 
@@
 
-The largest errors in the GLQ integration occur directly on top of the
-tesseroid but not exactly at it's center, as can be seen in
-Figure~\ref{fig:glqerrorsample}.
+We can use the spherical shell values to calculate the error of the tesseroid
+computations for varying values of $D$.
 
@@
 
-We calculate the effect of a tesseroid model of the spherical shell on a
-regular grid.
+We calculate the effect of a tesseroid model on a regular grid.
 
@@
 
+The computation point can be considered at any point with equal radial
+coordinate located outside the shell because of the symmetry of the shell.
+
+We calculate the difference between the shell values and the values computed
+from tesseroids.
+
+We keep only the largest difference.
+
+Calculating on a grid above a tesseroid is important because
+the largest errors in the GLQ integration occur directly on top of the
+tesseroid but not exactly at it's center, as can be seen in
+Figure~\ref{fig:glqerrorsample}.
+
+The difference between shell and tesseroid values would be underestimated
+if the computation was performed only at a single point on the polar axis.
+
 We repeated the tesseroid calculations with values of the distance-size ratio
-varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
+$D$ varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
 
 Figure~\ref{fig:tess_vs_shell} shows the maximum difference between the
-shell and tesseroid fields as a function of the distance-size ratio.
+shell and tesseroid fields as a function of $D$.
 
-Differences are given as a percentage of the shell value.
+The differences are given as a percentage of the shell value.
 
-The curves allow us to choose a value of the distance-size ratio that
-produces an acceptable error.
+The error versus $D$ curves allow us to choose a value of $D$ that
+results in an error below a chosen maximum.
 
@@
 
-All errors will be smaller than 0.1\% if one uses a distance-size ratio of 1.
-
-Even in the extreme case of a $30^\circ$ tesseroid with computation points at a
+For $D=1$ all errors are smaller than 0.1\%,
+even in the extreme case of a $30^\circ$ tesseroid with computation points at a
 height of 1 km above the tesseroid.
 
-A distance-size ratio of 1.5 is sufficient to achieve errors below 0.1\% in
-$g_z$.
+$D=1.5$ is sufficient to achieve errors below 0.1\% in $g_z$.
 
-A distance-size ratio of 8 is required to achieve errors below 0.1\% in
+$D=8$ is required to achieve errors below 0.1\% in
 the gravity gradient components.
 
-The error curve for the computation grid at 260 km height shows that
-the error is smaller at higher altitudes.
+The error curves for the computation at 260 km height shows that
+the error is smaller at higher altitudes for all fields.
 
-Computation at higher altitudes would not change the chosen values of the
-distance-size ratio for the potential and $g_z$.
+Computation at 260 km height does not produce errors above the 0.1\% threshold
+for the potential and $g_z$.
 
@@
 
-We computed the error curves for $g_{zz}$ only at different heights
+We computed the error curves for $g_{zz}$ at different heights
 (Figure~\ref{fig:gzz_error_height}).

### 2015-02-28 manuscript.tex
Typo fix

@@
 the number of point masses used
-and he evaluation of the trigonometric functions in
+and the evaluation of the trigonometric functions in
 equations~\ref{eq:tesspot}-\ref{eq:tesstensor} inside the inner loops.

### 2015-02-28 manuscript.tex
Remove old manuscript file and minor fixes

@@
 \righthead{Tesseroids}
-% manuscript number
-\ms{Submission}
+\ms{Submission} % manuscript number
 
@@
     email: leouieda@gmail.com
+    \footnotemark[2]Observat\'orio Nacional,
+    Rio de Janeiro, Brazil.
 }
@@
 The software is freely available
-on the Internet\footnote{http://www.leouieda.com/tesseroids}
-under the BSD 3-clause license.
+online\footnote{http://www.leouieda.com/tesseroids}
+under the BSD 3-clause open-source license.
 An archived version of the source code
@@
 (commonly referred to as ``popping'' the stack).
-Stacks are also known as ``Last-in-first-out'', or LIFO, structures.
+Stacks are also known as ``Last-in-first-out'', or LIFO, data structures.
 

### 2015-02-28 manuscript.tex
Restructured introduction
Still in the form of  loose phases. Need to tie them together.

@@
 
-We use the optimized formula of \citet{Grombein2013} with Cartesian integral
-kernels.
+Satellite missions dedicated to measuring the Earths gravity field
+(like CHAMP, GRACE, and GOCE)
+have provided geophysicists with almost uniform and global data coverage.
+
+These new data have enabled interpretations on regional and global scales
+\citep[e.g.][]{Reguzzoni2013,Braitenberg2015}.
+
+Modeling at such scales requires taking into account the curvature of the
+Earth and calculating gravity gradients as well as the traditional
+gravitational acceleration.
+
+A common approach to achieve this is
+to discretize the Earth into tesseroids (spherical prisms)
+instead of rectangular prisms.
+
+However, the integral formula for the gravitational effects of a tesseroid
+must be solved numerically.
+
+Approaches to this numerical integration include
+Taylor series expansion \citep{Heck2007, Grombein2013}
+and the Gauss-Legendre Quadrature \citep{Asgharzadeh2007}.
+
+Taylor series expansion produces accurate results at low latitudes but
+presents a decrease in accuracy towards the polar regions.
+
+This is attributed to tesseroids degenerating into an approximately triangular
+shape at the poles.
+
+The Gauss-Legendre Quadrature (GLQ) integration consists of approximating
+the volume integral by a weighted sum of the effect of point masses.
+
+An advantage of the GLQ approach is that the accuracy of integration can be
+controlled by the number of point masses used.
+
+A disadvantage is the increased computation time as the number of
+point masses increases.
+
+Thus, there is a trade-off between accuracy and computation time, a common
+theme in numerical methods.
+
+\citet{Wild-Pfeiffer2008} investigated the use of different mass elements,
+including tesseroids, to compute the gravitational effects of topographic
+masses.
+
+The author concludes that using tesseroids with GLQ integration gives the best
+results for near-zone computations.
+
+However, the question of how to determine the optimal parameters for GLQ
+integration remained open.
+
+Previous work by \citet{Ku1977} investigated the use of the Gauss-Legendre
+Quadrature in gravity forward modeling.
+
+\citet{Ku1977} numerically integrated the vertical component of the
+gravitational acceleration of right rectangular prisms.
+
+The author suggested that the accuracy of the GLQ integration depends on
+the ratio between distance to the computation point and the distance between
+adjacent point masses.
+
+Based on this, \citet{Ku1977} proposed an empirical criteria that the distance
+between adjacent point masses should be less than the distance to the computation point.
+
+\citet{Asgharzadeh2007} proposed using this criteria for the GLQ integration of
+the gravity gradient tensor of tesseroids.
+
+To our knowledge, there has been no investigation if the empirical criteria of
+\citet{Ku1977} is valid for gravity gradient components or for tesseroids.
+
+There has also been no attempt to quantify the error committed in the GLQ
+integration when applying the criteria of \citet{Ku1977}.
+
+\citet{Li2011} devised an algorithm to automatically enforce the criteria of
+\citet{Ku1977}.
+
+Their algorithm divides the tesseroid into smaller ones instead of increasing
+the number of point masses per tesseroid.
+
+This division effectively increases the total number of point masses
+because the number of points masses per tesseroid is fixed.
+
+The criterion used decide if a division is necessary is
+that the minimum distance to the computation point must be larger than
+the largest dimension of the tesseroid.
+
+This division is repeated recursively until all tesseroids obey the criteria.
+
+Then, GLQ integration is performed for each of the smaller tesseroids
+using the specified number of point masses.
+
+The advantage of this adaptive discretization over
+increasing the number of points masses is that the
+total total distribution of point masses will be greater
+only close to the computation point.
+
+This makes the adaptive discretization more computationally efficient.
+
+\citet{Grombein2013} developed optimized formula for the gravitational fields
+of tesseroids using Cartesian integral kernels.
 
@@
 
-The traditional spherical integral kernels suffer from a singularity at the
-poles \citep{Heck2007, Wild-Pfeiffer2008}.
-
 The Cartesian formula are numerically integrated using a Taylor series
@@
 
-The Taylor series expansion approach produces accurate results at low
-latitudes.
-
-The Taylor series expansion presents a decrease in accuracy towards the polar
-regions.
-
-The tesseroids have an approximately rectangular surface area at low latitudes
-but collapse into an approximately triangular shape at the poles.
-
 \citet{Grombein2013} use a near-zone separation to mitigate the increased error
-in high latitudes.
+at high latitudes.
 
@@
 
+The algorithm of \citet{Li2011} is similar to the near-zone separation used by
+\citet{Grombein2013}.
+
 The determination of an optimal size of the near-zone remains an open question
@@
 
-We use the Gauss-Legendre Quadrature (GLQ) to numerically integrate the Cartesian
-kernels instead of the Taylor series expansion \citep{Asgharzadeh2007}.
+We have implemented a modified version of the adaptive discretion of
+\citet{Li2011} into the open-source software package \textit{Tesseroids}.
 
-The GLQ integration consists of approximating the volume integral by a weighted sum of
-the effect of point masses.
-
-The point masses are distributed according to the roots of Legendre polynomials
-\citep{Hildebrand1987}.
-
-An advantage of the GLQ approach is that the accuracy of integration can be
-controlled by the number of point masses used.
-
-A disadvantage of the GLQ is the increased computation time as the number of
-point masses increases.
-
-There is a trade-off between accuracy and computation time.
-
-\citet{Ku1977} suggests that the accuracy of the GLQ integration depends on
-the ratio between distance to the computation point and the distance between
-adjacent point masses.
-
-\citet{Ku1977} proposes an empirical criteria that the distance between adjacent
-point masses should be less than the distance to the computation point.
-
-This empirical criteria was used by \citet{Asgharzadeh2007} as a way to
-choose the number of point masses used in the GLQ integration.
-
-They suggested using this criteria for gravitational attraction as well as
-the gravity gradient tensor (or Marussi tensor) of a tesseroid.
-
-\citet{Ku1977} presented results for the vertical component of the
-gravitational acceleration ($g_z$) caused by a right rectangular prism.
-
-To our knowledge, there has been no investigation if the empirical criteria of
-\citet{Ku1977} is valid for the second derivatives of the
-gravitational potential.
-
-There has also been no attempt to quantify the error committed in the GLQ
-integration of gravity gradients when applying the criteria of \citet{Ku1977}.
-
-\citet{Li2011} devised an automatic algorithm to enforce the criteria of
-\citet{Ku1977}.
-
-Their algorithm keeps the number of points masses per tesseroid fixed.
-
-They then check the ratio between the minimum distance to the computation point
-and the largest dimension of the tesseroid.
-
-If the ratio below a specified threshold value, the tesseroid is divided into
-smaller tesseroids.
-
-This division is repeated recursively until all tesseroids obey the criteria.
-
-The GLQ integration is then performed for each of the smaller tesseroids
-using the fixed number of point masses.
-
-The algorithm of \citet{Li2011} is similar to the near-zone separation used by
-\citet{Grombein2013}.
-
-The advantage of the adaptive discretization of \citet{Li2011} over simply
-increasing the number of points masses in the GLQ is that the
-amount of point masses will be greater only close to the computation point.
-
-This makes the adaptive discretization more computationally efficient.
-
-This article describes the software \textit{Tesseroids} and the implementation
-of the adaptive discretization algorithm.
+The software uses the Cartesian formula of \citet{Grombein2013} for improved
+performance and robustness.
 
@@
 
-We will attempt to quantify the error committed in the computation of the
-gravity gradients using the GLQ with adaptive discretization.
+This article describes the software \textit{Tesseroids} and the implementation
+of our modified adaptive discretization algorithm.
 
-We will establish an empirical relationship between the integration error and
-the ratio between the distance to the computation point and the dimensions of
-the tesseroid.
+We also present a numerical investigation of the error committed
+in the computations.
 
-This will allow one to choose the appropriate amount of discretization to
-achieve the desired accuracy.
+These results allow us calibrate the adaptive discretization algorithm
+separately for the gravitational potential, gravitational acceleration,
+as well as the gravity gradient tensor components.
 

### 2015-02-28 manuscript.tex
Tied the intro together. Need proof-reading

@@
 have provided geophysicists with almost uniform and global data coverage.
-
 These new data have enabled interpretations on regional and global scales
 \citep[e.g.][]{Reguzzoni2013,Braitenberg2015}.
-
 Modeling at such scales requires taking into account the curvature of the
@@
 gravitational acceleration.
-
 A common approach to achieve this is
-to discretize the Earth into tesseroids (spherical prisms)
+to discretize the Earth into tesseroids (Figure~\ref{fig:tesseroid})
 instead of rectangular prisms.
-
 However, the integral formula for the gravitational effects of a tesseroid
 must be solved numerically.
-
 Approaches to this numerical integration include
@@
 and the Gauss-Legendre Quadrature \citep{Asgharzadeh2007}.
-
 Taylor series expansion produces accurate results at low latitudes but
 presents a decrease in accuracy towards the polar regions.
-
 This is attributed to tesseroids degenerating into an approximately triangular
 shape at the poles.
-
 The Gauss-Legendre Quadrature (GLQ) integration consists of approximating
 the volume integral by a weighted sum of the effect of point masses.
-
 An advantage of the GLQ approach is that the accuracy of integration can be
 controlled by the number of point masses used.
-
 A disadvantage is the increased computation time as the number of
 point masses increases.
-
 Thus, there is a trade-off between accuracy and computation time, a common
 theme in numerical methods.
-
 \citet{Wild-Pfeiffer2008} investigated the use of different mass elements,
@@
 masses.
-
 The author concludes that using tesseroids with GLQ integration gives the best
 results for near-zone computations.
-
 However, the question of how to determine the optimal parameters for GLQ
@@
 Quadrature in gravity forward modeling.
-
 \citet{Ku1977} numerically integrated the vertical component of the
 gravitational acceleration of right rectangular prisms.
-
 The author suggested that the accuracy of the GLQ integration depends on
@@
 adjacent point masses.
-
 Based on this, \citet{Ku1977} proposed an empirical criteria that the distance
 between adjacent point masses should be less than the distance to the computation point.
-
 \citet{Asgharzadeh2007} proposed using this criteria for the GLQ integration of
 the gravity gradient tensor of tesseroids.
-
 To our knowledge, there has been no investigation if the empirical criteria of
 \citet{Ku1977} is valid for gravity gradient components or for tesseroids.
-
 There has also been no attempt to quantify the error committed in the GLQ
@@
 \citet{Ku1977}.
-
 Their algorithm divides the tesseroid into smaller ones instead of increasing
 the number of point masses per tesseroid.
-
 This division effectively increases the total number of point masses
 because the number of points masses per tesseroid is fixed.
-
 The criterion used decide if a division is necessary is
@@
 the largest dimension of the tesseroid.
-
 This division is repeated recursively until all tesseroids obey the criteria.
-
 Then, GLQ integration is performed for each of the smaller tesseroids
 using the specified number of point masses.
-
 The advantage of this adaptive discretization over
@@
 only close to the computation point.
-
 This makes the adaptive discretization more computationally efficient.
@@
 of tesseroids using Cartesian integral kernels.
-
-The Cartesian kernels are faster to compute.
-
-The Cartesian kernels don't have singularities in the poles.
-
+These formula are faster to compute and don't have singularities at the poles
+like their spherical counterparts.
 The Cartesian formula are numerically integrated using a Taylor series
-expansion as proposed by \citet{Heck2007}.
-
+expansion as per \citet{Heck2007}.
 \citet{Grombein2013} use a near-zone separation to mitigate the increased error
 at high latitudes.
-
-In the so called "near-zone" of the computation point they use a finner
+In the so called ``near-zone'' of the computation point they use a finner
 discretization with smaller tesseroids.
-
-This is accomplished by dividing the tesseroids along the horizontal
+This is accomplished by dividing the tesseroids along their horizontal
 dimensions.
-
-The algorithm of \citet{Li2011} is similar to the near-zone separation used by
-\citet{Grombein2013}.
-
-The determination of an optimal size of the near-zone remains an open question
-\citep{Grombein2013}.
+However, the determination of an optimal size of the near-zone remains an
+open question \citep{Grombein2013}.
 
@@
 \citet{Li2011} into the open-source software package \textit{Tesseroids}.
-
 The software uses the Cartesian formula of \citet{Grombein2013} for improved
 performance and robustness.
-
 Previous versions of \textit{Tesseroids} have been used by, e.g.,
@@
 Braitenberg2011, Fullea2014}.
-
 This article describes the software \textit{Tesseroids} and the implementation
 of our modified adaptive discretization algorithm.
-
 We also present a numerical investigation of the error committed
 in the computations.
-
 These results allow us calibrate the adaptive discretization algorithm

### 2015-03-01 manuscript.tex
Update website url
hyperlinks, section on reproducing, split long eqs

@@
-\documentclass[paper,twocolumn]{geophysics}
+\pdfoutput=1
+%\documentclass[paper,twocolumn]{geophysics}
 %\documentclass[manuscript]{geophysics}
 %\documentclass[manuscript,endfloat]{geophysics}
+\documentclass[twocolumn]{article}
+\usepackage{natbib}
 \usepackage{graphicx}
 \usepackage{amsmath}
+\usepackage{hyperref}
 \newcommand{\mitbf}[1]{\mathbf{{#1}}}
 
-\begin{document}
-
+\hypersetup{
+    colorlinks=true,
+    citecolor=blue,
+    linkcolor=blue,
+    urlcolor=blue,
+    pdfauthor={Uieda et al.},
+    pdftitle={Tesseroids: forward modeling gravitational fields in spherical
+              coordinates}
+}
 \title{
@@
 }
-\righthead{Tesseroids}
-\ms{Submission} % manuscript number
-
 \author{
@@
 
-\address{
+\begin{document}
+
+\maketitle
+
+%\righthead{Tesseroids}
+%\ms{Submission} % manuscript number
+
+%\address{
+    \noindent
     \footnotemark[1]Universidade do Estado do Rio de Janeiro,
@@
     Rio de Janeiro, Brazil.
-}
-
-\maketitle
+%}
 
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figs/tesseroid}
+    \includegraphics[width=\columnwidth]{figs/tesseroid}
     \caption{
@@
         \int\limits_{r_1}^{r_2}
+        I_{\alpha\beta}
+        \ \kappa\ dr^\prime d\phi^\prime d\lambda^\prime.
+    \label{eq:tesstensor}
+\end{equation}
+\begin{equation}
+    I_{\alpha\beta} =
     \left(
@@
     \right)
-        \kappa\ dr^\prime d\phi^\prime d\lambda^\prime.
-    \label{eq:tesstensor}
+    \label{eq:tesstensorkernel}
 \end{equation}
@@
 
-\begin{eqnarray}
-    \Delta_x &=& r^\prime\cos\phi\sin\phi^\prime - \sin\phi\cos\phi^\prime
-                 \cos(\lambda^\prime - \lambda), \\
-    \Delta_y &=& r^\prime \cos \phi^\prime \sin(\lambda^\prime - \lambda), \\
-    \Delta_z &=& r^\prime \cos \psi - r,\\
-    \kappa &=& {r^\prime}^2 \cos \phi^\prime, \\
-    \ell &=& \sqrt{{r^\prime}^2 + r^2 - 2 r^\prime r \cos \psi}, \\
-    \cos\psi &=& \sin\phi\sin\phi^\prime + \cos\phi\cos\phi^\prime
+\begin{equation}
+    \Delta_x = r^\prime(\cos\phi\sin\phi^\prime - \sin\phi\cos\phi^\prime
+               \cos(\lambda^\prime - \lambda)),
+\end{equation}
+\begin{equation}
+    \Delta_y = r^\prime \cos \phi^\prime \sin(\lambda^\prime - \lambda),
+\end{equation}
+\begin{equation}
+    \Delta_z = r^\prime \cos \psi - r,
+\end{equation}
+\begin{equation}
+    \kappa = {r^\prime}^2 \cos \phi^\prime,
+\end{equation}
+\begin{equation}
+    \ell = \sqrt{{r^\prime}^2 + r^2 - 2 r^\prime r \cos \psi},
+\end{equation}
+\begin{equation}
+    \cos\psi = \sin\phi\sin\phi^\prime + \cos\phi\cos\phi^\prime
                  \cos(\lambda^\prime - \lambda).
-\end{eqnarray}
+\end{equation}
 
@@
     f(r^\prime, \lambda^\prime, \phi^\prime)
-    dr^\prime d\lambda^\prime d\phi^\prime
+    d\Omega
     \approx
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figs/vary-height-and-order}
+    \includegraphics[width=0.7\columnwidth]{figs/vary-height-and-order}
     \caption{
@@
 The software is freely available
-online\footnote{http://www.leouieda.com/tesseroids}
+online\footnote{
+\href{http://tesseroids.leouieda.com}{http://tesseroids.leouieda.com}}
 under the BSD 3-clause open-source license.
@@
 
+\subsection{Reproducing analysis and results}
+
 All figures and numerical experiments in this article
-were produced by IPython notebooks\footnote{http://ipython.org}
+were produced by IPython notebooks\footnote{
+\href{http://ipython.org}{http://ipython.org}}
 \citep{Perez2007}.
@@
 with text, equations, and the figures generated by the code.
-These files,
+The notebook files,
 as well as instructions for installing and running the analysis,
@@
 Alternatively,
-the notebooks and can also be downloaded
-or viewed in a non-executable form from
-an online repository\footnote{https://github.com/pinga-lab/paper-tesseroids}.
+all accompanying material can be downloaded
+from an online source code repository\footnote{
+\href{https://github.com/pinga-lab/paper-tesseroids}{https://github.com/pinga-lab/paper-tesseroids}}
+or a persistent archive\footnote{
+\href{}{Zenodo DOI}}.
 
@@
     d = \left[
-        r^2 + r_t^2 - 2 r r_t
-        \cos(\sin\phi\sin\phi_t + \cos\phi\cos\phi_t
-            \cos(\lambda - \lambda_t))
+        r^2 + r_t^2 - 2 r r_t \cos\psi_t
         \right]^{\frac{1}{2}}.
     \label{eq:distance}
+\end{equation}
+\begin{equation}
+    \cos\psi_t =
+        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t)
 \end{equation}
@@
 \end{equation}
+
+\noindent
+$L_\lambda$ and $L_\phi$ are arc-distances measured along the top surface of
+the tesseroid (Figure~\ref{fig:division}a).
+Specifically, $L_\lambda$ is measured long the middle latitude of the
+tesseroid ($\phi_t$).
 
@@
 4, 38, and 936.
-Code to perform the adaptive discretization
-and produce Figure~\ref{fig:division} is available in
-the IPython notebook \texttt{methods\_figures.ipynb}.
 
@@
     \centering
-    \includegraphics{figs/tesseroid-split}
+    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
     \caption{

### 2015-03-01 manuscript.tex
Added citations to pandas, mpl, Mayavi

@@
 \newcommand{\mitbf}[1]{\mathbf{{#1}}}
+
+% To leave comments on Overleaf
+\usepackage[colorinlistoftodos]{todonotes}
 
@@
 \section{Introduction}
+
+\todo[inline]{Here's a comment!}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid}
+    \includegraphics[width=0.4\textwidth]{figs/tesseroid}
     \caption{
@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{figs/vary-height-and-order}
+    \includegraphics[width=0.4\textwidth]{figs/vary-height-and-order}
     \caption{
@@
 with text, equations, and the figures generated by the code.
+We used \textit{pandas} by \citet{Mckinney2010}
+to perform the data analysis,
+\textit{matplotlib} by \citet{Hunter2007} for graphs,
+and \textit{Mayavi} by \citet{Ramachandran2011} for 3D graphs.
+
 The notebook files,
@@
 or a persistent archive\footnote{
-\href{}{Zenodo DOI}}.
+\href{}{\todo[inline]{Replace with the Zenodo DOI for the repo.}}}.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
+    \includegraphics[width=0.4\textwidth]{figs/tesseroid-split}
     \caption{

### 2015-03-01 manuscript.tex
Adding text to results section

@@
 
-$D$ controls both the accuracy of the integration and the computation time.
+$D$ indirectly controls both the accuracy of the integration
+and the computation time.
 
-The value chosen for $D$ determines how many divisions are made.
+The specific value chosen for $D$ determines how many divisions will be made
+(Figure~\ref{fig:division}).
 
@@
 
-To our knowledge,
-there has been no further investigation
-into the relationship between
-values of the distance-size ratio and the computation error.
+In this section, we investigate
+the relationship between
+values of the distance-size ratio
+and the computation error.
 
-It is also not clear whether
-the same value of $D$ yields compatible error levels
-for the gravitational attraction
-and the gradient tensor components.
+We perform the error analysis for the gravitational potential,
+acceleration, and gradient tensor components
+to evaluate if the same value of $D$ yields compatible error levels
+for different fields.
 
-We must determine the relationship between the distance-size ratio $D$ and
-the computation error.
+We establish a homogeneous spherical shell as a reference against which to
+compare the computed tesseroid fields.
 
-We need a reference against which to compare the computed tesseroid fields.
+The shell has analytical solutions for the gravitational potential,
+attraction, and gradient tensor along the polar axis
+\citep{Grombein2013}.
 
-The spherical shell is a convenient reference model.
+We used a spherical shell with a thickness of 1 km,
+density of $2670\ kg.m^{-3}$,
+bottom at height 0 km above the reference sphere,
+and top at 1 km height.
 
-The spherical shell has a analytical solutions for the gravitational potential,
-attraction, and gradient tensor along the polar axis.
+We discretized the spherical shell along the horizontal dimensions into
+a regular mesh of tesseroids.
 
-\citep{Grombein2013} provides the equations for
-the gravitational potential, attraction, and gravity gradient components
-of a homogeneous spherical shell.
+Figure~\ref{fig:glqerrorsample} shows that the largest errors are concentrated
+on top of the tesseroid.
 
-We used a spherical shell with a thickness of 1 km.
+If we calculate the tesseroid fields at a single point, that point might not
+coincide with the point of largest error.
 
-The density of the shell is $2670\ kg.m^{-3}$.
+Performing the error analysis on calculations at a single point will not be
+representative of the true error.
 
-The bottom of the shell is at height 0 km above the reference sphere.
+Fortunately, the symmetry of the shell allows us to consider the computation point at any
+geographic coordinate.
 
-The top of the shell is at 1 km height.
+We calculate the effect of the tesseroid model of the spherical shell
+on a regular grid.
 
-We can use the spherical shell values to calculate the error of the tesseroid
-computations for varying values of $D$.
+The effect of the shell will be same along the entire grid.
 
-We discretized the spherical shell into tesseroids
-with dimensions of $1^\circ \times 1^\circ \times 1\ km$.
-
-We calculate the effect of a tesseroid model on a regular grid.
+We perform the analysis on the largest error.
 

### 2015-03-01 manuscript.tex
Move Reproducing to the end of Implementation

@@
 %\documentclass[manuscript,endfloat]{geophysics}
-\documentclass[twocolumn]{article}
+\documentclass{article}
 \usepackage{natbib}
@@
 written in the C programming language.
-The command-line interface follows the Unix principals of
-(1) making each program do one thing well,
-(2) using plain text for input and output,
-(3) making each program a ``filter'' that can be used in conjunction with other
-programs.
 The package includes programs to calculate
@@
 and the adaptive discretization of tesseroids.
-
-\subsection{Reproducing analysis and results}
-
-All figures and numerical experiments in this article
-were produced by IPython notebooks\footnote{
-\href{http://ipython.org}{http://ipython.org}}
-\citep{Perez2007}.
-These files combine the analysis and figure generation source code
-with text, equations, and the figures generated by the code.
-We used \textit{pandas} by \citet{Mckinney2010}
-to perform the data analysis,
-\textit{matplotlib} by \citet{Hunter2007} for graphs,
-and \textit{Mayavi} by \citet{Ramachandran2011} for 3D graphs.
-
-The notebook files,
-as well as instructions for installing and running the analysis,
-are distributed along with the source code archive
-that accompanies this article.
-Alternatively,
-all accompanying material can be downloaded
-from an online source code repository\footnote{
-\href{https://github.com/pinga-lab/paper-tesseroids}{https://github.com/pinga-lab/paper-tesseroids}}
-or a persistent archive\footnote{
-\href{}{\todo[inline]{Replace with the Zenodo DOI for the repo.}}}.
 
@@
 
+\subsection{Reproducing analysis and results}
+
+All figures and numerical experiments in this article
+were produced in IPython notebooks\footnote{
+\href{http://ipython.org}{http://ipython.org}}
+\citep{Perez2007}.
+These files combine the analysis and figure generation source code
+with text, equations, and the figures generated by the code.
+We used \textit{pandas} by \citet{Mckinney2010}
+to perform the data analysis,
+\textit{matplotlib} by \citet{Hunter2007} for graphs,
+and \textit{Mayavi} by \citet{Ramachandran2011} for 3D graphs.
+
+The notebook files,
+as well as instructions for installing and running the analysis,
+are distributed along with the source code archive
+that accompanies this article.
+Alternatively,
+all accompanying material can be downloaded
+from an online source code repository\footnote{
+\href{https://github.com/pinga-lab/paper-tesseroids}{https://github.com/pinga-lab/paper-tesseroids}}
+or a persistent archive\footnote{
+\href{}{\todo[inline]{Replace with the Zenodo/figshare DOI for the repo.}}}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+If we calculate the tesseroid fields at a single point, that point might not
+coincide with the point of largest error.
+
 Figure~\ref{fig:glqerrorsample} shows that the largest errors are concentrated
 on top of the tesseroid.
-
-If we calculate the tesseroid fields at a single point, that point might not
-coincide with the point of largest error.
 
@@
 
-We perform the analysis on the largest error.
+We compute the differences between the effects of the shell and the tesseroid
+model on the grid and keep only the largest error.
 

### 2015-03-02 manuscript.tex
Placeholder for the 2 analysis figures
The error x distance-size ratio curves for V, gz and gzz.
The curves for different heights for gzz.

@@
 \pdfoutput=1
-%\documentclass[paper,twocolumn]{geophysics}
+\documentclass[paper,twocolumn]{geophysics}
 %\documentclass[manuscript]{geophysics}
 %\documentclass[manuscript,endfloat]{geophysics}
-\documentclass{article}
+%\documentclass[twocolumn]{article}
 \usepackage{natbib}
@@
 
-% To leave comments on Overleaf
+% To leave comments
 \usepackage[colorinlistoftodos]{todonotes}
@@
 
-%\righthead{Tesseroids}
-%\ms{Submission} % manuscript number
+\righthead{Tesseroids}
+\ms{Submission} % manuscript number
 
-%\address{
+\address{
     \noindent
@@
     Rio de Janeiro, Brazil.
-%}
+}
 
@@
 in the open-source package \textit{Tesseroids}.
-The software is freely available
-online\footnote{
+It is freely available
+online through the software website\footnote{
 \href{http://tesseroids.leouieda.com}{http://tesseroids.leouieda.com}}
@@
 An archived version of the source code
-is available as part of this article.
+is available as part of this article
+and online\footnote{\href{}{\todo[inline]{insert a ZENODO doi}}}.
 
@@
 
+\begin{figure}
+    %\centering
+    %\includegraphics[width=0.4\textwidth]{figs/distance-size-curves}
+    \missingfigure{}
+    \caption{
+        The maximum difference between the computed tesseroid and shell effects
+        as a function of the distance-size ratio $D$
+        for (a) the gravitational potential, (b) $g_z$, and (c) $g_{zz}$.
+        The difference is given as a percentage of the shell effect.
+        Curves correspond to the different tesseroid models:
+        $1^\circ \times 1^\circ$ tesseroids with computation grid at
+        2 km height at the polar region,
+        $1^\circ \times 1^\circ$ tesseroids with computation grid at
+        2 km height at the equatorial region,
+        $1^\circ \times 1^\circ$ tesseroids with computation grid at
+        260 km height at the polar region,
+        and
+        $30^\circ \times 30^\circ$ tesseroids with computation grid at
+        2 km height at the polar region.
+    }
+    \label{fig:dist-size-curves}
+\end{figure}
+
+\begin{figure}
+    %\centering
+    %\includegraphics[width=0.4\textwidth]{figs/gzz-with-height}
+    \missingfigure{}
+    \caption{
+        Variation of $g_{zz}$ difference curves for computation grids at
+        varying heights.
+    }
+    \label{fig:gzz-with-height}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-03-02 manuscript.tex
Outline of results section
Need to review and join sentences.
Figures are missing and have only placeholders

@@
 
-$D$ indirectly controls both the accuracy of the integration
-and the computation time.
-
 The specific value chosen for $D$ determines how many divisions will be made
@@
 
-\citet{Li2011} provide a few examples of values for this ratio in the case of
-the vertical component of the gravitational attraction.
+Thus, $D$ indirectly controls both the accuracy of the integration
+and the computation time.
 
-In this section, we investigate
-the relationship between
-values of the distance-size ratio
-and the computation error.
+In this section, we investigate the relationship between
+values of the distance-size ratio and the integration error.
 
-We perform the error analysis for the gravitational potential,
+We perform the analysis for the gravitational potential,
 acceleration, and gradient tensor components
@@
 
-We establish a homogeneous spherical shell as a reference against which to
-compare the computed tesseroid fields.
+The reference against which we compare the computed tesseroid fields
+is a homogeneous spherical shell.
 
-The shell has analytical solutions for the gravitational potential,
-attraction, and gradient tensor along the polar axis
-\citep{Grombein2013}.
+The shell has analytical solutions along the polar axis
+\citep{Grombein2013}
+and can be perfectly discretized into tesseroids.
 
-We used a spherical shell with a thickness of 1 km,
+We chose a spherical shell with a thickness of 1 km,
 density of $2670\ kg.m^{-3}$,
@@
 
-We discretized the spherical shell along the horizontal dimensions into
-a regular mesh of tesseroids.
+We produced tesseroid models of the shell by discretizing it along the
+horizontal dimensions into a regular mesh.
 
-If we calculate the tesseroid fields at a single point, that point might not
-coincide with the point of largest error.
-
-Figure~\ref{fig:glqerrorsample} shows that the largest errors are concentrated
+Figure~\ref{fig:glqerrorsample} shows that the largest errors are spread over
 on top of the tesseroid.
 
-Performing the error analysis on calculations at a single point will not be
-representative of the true error.
+Thus, calculating the tesseroid fields at a single point might not
+capture the point of largest error.
 
-Fortunately, the symmetry of the shell allows us to consider the computation point at any
-geographic coordinate.
+Instead, we calculate the effect of the tesseroid model on a regular grid
+of $10 \times 10$ points.
 
-We calculate the effect of the tesseroid model of the spherical shell
-on a regular grid.
+Fortunately, the symmetry of the shell allows us to consider the computation
+point at any geographic coordinate.
 
-The effect of the shell will be same along the entire grid.
+Therefore, the effect of the shell will be same along the entire grid.
 
 We compute the differences between the effects of the shell and the tesseroid
-model on the grid and keep only the largest error.
+model on the grid and consider only the largest error.
 
-The grid is at a constant height of 2 km (1 km above the shell).
+By placing the grid on top of a particular tesseroid, we increase the chances
+of capturing the true largest integration error.
 
-The grid is located on top of a tesseroid located at the North pole.
-
-The computation point can be considered at any point with equal radial
-coordinate located outside the shell because of the symmetry of the shell.
-
-We calculate the difference between the shell values and the values computed
-from tesseroids.
-
-We keep only the largest difference.
-
-Calculating on a grid above a tesseroid is important because
-the largest errors in the GLQ integration occur directly on top of the
-tesseroid but not exactly at it's center, as can be seen in
-Figure~\ref{fig:glqerrorsample}.
-
-The difference between shell and tesseroid values would be underestimated
-if the computation was performed only at a single point on the polar axis.
-
-We repeated the tesseroid calculations with values of the distance-size ratio
+We calculate the errors for values of the distance-size ratio
 $D$ varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
 
-Figure~\ref{fig:tess_vs_shell} shows the maximum difference between the
-shell and tesseroid fields as a function of $D$.
+We repeated the analysis in four different numerical experiments,
+each with computation grids at different locations and
+different tesseroid model sizes.
+
+Table~\ref{tbl:experiment} describes the different numerical
+experiments and the corresponding parameters of the computation grid and
+tesseroid model.
+
+Figure~\ref{fig:dist-size-curves} shows
+the maximum difference between the shell and tesseroid fields
+as a function of $D$ for the four numerical experiments.
 
@@
 
-The error versus $D$ curves allow us to choose a value of $D$ that
-results in an error below a chosen maximum.
+We established a maximum tolerated error of $0.1\%$, represented by the
+horizontal solid lines in Figure~\ref{fig:dist-size-curves}.
 
-We chose 0.1\% as the maximum allowed error.
+For the potential $V$, a distance-size ratio $D=1$ guarantees that the curves
+for all experiments are below the $0.1\%$ error threshold.
 
-We performed the computations again for three different scenarios: at 260 km
-height (the mean orbit height of the GOCE satellite), on top of a tesseroid at
-the equator, and for a $30^\circ \times 30^\circ \times 1\ km$ tesseroid model
-of the shell.
+For $g_z$, the same is achieved with $D=1.5$.
 
-We performed these computations to investigate if the GLQ error decays the same
-way under different scenarios.
+Conversely, $g_{zz}$ requires a value of $D=8$ to achieve an error level of
+$0.1\%$.
 
-The error curves for the gravitational potential show increased errors at the
-equator and for a $30^\circ$ tesseroid.
+For a computation height of 260 km, the error curve for $G_{zz}$
+intercepts the error threshold line at $D=2.5$.
 
-For $D=1$ all errors are smaller than 0.1\%,
-even in the extreme case of a $30^\circ$ tesseroid with computation points at a
-height of 1 km above the tesseroid.
+This behavior suggests that the error curves for $g_{zz}$ might depend on the
+computation height.
 
-$D=1.5$ is sufficient to achieve errors below 0.1\% in $g_z$.
+To test this hypothesis, we computed the error curves for $g_{zz}$ at
+heights 2, 10, 50, 150, and 260 km.
 
-$D=8$ is required to achieve errors below 0.1\% in
-the gravity gradient components.
+Figure~\ref{fig:gzz-with-height} shows the results for $g_{zz}$ at varying
+computation heights.
 
-The error curves for the computation at 260 km height shows that
-the error is smaller at higher altitudes for all fields.
+Notice that the distance-size ratio needed to achieve $0.1\%$ accuracy
+decreases as the computation height increases.
 
-Computation at 260 km height does not produce errors above the 0.1\% threshold
-for the potential and $g_z$.
+For example, computation at 260 km height requires $D=2.5$ whereas at 10 km
+height a value of $D=5.5$ is required to achieve the same accuracy.
 
-The computation height changes the value of the distance-size ratio required
-for the gravity gradient components.
+One can take advantage of this behavior to reduce the distance-size ratio for
+computations of the gravity gradient tensor at high altitudes,
+saving computation time.
 
-We computed the error curves for $g_{zz}$ at different heights
-(Figure~\ref{fig:gzz_error_height}).
+We have implemented the values of the distance-size ratio producing $0.1\%$
+accuracy as defaults for the software \textit{Tesseroids}.
 
-The distance-size ratio can be reduced to 4 if the computation is done at 50 km
-height.
+We chose the conservative value of $D=8$ for the gravity gradient components
+as a fail-safe alternative.
 
-One can use a distance-size ratio of 2.5 for computations at 260 km height.
+Users can control the value of $D$ used in the computations through
+command-line arguments to achieve greater performance as a trade-off for
+accuracy.
 
-These reductions in the distance-size ratio can reduce the computation time
-required to achieve the desired.
-
-The software \textit{Tesseroids} includes default values of the distance-size
-ratio.
-
-We used the values required to achieve 0.1\% accuracy: 1 for the potential,
-1.5 for the gravitational attraction, and 8 for the gravity gradient
-components.
-
-The value of $D=8$ for the gravity gradients is conservative.
-
-We chose it to ensure the accuracy of the computation in most cases.
-
-The experienced user of the software can set smaller values using command-line
-options.
-
-The smaller values result in less computation time.
+\begin{table*}
+\centering
+\begin{tabular}{cccc}
+    \hline
+    ~             & Grid location & Grid height & Tesseroid size \\
+    \hline
+    Experiment 1  & 89N--90N/0E--1E & 2 km & $1^\circ \times 1^\circ$ \\
+    Experiment 2  & 0N--1N/0E--1E & 2 km & $1^\circ \times 1^\circ$ \\
+    Experiment 3  & 89N--90N/0E--1E & 260 km & $1^\circ \times 1^\circ$ \\
+    Experiment 4  & 60N--90N/0E--30E & 2 km & $30^\circ \times 30^\circ$ \\
+    \hline
+\end{tabular}
+\caption{Parameters of the numerical experiments.}
+\label{tbl:experiment}
+\end{table*}
 

### 2015-03-02 manuscript.tex
Error X ratio figure
Included gzz with height plot and better captions

@@
 \begin{figure}
-    %\centering
-    %\includegraphics[width=0.4\textwidth]{figs/distance-size-curves}
-    \missingfigure{}
+    \centering
+    \includegraphics[width=0.4\textwidth]{figs/distance-size-curves}
+    %\missingfigure{}
     \caption{
@@
         The difference is given as a percentage of the shell effect.
-        Curves correspond to the different tesseroid models:
-        $1^\circ \times 1^\circ$ tesseroids with computation grid at
-        2 km height at the polar region,
-        $1^\circ \times 1^\circ$ tesseroids with computation grid at
-        2 km height at the equatorial region,
-        $1^\circ \times 1^\circ$ tesseroids with computation grid at
-        260 km height at the polar region,
-        and
-        $30^\circ \times 30^\circ$ tesseroids with computation grid at
-        2 km height at the polar region.
+        Curves correspond to the different tesseroid models and computation
+        grids shown in Table~\ref{tbl:experiment}.
+        The horizontal solid black line marks the established error threshold
+        of $0.1\%$.
     }
@@
 \begin{figure}
-    %\centering
-    %\includegraphics[width=0.4\textwidth]{figs/gzz-with-height}
-    \missingfigure{}
+    \centering
+    \includegraphics[width=0.4\textwidth]{figs/gzz-with-height}
+    %\missingfigure{}
     \caption{
-        Variation of $g_{zz}$ difference curves for computation grids at
-        varying heights.
+        Difference between the computed $g_{zz}$ for the spherical shell and
+        the tesseroid model at different heights. Curves show the maximum
+        difference as a percentage of the shell value.
+        The horizontal solid black line marks the established error threshold
+        of $0.1\%$.
     }

### 2015-03-02 manuscript.tex
Finished analysis sections
New to revise the text  later.

@@
 
-\righthead{Tesseroids}
+\righthead{Tesseroids: modeling in spherical coordinates}
 \ms{Submission} % manuscript number
@@
 is available as part of this article
-and online\footnote{\href{}{\todo[inline]{insert a ZENODO doi}}}.
+and online\footnote{
+\href{}{\todo[inline]{Replace with the Zenodo DOI for the Tesseroids
+source code.}}}.
 
@@
 
-\subsection{Reproducing analysis and results}
+\subsection{Reproducing the analysis and results}
 
@@
 with text, equations, and the figures generated by the code.
-We used \textit{pandas} by \citet{Mckinney2010}
+We used the libraries
+\textit{pandas} by \citet{Mckinney2010}
 to perform the data analysis,
@@
 or a persistent archive\footnote{
-\href{}{\todo[inline]{Replace with the Zenodo/figshare DOI for the repo.}}}.
+\href{}{\todo[inline]{Replace with the Zenodo/figshare DOI for the repository.}}}.
 
@@
 is the distance-size ratio $D$ (equation~\ref{eq:condition}).
-
 The specific value chosen for $D$ determines how many divisions will be made
 (Figure~\ref{fig:division}).
-
 Thus, $D$ indirectly controls both the accuracy of the integration
 and the computation time.
-
 In this section, we investigate the relationship between
-values of the distance-size ratio and the integration error.
-
+the distance-size ratio and the integration error.
 We perform the analysis for the gravitational potential,
@@
 is a homogeneous spherical shell.
-
 The shell has analytical solutions along the polar axis
@@
 and can be perfectly discretized into tesseroids.
-
 We chose a spherical shell with a thickness of 1 km,
@@
 and top at 1 km height.
-
 We produced tesseroid models of the shell by discretizing it along the
@@
 on top of the tesseroid.
-
 Thus, calculating the tesseroid fields at a single point might not
 capture the point of largest error.
-
 Instead, we calculate the effect of the tesseroid model on a regular grid
 of $10 \times 10$ points.
-
 Fortunately, the symmetry of the shell allows us to consider the computation
-point at any geographic coordinate.
-
+point at any geocentric coordinate.
 Therefore, the effect of the shell will be same along the entire grid.
-
 We compute the differences between the effects of the shell and the tesseroid
 model on the grid and consider only the largest error.
-
 By placing the grid on top of a particular tesseroid, we increase the chances
@@
 $D$ varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
-
-We repeated the analysis in four different numerical experiments,
+Furthermore, we repeated the analysis in four different numerical experiments,
 each with computation grids at different locations and
 different tesseroid model sizes.
-
 Table~\ref{tbl:experiment} describes the different numerical
@@
 the maximum difference between the shell and tesseroid fields
-as a function of $D$ for the four numerical experiments.
-
+as a function of $D$ for the four experiments.
 The differences are given as a percentage of the shell value.
-
 We established a maximum tolerated error of $0.1\%$, represented by the
 horizontal solid lines in Figure~\ref{fig:dist-size-curves}.
+Only results for the gravitational potential, $g_z$, and $g_{zz}$ are shown.
+The results for the other diagonal components of the gravity gradient tensor
+are similar to $g_{zz}$.
+Figures for these components can be found in the supplementary material
+(see section "Reproducing the analysis and results").
 
@@
 for all experiments are below the $0.1\%$ error threshold.
-
 For $g_z$, the same is achieved with $D=1.5$.
-
 Conversely, $g_{zz}$ requires a value of $D=8$ to achieve an error level of
 $0.1\%$.
-
 For a computation height of 260 km, the error curve for $G_{zz}$
 intercepts the error threshold line at $D=2.5$.
-
 This behavior suggests that the error curves for $g_{zz}$ might depend on the
 computation height.
-
 To test this hypothesis, we computed the error curves for $g_{zz}$ at
 heights 2, 10, 50, 150, and 260 km.
-
 Figure~\ref{fig:gzz-with-height} shows the results for $g_{zz}$ at varying
 computation heights.
-
 Notice that the distance-size ratio needed to achieve $0.1\%$ accuracy
 decreases as the computation height increases.
-
 For example, computation at 260 km height requires $D=2.5$ whereas at 10 km
 height a value of $D=5.5$ is required to achieve the same accuracy.
-
 One can take advantage of this behavior to reduce the distance-size ratio for
@@
 We have implemented the values of the distance-size ratio producing $0.1\%$
-accuracy as defaults for the software \textit{Tesseroids}.
-
+accuracy determined above as defaults for the software \textit{Tesseroids}.
 We chose the conservative value of $D=8$ for the gravity gradient components
 as a fail-safe alternative.
-
 Users can control the value of $D$ used in the computations through
-command-line arguments to achieve greater performance as a trade-off for
+command-line arguments to achieve greater performance at the cost of
 accuracy.

### 2015-03-03 manuscript.tex
Mention v1.2 of Tesseroids.

@@
 with adaptive discretization
-in the open-source package \textit{Tesseroids}.
-It is freely available
-online through the software website\footnote{
-\href{http://tesseroids.leouieda.com}{http://tesseroids.leouieda.com}}
-under the BSD 3-clause open-source license.
+in version 1.2 of the open-source package \textit{Tesseroids}.
+It is freely available online
+through a persistent archive\footnote{
+\href{}{\todo[inline]{Zenodo DOI}}}
+or the software website\footnote{
+\href{http://tesseroids.leouieda.com}{http://tesseroids.leouieda.com}}.
+The software is distributed under the BSD 3-clause open-source license.
 An archived version of the source code
-is available as part of this article
-and online\footnote{
-\href{}{\todo[inline]{Replace with the Zenodo DOI for the Tesseroids
-source code.}}}.
+is also available as part of this article.
 

### 2015-03-06 manuscript.tex
Outline of conclusions

@@
 \citet{Grombein2013} formulated Cartesian kernels for the volume integrals
-that define the tesseroid gravitational potential, gravitational attraction,
+that define the tesseroid gravitational potential, gravitational acceleration,
 and Marussi tensor, respectively,
@@
 
+We have presented the open-source software package \textit{Tesseroids}.
+
+It consists of command-line programs written in C.
+
+The programs perform the forward modeling of gravitational fields in spherical
+coordinates.
+
+The fields are calculated from a mass model composed of spherical prisms, the
+so-called tesseroids.
+
+The volume integrals of the gravitational fields of a tesseroid are solved
+numerically using the Gauss-Legendre Quadrature (GLQ).
+
+The GLQ approximates the volume integral by a weighted sum of point mass
+effects.
+
+The error of the GLQ integration increases as the computation point gets closer
+to the tesseroid.
+
+The accuracy of the GLQ integration can be increased by using more point
+masses.
+
+Using more point masses also increases the computation time.
+
+The tesseroid can be divided into smaller tesseroids as an alternative to
+increasing the number of point masses in the GLQ.
+
+The effect of each tesseroid is computed using the original number of point
+masses.
+
+This effectively increases the total number of point masses used.
+
+We have implemented and improved upon an adaptive discretization algorithm to
+achieve an optimal division of tesseroids.
+
+Tesseroids are divided into more parts closer to the computation point.
+
+This produces an optimal distribution of point masses needed to produce
+accurate results.
+
+Our implementation of the adaptive discretization uses a ``stack'' data
+structure in place of the original recursive implementation.
+
+The stack-based algorithm is computationally faster and allows more control
+over errors when too many divisions are necessary (and the stack
+``overflows'').
+
+The criterion for dividing a tesseroid is based on the ratio the distance
+(measured between the computation point and the geometric center of the
+tesseroid) and the dimensions of the tesseroid.
+
+The tesseroids is divided in two along a particular direction if this ratio
+is larger than a parameter called the distance-size ratio ($D$).
+
+Thus, by the end of the adaptive discretization algorithm, all tesseroids will
+have dimensions smaller than $D$ times the distance to the computation point.
+
+The value of $D$ indirectly controls the accuracy of the integration as well as
+the computation time.
+
+We performed an error analysis to determine the optimal value of $D$ to achieve
+a target accuracy.
+
+We used a spherical shell as a reference to determine the computation error of
+our algorithm for different values of the distance-size ratio.
+
+The spherical shell is convenient because there are analytical solutions for
+its gravitational fields and it can be perfectly discretized into tesseroids.
+
+We performed numerical experiments under different conditions.
+
+We varied the size of the tesseroids used to discretize the shell, the height
+of the computation points, and the location of the computation points (near the
+pole or the equator).
+
+We established a maximum allowed error of 0.1\% of the spherical shell
+solutions.
+
+The plots of the computation error per distance-size ratio reveal that the
+values of $D$ required achieve 0.1\% error are
+1 for the potential, 1.5 for the gravitational acceleration,
+and 8 for the gravity gradients.
+
+Previous assumptions were that accurate results are
+guaranteed if the following condition is met:
+the distance to the tesseroid is larger than the distance between
+point masses (approximately the size of the tesseroid for a second-order GLQ).
+
+This same condition was applied indiscriminately to both the gravitational
+acceleration and the gravity gradients.
+
+This condition is equivalent to using a distance-size ratio of 1.5.
+
+Our results show that using $D=1$ produces an acceptable error of
+approximately 0.1\% for the gravitational acceleration.
+
+However, the same condition applied to the gravity gradients produces
+an error of the order of $10^2\%$.
+
+Therefore, the higher frequency gravity gradients require a finer
+discretization of the tesseroids.
+
+The behavior of the error for the gravity gradients varies with the computation
+height.
+
+The distance-size ratio required for 0.1\% error decreases with height.
+
+This can be attributed to the higher sensitivity of gravity gradient to height
+variations.
+
+The integration kernel for gravity gradients has high frequency components
+and thus requires more point masses for numerical integration.
+
+We chose the value $D=8$ as a conservative default.
+
+The values of the distance-size ratio determined in the error analysis were
+incorporated as defaults in the software \textit{Tesseroids}.
+
+The value of $D$ used can be controlled by a command-line argument.
+
+In situations that require many tesseroid divisions in the adaptive
+discretization, the stack used in the algorithm will overflow and further
+divisions become impossible.
+
+The current version of \textit{Tesseroids} (1.2) warns the user that
+this situation occurred and proceeds with the GLQ integration without
+division.
+
+Future improvements to the algorithm include a better way to handle situations
+that require many tesseroid divisions in the adaptive discretization.
+
+An alternative would be to replace the tesseroid by an equivalent
+right-rectangular prism and compute its effects instead.
+
+Also, the computation time increases drastically as the computation point
+gets closer to the tesseroid.
+
+This effect can be prohibitive for computing the gravity gradients at
+relatively low heights (e.g., for ground or airborne surveys).
+
+Further investigation of different criteria for dividing the tesseroids could
+yield better performance through a reduced number of divisions.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-03-07 manuscript.tex
Finish first try at the conclusions

--- 2015-03-03 manuscript.tex
@@
 
+We have presented the open-source software \textit{Tesseroids}.
+It consists of command-line programs,
+written in the C programming language,
+to perform the forward modeling of gravitational fields in spherical
+coordinates.
+The fields are calculated from a mass model composed of spherical prisms, the
+so-called tesseroids.
+The volume integrals of the gravitational fields of a tesseroid are solved
+numerically using the Gauss-Legendre Quadrature (GLQ).
+The GLQ approximates the volume integrals by weighted sums of point mass
+effects.
+The error of the GLQ integration increases as the computation point gets closer
+to the tesseroid.
+To counter this effect,
+the accuracy of the GLQ integration can be increased by using more point
+masses or by dividing each tesseroid into smaller ones.
+
+We have implemented and improved upon an adaptive discretization algorithm to
+achieve an optimal division of tesseroids.
+Tesseroids are divided into more parts closer to the computation point,
+where more point masses are needed.
+Our implementation of the adaptive discretization uses a ``stack'' data
+structure in place of the originally proposed recursive implementation.
+The stack-based algorithm is computationally faster and allows more control
+over errors when too many divisions are necessary.
+The adaptive discretization is controlled by
+a scalar called the distance-size ratio ($D$).
+The algorithm ensures that all tesseroids will
+have dimensions smaller than $D$ times the distance to the computation point.
+The value of $D$ indirectly controls the accuracy of the integration as well as
+the computation time.
+
+We performed an error analysis to determine
+the optimal value of $D$ required to achieve a target accuracy.
+We used a spherical shell as a reference to calculate the computation error of
+our algorithm for different values of $D$.
+Our results show that the values of $D$ required to achieve a maximum error
+of 0.1\% of the shell values are
+1 for the gravitational potential, 1.5 for the gravitational acceleration,
+and 8 for the gravity gradients.
+Previous assumptions were that accurate results are guaranteed if
+the distance to the tesseroid is larger than
+the distance between point masses.
+This same condition was previously applied indiscriminately
+to both the gravitational acceleration and the gravity gradients.
+That assumption is equivalent to using $D=1.5$ for all fields.
+Our results show that this is valid for the gravitational
+acceleration and results in a 0.1\% computation error.
+This is expected because the original study that determined the above condition
+was performed on the vertical component of gravitational acceleration.
+However, applying the same condition to the gravity gradients produces
+an error of the order of $10^2\%$.
+
+For the gravity gradients in particular,
+the distance-size ratio required for 0.1\% error decreases with height.
+This can be attributed to the higher sensitivity of gravity gradients
+to height variations because they decay with the distance cubed.
+This behavior causes the integration kernel to have more high frequency
+components, thus requiring more point masses for numerical integration.
+
+The values of the distance-size ratio determined above were
+incorporated as defaults in the software \textit{Tesseroids}.
+We chose the value $D=8$ for the gravity gradients as a conservative default.
+If the user desires, the value of $D$ used can be controlled by a command-line
+argument.
+
+In situations that require many tesseroid divisions,
+the stack used in the algorithm will overflow and further
+divisions become impossible.
+The current implementation warns the user that
+the overflow occurred and proceeds with the GLQ integration without division.
+Future improvements to the algorithm include a better way to handle such
+situations as they arise.
+An alternative would be to replace the tesseroid by an equivalent
+right rectangular prism and compute its effects instead.
+This would allow accurate computations at smaller distances.
+Furthermore,
+the computation time increases drastically as the computation point
+gets closer to the tesseroid.
+This effect can be prohibitive for computing the gravity gradients at
+relatively low heights (e.g., for terrain corrections of
+ground or airborne surveys).
+Further investigation of different criteria for dividing the tesseroids could
+yield better performance through a reduced number of divisions.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-03-07 manuscript.tex
Replace hrefs with urls

@@
 \address{
-    \noindent
 \footnotemark[1]Universidade do Estado do Rio de Janeiro,
@@
 in version 1.2 of the open-source package \textit{Tesseroids}.
-It is freely available online
-through a persistent archive\footnote{
-\href{}{\todo[inline]{Zenodo DOI}}}
-or the software website\footnote{
-\href{http://tesseroids.leouieda.com}{http://tesseroids.leouieda.com}}.
-The software is distributed under the BSD 3-clause open-source license.
+It is freely available online\footnote{
+\url{http://tesseroids.leouieda.com}}\footnote{
+\url{}\todo[inline]{Zenodo DOI}}
+under the BSD 3-clause open-source license.
 An archived version of the source code
@@
 were produced in IPython notebooks\footnote{
-\href{http://ipython.org}{http://ipython.org}}
+\url{http://ipython.org}}
 \citep{Perez2007}.
@@
 as well as instructions for installing and running the analysis,
-are distributed along with the source code archive
-that accompanies this article.
+are distributed along with
+the source code archive that accompanies this article.
 Alternatively,
-all accompanying material can be downloaded
-from an online source code repository\footnote{
-\href{https://github.com/pinga-lab/paper-tesseroids}{https://github.com/pinga-lab/paper-tesseroids}}
-or a persistent archive\footnote{
-\href{}{\todo[inline]{Replace with the Zenodo/figshare DOI for the repository.}}}.
+all accompanying material is also available online\footnote{
+\url{https://github.com/pinga-lab/paper-tesseroids}}\footnote{
+\url{}\todo[inline]{Replace with the Zenodo/figshare DOI for the repository.}}.
 

### 2015-03-07 manuscript.tex
Replaced textit with emph

@@
 \pdfoutput=1
-\documentclass[paper,twocolumn]{geophysics}
+\documentclass[paper,twocolumn,twoside]{geophysics}
 %\documentclass[manuscript]{geophysics}
@@
 We have implemented a modified version of the adaptive discretion of
-\citet{Li2011} into the open-source software package \textit{Tesseroids}.
+\citet{Li2011} into the open-source software package \emph{Tesseroids}.
 The software uses the Cartesian formula of \citet{Grombein2013} for improved
 performance and robustness.
-Previous versions of \textit{Tesseroids} have been used by, e.g.,
+Previous versions of the software have been used by, e.g.,
 \citet{Alvarez2012, Bouman2013, Bouman2013a, Mariani2013, Braitenberg2014,
 Braitenberg2011, Fullea2014}.
-This article describes the software \textit{Tesseroids} and the implementation
+This article describes the software design and the implementation
 of our modified adaptive discretization algorithm.
@@
 with adaptive discretization
-in version 1.2 of the open-source package \textit{Tesseroids}.
+in version 1.2 of the open-source package \emph{Tesseroids}.
 It is freely available online\footnote{
@@
 
-\textit{Tesseroids} consists of command-line programs
+\emph{Tesseroids} consists of command-line programs
 written in the C programming language.
@@
 We used the libraries
-\textit{pandas} by \citet{Mckinney2010}
+\emph{pandas} by \citet{Mckinney2010}
 to perform the data analysis,
-\textit{matplotlib} by \citet{Hunter2007} for graphs,
-and \textit{Mayavi} by \citet{Ramachandran2011} for 3D graphs.
+\emph{matplotlib} by \citet{Hunter2007} for graphs,
+and \emph{Mayavi} by \citet{Ramachandran2011} for 3D graphs.
 
@@
 We have implemented the values of the distance-size ratio producing $0.1\%$
-accuracy determined above as defaults for the software \textit{Tesseroids}.
+accuracy determined above as defaults for the software \emph{Tesseroids}.
 We chose the conservative value of $D=8$ for the gravity gradient components
@@
 
-We have presented the open-source software \textit{Tesseroids}.
+We have presented the open-source software \emph{Tesseroids}.
 It consists of command-line programs,
@@
 The values of the distance-size ratio determined above were
-incorporated as defaults in the software \textit{Tesseroids}.
+incorporated as defaults in the software \emph{Tesseroids}.
 We chose the value $D=8$ for the gravity gradients as a conservative default.

### 2015-03-07 manuscript.tex
Replace texttt by emph

@@
 
-The \texttt{src} folder of the source code archive
+The \emph{src} folder of the source code archive
 contains the C files that build the command-line programs
-(e.g., \texttt{tessgz.c}).
-The \texttt{src/lib} folder contains
+(e.g., \emph{tessgz.c}).
+The \emph{src/lib} folder contains
 the source files that implement the numerical computations.
@@
 
-The source file \texttt{src/lib/glq.c}
+The source file \emph{src/lib/glq.c}
 contains the code necessary to perform
@@
 The root-finder is implemented in functions
-\texttt{glq\_nodes} and \texttt{glq\_next\_root}.
+\emph{glq\_nodes} and \emph{glq\_next\_root}.
 The computed roots will be in the range $[-1, 1]$
@@
 (the physical boundaries of the tesseroid)
-using function \texttt{glq\_set\_limits} (see equation~\ref{eq:glq_scaling}).
+using function \emph{glq\_set\_limits} (see equation~\ref{eq:glq_scaling}).
 
 The GLQ weights (equation~\ref{eq:glq_weights})
-are computed by function \texttt{glq\_weights}.
+are computed by function \emph{glq\_weights}.
 Both the computed roots and weights are stored in a data structure
-(a C \texttt{struct}) called \texttt{GLQ}.
-Function \texttt{glq\_new}
+(a C \emph{struct}) called \emph{GLQ}.
+Function \emph{glq\_new}
 handles memory allocation,
 calculates the roots and weights,
-and returns the complete \texttt{GLQ} structure.
+and returns the complete \emph{GLQ} structure.
 
 The numerical integration of the tesseroid gravitational fields
-is performed by the functions in module \texttt{src/lib/grav\_tess.c}.
-Functions \texttt{tess\_pot}, \texttt{tess\_gx}, \texttt{tess\_gy}, and so on,
+is performed by the functions in module \emph{src/lib/grav\_tess.c}.
+Functions \emph{tess\_pot}, \emph{tess\_gx}, \emph{tess\_gy}, and so on,
 compute the gravitational fields of a single tesseroid
 on a single computation point.
-These functions require three \texttt{GLQ} structures,
+These functions require three \emph{GLQ} structures,
 each containing the roots and weights
@@
 on a single computation point.
-The algorithm is implemented in function \texttt{calc\_tess\_model\_adapt}
-of the file \texttt{src/lib/grav\_tess.c}.
+The algorithm is implemented in function \emph{calc\_tess\_model\_adapt}
+of the file \emph{src/lib/grav\_tess.c}.
 The stack of tesseroids is represented by
-the \texttt{stack} variable,
-an array of \texttt{TESSEROID} structures.
+the \emph{stack} variable,
+an array of \emph{TESSEROID} structures.
 We must define a maximum size for the stack to allocate memory for it.
@@
 (or sufficiently close to) the surface of the tesseroid.
-We use the integer \texttt{stktop}
+We use the integer \emph{stktop}
 to keep track of the index of
@@
 The initialization is done by copying the tesseroid into the stack
-and setting \texttt{stktop} to zero (the first element).
+and setting \emph{stktop} to zero (the first element).
 Once the stack is initialized, the steps of the algorithm are:
@@
     \item ``Pop'' the stack (i.e., take the last tesseroid from it).
-        This will cause \texttt{stktop} to be reduced by one.
+        This will cause \emph{stktop} to be reduced by one.
         This tesseroid is the one that will be evaluated in the following
@@
              the condition.
-        \item Check if the number of new elements plus \texttt{stktop}
+        \item Check if the number of new elements plus \emph{stktop}
              is smaller than the maximum stack size.
@@
     \item Repeat the above steps until the stack is empty
-        (\texttt{stktop} is equal to -1).
+        (\emph{stktop} is equal to -1).
 \end{enumerate}

### 2015-03-07 manuscript.tex
First try at the abstract

@@
 \begin{abstract}
+We present the open-source software \emph{Tesseroids},
+a set of command-line programs to perform the forward modeling
+of gravitational fields in spherical coordinates.
+The fields are calculated numerically using
+the Gauss-Legendre Quadrature (GLQ).
+We have improved upon an adaptive discretization algorithm
+to guarantee the accuracy of the GLQ integration.
+Our implementation of adaptive discretization
+uses a  ``stack'' based algorithm
+instead of recursion to achieve higher performance
+and more control over execution errors and corner cases.
+The algorithm is controlled by
+a scalar called the distance-size ratio ($D$)
+that determines the accuracy of the integration as well as
+the computation time.
+We determined optimal values of $D$
+for the gravitational potential, gravitational acceleration,
+and gravity gradient tensor
+by comparing the computed tesseroids effects
+with those of a homogeneous spherical shell.
+Contrary to previous assumptions,
+our results show that the potential and its first and second derivatives
+require different values of $D$ to achieve a target accuracy of 0.1\%.
+These values were incorporated as defaults in the software.
 \end{abstract}

### 2015-03-07 manuscript.tex
Acknowledge capes, cnpq and faperj

@@
 
-\todo[inline]{Here's a comment!}
+\todo[inline]{This is an example of how to leave a comment in a Latex
+document.}
 
@@
 
+The authors were supported in this research by
+a fellowship (VCFB) from
+Conselho Nacional de Desenvolvimento Cient\'ifico e Tecnol\'ogico (CNPq)
+and a scholarship (LU) from
+Coordena\c{c}\~ao de Aperfei\c{c}oamento de Pessoal de N\'ivel Superior
+(CAPES),
+Brazil.
+Additional support for the authors was provided by
+the Brazilian agencies CNPq (grant 471693/2011-1) and
+FAPERJ (grant E-26/103.175/2011).
+
 \bibliographystyle{seg}

### 2015-03-07 manuscript.tex
Add Valeria and Carla as authors
Better caption for the experiments table

@@
     Leonardo Uieda\footnotemark[1]\footnotemark[2],
-    et al.
+    Val\'eria C. F. Barbosa\footnotemark[2],
+    and
+    Carla Braitenberg\footnotemark[3]
 }
@@
 email: leouieda@gmail.com
+\\
 \footnotemark[2]Observat\'orio Nacional,
 Rio de Janeiro, Brazil.
+\\
+\footnotemark[3]Department of Mathematics and Geosciences, University of
+Trieste, Trieste, Italy
 }
@@
 \end{tabular}
-\caption{Parameters of the numerical experiments.}
+\caption{
+    Parameters of the numerical experiments to quantify the accuracy of
+    the numerical integration.
+    All grids had $10\times10$ regularly spaced computation points at a
+    constant height.
+    Tesseroids used to discretize the spherical shell had 1 km thickness and
+    the horizontal dimensions shown in the table.
+}
 \label{tbl:experiment}

### 2015-03-12 manuscript.tex
After some corrections and comments by @valcris

@@
 the volume integral by a weighted sum of the effect of point masses.
-An advantage of the GLQ approach is that the accuracy of integration can be
+An advantage of the GLQ approach is that it can be
 controlled by the number of point masses used.
+The larger the number of point masses, the better the accuracy of GLQ
+integration.
 A disadvantage is the increased computation time as the number of
 point masses increases.
-Thus, there is a trade-off between accuracy and computation time, a common
-theme in numerical methods.
+Thus, there is a trade-off between accuracy and computation time.
+This is a common theme in numerical methods.
 \citet{Wild-Pfeiffer2008} investigated the use of different mass elements,
@@
 
-Previous work by \citet{Ku1977} investigated the use of the Gauss-Legendre
-Quadrature in gravity forward modeling.
+Previous work by \citet{Ku1977} investigated the use of the GLQ in gravity forward modeling.
 \citet{Ku1977} numerically integrated the vertical component of the
@@
 between adjacent point masses should be less than the distance to the computation point.
-\citet{Asgharzadeh2007} proposed using this criteria for the GLQ integration of
+\citet{Asgharzadeh2007} used  this criteria for the GLQ integration of
 the gravity gradient tensor of tesseroids.
-To our knowledge, there has been no investigation if the empirical criteria of
-\citet{Ku1977} is valid for gravity gradient components or for tesseroids.
+To our knowledge, the analysis if this ad hoc criteria of
+\citet{Ku1977} works well for gravity gradient components or for tesseroids has never been done before.
+
 There has also been no attempt to quantify the error committed in the GLQ
@@
 because the number of points masses per tesseroid is fixed.
-The criterion used decide if a division is necessary is
-that the minimum distance to the computation point must be larger than
-the largest dimension of the tesseroid.
-This division is repeated recursively until all tesseroids obey the criteria.
+A tesseroid is divided if the minimum distance to the computation point
+is smaller than the largest dimension of the tesseroid.
+\todo[inline]{VAL:
+Leo, I have modified the above sentence because I have not understood the original meaning.
+}
+\todo[inline]{LEO: There was a mistake in that sentence.
+I've rewritten it. Is it better now?}
+This division is repeated recursively until all tesseroids obey the criterion.
 Then, GLQ integration is performed for each of the smaller tesseroids
@@
 increasing the number of points masses is that the
-total total distribution of point masses will be greater
+total distribution of point masses will be greater
 only close to the computation point.
@@
 of tesseroids using Cartesian integral kernels.
-These formula are faster to compute and don't have singularities at the poles
+These formulas are faster to compute and do not have singularities at the poles
 like their spherical counterparts.
@@
 at high latitudes.
-In the so called ``near-zone'' of the computation point they use a finner
-discretization with smaller tesseroids.
+In the so called ``near-zone'' of the computation point they use a finer
+discretization composed by smaller tesseroids.
 This is accomplished by dividing the tesseroids along their horizontal
@@
 Braitenberg2011, Fullea2014}.
+
+\todo[inline]{Valeria: Leo,  what does the above sentence mean? Does it mean that these mentioned authors have already used a former version of this software?}
+\todo[inline]{Leo: exactly. These are the papers I found that mentioned using Tesseroids.}
+
 This article describes the software design and the implementation
@@
     \sum\limits_{k=1}^{N^\lambda}
+    W_i^r W_j^\phi W_k^\lambda
     f(r_i, \phi_j, \lambda_k),
@@
 of the GLQ (black dots) and does not resemble the effect of a tesseroid.
-\citet{Ku1977} determined an empirical criterion that the distance between
+\citet{Ku1977} determined an ad hoc criterion that the distance between
 point masses (quadrature nodes) should be smaller than the minimum distance to
@@
         the result resembles that of
-        four point mases instead of a single tesseroid.
+        four point masses instead of a single tesseroid.
         This effect was shown by \citet{Ku1977}.
-        c) At $150\ km$ but with a higher GLQ order of 30,
-        the result is similar to that expected for a single mass source.
+        c) At $150\ km$ but with a higher GLQ order of 30.
+In (c) the horizontal locations of the point masses were not shown.
+Notice that the result shown in (c) is similar to that expected for a single mass source.
     }
@@
 \citet{Li2011} proposed an alternative method
-for decreasing the distance between point masses
-and achieve an accurate integration.
+for decreasing the distance between point masses on the quadrature nodes
+aiming at achieving an accurate integration.
 Instead of increasing the GLQ order,
 they keep it fixed to a given number
-and divide the tesseroid into smaller parts.
+and divide the tesseroid into smaller volumes.
 The sum of the effects of the smaller tesseroids
@@
 of the smaller size of the tesseroids.
-The criterion for dividing a tesseroid or not
-is that the distance to the computation point
-must be smaller than a constant times the size of the tesseroid.
-This is analogous to the criterion of \citet{Ku1977}
+The criterion for dividing a tesseroid is that the distance to the computation point
+should be smaller than a constant times the size of the tesseroid.
+This is analogous to the criterion proposed by \citet{Ku1977}
 because the size of the tesseroid serves as a proxy
@@
 This procedure is repeated recursively
-(i.e., to each of the smaller tesseroids and so on)
 until all tesseroids are within the acceptable
@@
 that the number of point masses is only increased
-in regions of the tesseroid that are
+in regions where the tesseroid are
 closer to the computation point.
-Alternatively,
+Notice that the alternative approach of
 simply increasing the order of the GLQ
 would increase the number of point masses
-evenly throughout the whole tesseroid.
+evenly throughout the whole tesseroids.
 
@@
 as well as model and computation grid generation.
+\todo[inline]{Valeria: Leo, I did not understand "as well as model and computation grid generation"}
+\todo[inline]{Leo: I mean that there are programs to generate the model files and the grid files. This can be removed without any problem.}
 All programs receive input through
@@
 calculate on it $g_z$ and $g_{zz}$ caused by
-the tesseroids in a file ``MODELFILE'',
-and save the results to a file called ``OUTPUT''
+the tesseroids in a file "MODELFILE'',
+and save the results to a file called "OUTPUT''
 is:
@@
 Our implementation of the adaptive discretization algorithm
-differs in a few ways from the original.
+differs in a few ways from one proposed by \citet{Li2011}.
 In \citet{Li2011},
@@
 Our definition of the dimensions of the tesseroid
-(the ``side lengths'' of \citet{Li2011})
+(the "side lengths'' of \citet{Li2011})
 along longitude, latitude, and radius, respectively, are
@@
 \begin{equation}
-    \frac{d}{L_i} < D,
+    \frac{d}{L_i} > D,
     \label{eq:condition}
 \end{equation}
+\todo[inline]{Leo: the condition was wrong. It had a $<$ when it should be a $>$. I fixed it.}
 
@@
 $D$ is a positive scalar
-hereafter referred to as the ``distance-size ratio''.
+hereafter referred to as the "distance-size ratio''.
 If the inequality holds for all three dimensions,
@@
         (b) 1, (c) 2, and (d) 6.
-        $d$ is the Cartesian distance between
-        the center of the top face of the tesseroid
-        and $P$.
         $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the tesseroid.
@@
         steps.
-    \item Compute the distance $d$ between
+    \item Compute the distance $d$ (equation~\ref{eq:distance}) between
         the geometric center of the tesseroid and
-        the computation point using equation~\ref{eq:distance}.
+        the computation point.
     \item Compute the dimensions of the tesseroid $L_\lambda$, $L_\phi$,
@@
         dimension of the tesseroid.
-    \item If none of the dimensions fail the above condition,
-        compute the gravitational effect of the tesseroid using the
+
+\todo[inline]{  Regarding the Item 5  - I think that it is confusing.
+
+In the item 5 is written " If none of the dimensions fail the above condition, ...."
+
+This  means that all the dimensions hold the inequality (21). Ok ?
+
+So, Item 5 could be written directly,
+
+5. If all the dimensions hold the inequality (21), hence the tesseroid is not divided and the gravitational effect of the tesseroid is computed by using the
+        Gauss-Legendre Quadrature.....}
+\todo[inline]{Leo: I agree completely. I have made the modifications below.}
+
+    \item If all dimensions hold the inequality \ref{eq:condition},
+        the tesseroid is not divided and its
+        gravitational effect is computed using the
         Gauss-Legendre Quadrature
@@
              the condition.
-        \item Check if the number of new elements plus \emph{stktop}
-             is smaller than the maximum stack size.
-             If it is not (i.e., the stack is full),
+
+\todo[inline]{   Regarding the Item 6b  - I am not sure, but I think that it might be wrong or written in an awkward style
+
+Item 6b is written - "Check if the number of new elements plus \emph{stktop} is smaller than the maximum stack size."
+Based on item 6c, I think that in 6b you are check if the stack is full, ok ?
+
+My suggestion is to merge items 6b and 6c, for example:
+
+If the stack is full, place the smaller tesseroids into the stack; otherwise warn the user of "stak overflow" and compute the effect of the larger tesseroid (step 5)
+
+}
+\todo[inline]{Leo: Yes, that sentence was a bit confusing.
+The idea is checking if the newly generated tesseroids
+fit in the stack.
+I just wanted to explain how to check for that
+(the number of new elements ...).
+I'll rephrase it.}
+\todo[inline]{Leo: By the way, the proper way of writing " in LaTeX is ``like this''.}
+
+        \item Check if there is room in the stack
+            for the new tesseroids
+            (i.e.,the number of new elements plus \emph{stktop}
+             is smaller than the maximum stack size).
+             If there isn't,
              warn the user of a ``stack overflow''
              and compute the effect of the larger tesseroid (step 5).
-        \item If the stack is not full, place the smaller tesseroids into
+             If there is room in the stack,
+             place the smaller tesseroids into
              the stack.
@@
 the maximum number of consecutive recursive calls
-(i.e., the ``recursion depth'').
+(i.e., the "recursion depth").
 This limit may vary with programming language,
@@
 
-\subsection{Reproducing the analysis and results}
-
-All figures and numerical experiments in this article
-were produced in IPython notebooks\footnote{
-\url{http://ipython.org}}
+\subsection{Code for figures and error analysis}
+
+The error analysis and all figures in this article
+were produced in IPython notebooks
 \citep{Perez2007}.
-These files combine the analysis and figure generation source code
-with text, equations, and the figures generated by the code.
-We used the libraries
-\emph{pandas} by \citet{Mckinney2010}
-to perform the data analysis,
-\emph{matplotlib} by \citet{Hunter2007} for graphs,
-and \emph{Mayavi} by \citet{Ramachandran2011} for 3D graphs.
-
-The notebook files,
-as well as instructions for installing and running the analysis,
-are distributed along with
+The notebook files combine source code in various programming languages,
+program execution,
+text, equations,
+and the figures generated by the code
+into a single document.
+We used the following Python language libraries
+to perform the error analysis and generate figures:
+\emph{pandas} by \citet{Mckinney2010},
+\emph{matplotlib} by \citet{Hunter2007} for 2D figures and maps,
+and \emph{Mayavi} by \citet{Ramachandran2011} for 3D figures.
+
+\todo[inline]{ VAL: The code was written in C., was not it? I think that there is a  version in Python. However, the readers do not known. }
+
+\todo[inline]{Leo: The code that we are discussing in this paper was written in C.
+The figures were generated in IPython notebooks in Python.
+The error analysis was done in IPython notebooks as well.
+However, it does NOT use the Python version of the tesseroid code.
+It runs the C programs and uses Python to calculate the differences
+The modeling that is written in Python is not included or discussed here.
+Readers do not need to know that it exists.}
+
+\todo[inline]{VAL: The above sentence is confusing:
+1) "These files combine the analysis and figure.... ",
+What are the files? Probably, they are the IPython notebooks. Ok?
+Why are you saying "the analysis"
+In the last paragraph, you said "running the analysis".
+What is the analysis? Is the test to evaluate the accuracy?
+I think that you want to say
+"The Ipython nootebook files combine source codes in python, code executions, text, equations, dnd the figures generated by the code into a single document"
+You need keeping in mind that you know about Ipython nootebook. However, the readers cannot know.}
+
+\todo[inline]{LEO: Yes, that part is a bit confusing now that I read it again.
+By analysis, I mean the error analysis.
+That is done in \href{http://nbviewer.ipython.org/github/pinga-lab/paper-tesseroids/blob/master/notebooks/tesseroid_vs_spherical_shell.ipynb}{this IPython notebook}.
+This file will be  available  with the article.
+There is a link to this online version in \url{https://github.com/pinga-lab/paper-tesseroids}
+which is mentioned in the text.
+There are two other notebooks that make the figures for the methods and the final results.
+There are links to those two in the above link  as well.
+I'll rewrite the paragraph to make it more clear.}
+
+The IPython notebooks
+and the data generated for the error analysis,
+as well as instructions for installing the software
+and running the programs,
+are also included in
 the source code archive that accompanies this article.
 Alternatively,
-all accompanying material is also available online\footnote{
+all accompanying material is available
+in online repositories\footnote{
 \url{https://github.com/pinga-lab/paper-tesseroids}}\footnote{
@@
 
+\todo[inline]{ VAL: The above paragraph is great !}
+
 The reference against which we compare the computed tesseroid fields
@@
 Instead, we calculate the effect of the tesseroid model on a regular grid
-of $10 \times 10$ points.
+of $10 \times 10$ points at different geographic locations
+(Table~\ref{tbl:experiment}).
+\todo[inline]{ VAL: A grid of 10 X  10  point covers a are of ??? X ???? km}
+\todo[inline]{LEO: The grid location varied in each experiment of the numerical analysis. Table 1 shows the locations of the grids.
+This is discussed below.
+I'll add this information to the text.}
 Fortunately, the symmetry of the shell allows us to consider the computation
@@
 We compute the differences between the effects of the shell and the tesseroid
-model on the grid and consider only the largest error.
-By placing the grid on top of a particular tesseroid, we increase the chances
-of capturing the true largest integration error.
+model on the grid.
+However, we will consider only the largest error in our analysis.
+\todo[inline]{ VAL: I rewrite the above sentence by:
+"We compute the differences between the effects of the shell and the tesseroid model on the grid; however, we consider  the largest error only.  }
+\todo[inline]{LEO: Rewritten. Is it better now?}
+We placed the grid on top of a particular tesseroid
+to increase the chances of capturing the true largest integration error.
+\todo[inline]{ VAL:I rewrite the above sentence as:
+"We decided for the nonconservative procedure of placing the grid on top of a particular tesseroid to increase the chances
+of capturing the true largest integration error"}
+\todo[inline]{LEO: Changed. I don't think this is really
+``nonconservative'' so I left it out.}
+
+\todo[inline]{ I've stopped here - VALERIA}
 

### 2015-03-12 manuscript.tex
Added back proper quotation marks

@@
 controlled by the number of point masses used.
-The larger the number of point masses, the better the accuracy of GLQ
-integration.
+The larger the number of point masses, the better the accuracy of GLQ integration.
 A disadvantage is the increased computation time as the number of
@@
 the gravity gradient tensor of tesseroids.
-To our knowledge, the analysis if this ad hoc criteria of
-\citet{Ku1977} works well for gravity gradient components or for tesseroids has never been done before.
-
+To our knowledge, an analysis of how well this ad hoc criteria of
+\citet{Ku1977} works for gravity gradient components or for tesseroids
+has never been done before.
 There has also been no attempt to quantify the error committed in the GLQ
 integration when applying the criteria of \citet{Ku1977}.
+\todo[inline]{LEO: Valeria, I've made some modifications to your correction.}
 
@@
         In (c) the horizontal locations of the point masses were not shown.
-Notice that the result shown in (c) is similar to that expected for a single mass source.
+        Notice that the results shown in (c) are similar to that expected
+        for a single mass source.
     }
@@
 that the number of point masses is only increased
-in regions where the tesseroid are
+in parts of the tesseroid that are
 closer to the computation point.
@@
 would increase the number of point masses
-evenly throughout the whole tesseroids.
+evenly throughout the whole tesseroid.
+\todo[inline]{LEO: Valeria, this correction you made is wrong.
+It is not ``in regions where the tesseroids are closer''.
+It is actually ``in parts of the tesseroid that are closer''.
+This is all happening inside a single tesseroid.
+I have corrected the text above.}
 
@@
 the gravitational fields of tesseroids and
-rectangular prisms (in both Cartesian and spherical coordinates),
-as well as model and computation grid generation.
+rectangular prisms (in both Cartesian and spherical coordinates).
 \todo[inline]{Valeria: Leo, I did not understand "as well as model and computation grid generation"}
@@
 All programs receive input through
-command-line arguments and the standard input channel (STDIN)
-and output the results through the standard output channel (STDOUT).
+command-line arguments and the standard input channel (``STDIN'')
+and output the results through the standard output channel (``STDOUT'').
 For example,
@@
 calculate on it $g_z$ and $g_{zz}$ caused by
-the tesseroids in a file "MODELFILE'',
-and save the results to a file called "OUTPUT''
+the tesseroids in a file ``MODELFILE'',
+and save the results to a file called ``OUTPUT''
 is:
@@
 Our definition of the dimensions of the tesseroid
-(the "side lengths'' of \citet{Li2011})
+(the ``side lengths'' of \citet{Li2011})
 along longitude, latitude, and radius, respectively, are
@@
 $D$ is a positive scalar
-hereafter referred to as the "distance-size ratio''.
+hereafter referred to as the ``distance-size ratio''.
 If the inequality holds for all three dimensions,
@@
 the maximum number of consecutive recursive calls
-(i.e., the "recursion depth").
+(i.e., the ``recursion depth'').
 This limit may vary with programming language,
@@
 of $10 \times 10$ points at different geographic locations
-(Table~\ref{tbl:experiment}).
+(see Table~\ref{tbl:experiment}).
 \todo[inline]{ VAL: A grid of 10 X  10  point covers a are of ??? X ???? km}

### 2015-03-13 manuscript.tex
More comments and corrections from @valcris

@@
 Rio de Janeiro, Brazil.
+email: valcris@on.br
 \\
@@
 However, we will consider only the largest error in our analysis.
+
 \todo[inline]{ VAL: I rewrite the above sentence by:
@@
 
-\todo[inline]{ I've stopped here - VALERIA}
-
 We calculate the errors for values of the distance-size ratio
 $D$ varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
-Furthermore, we repeated the analysis in four different numerical experiments,
+Furthermore, we repeated the error analysis in four different numerical experiments,
 each with computation grids at different locations and
@@
 tesseroid model.
+
+ \todo[inline]{ Val: Regarding the visual aspect of Figure 4.
+I think that  the dashed lines are yielding a visual pollution.
+However, they are helping us to identify the values in the axes.
+Regarding the visual aspect of Figure 4.
+A possibility would be keep the dashed lines, but with  other color (e.g., light gray  dashed lines) }
 
@@
 $0.1\%$.
-For a computation height of 260 km, the error curve for $G_{zz}$
+For a computation height of 260 km, the error curve for $g_{zz}$
 intercepts the error threshold line at $D=2.5$.
@@
 computation heights.
-Notice that the distance-size ratio needed to achieve $0.1\%$ accuracy
+Notice that the distance-size ratio required to achieve $0.1\%$ accuracy
 decreases as the computation height increases.
@@
 
+\todo[inline]{ VAL: I think that we discussed that similar results that are achieved in gzz (Figure 5) also are I achieved in gxx, gyy.... I think that this information could be written.}
+
 \begin{table*}
 \centering
-\begin{tabular}{cccc}
+\begin{tabular}{lccc}
     \hline
@@
     \hline
-    Experiment 1  & 89N--90N/0E--1E & 2 km & $1^\circ \times 1^\circ$ \\
-    Experiment 2  & 0N--1N/0E--1E & 2 km & $1^\circ \times 1^\circ$ \\
-    Experiment 3  & 89N--90N/0E--1E & 260 km & $1^\circ \times 1^\circ$ \\
-    Experiment 4  & 60N--90N/0E--30E & 2 km & $30^\circ \times 30^\circ$ \\
+    Experiment 1 (pole)            & 89N--90N/0E--1E  & 2 km   & $1^\circ \times 1^\circ$ \\
+    Experiment 2 (equator)         & 0N--1N/0E--1E    & 2 km   & $1^\circ \times 1^\circ$ \\
+    Experiment 3 (260 km)          & 89N--90N/0E--1E  & 260 km & $1^\circ \times 1^\circ$ \\
+    Experiment 4 ($30^\circ$ size) & 60N--90N/0E--30E & 2 km   & $30^\circ \times 30^\circ$ \\
     \hline
@@
 
+\todo[inline]{ I am here. I did not read the captions of the Figures 4 and 5 and the caption of the Table 1}
+
 We have presented the open-source software \emph{Tesseroids}.
@@
 and 8 for the gravity gradients.
-Previous assumptions were that accurate results are guaranteed if
+Previous assumptions in the literature
+were that accurate results are guaranteed if
 the distance to the tesseroid is larger than
 the distance between point masses.
-This same condition was previously applied indiscriminately
+This condition was previously applied indiscriminately
 to both the gravitational acceleration and the gravity gradients.
@@
 Additional support for the authors was provided by
-the Brazilian agencies CNPq (grant 471693/2011-1) and
-FAPERJ (grant E-26/103.175/2011).
+the Brazilian agency FAPERJ (grant E-26/103.175/2011).
 

### 2015-03-13 manuscript.tex
Removed comments from the paper

@@
 
-\todo[inline]{This is an example of how to leave a comment in a Latex
-document.}
-
 Satellite missions dedicated to measuring the Earths gravity field
@@
 controlled by the number of point masses used.
-The larger the number of point masses, the better the accuracy of GLQ integration.
+The larger the number of point masses,
+the better the accuracy of GLQ integration.
 A disadvantage is the increased computation time as the number of
@@
 
-Previous work by \citet{Ku1977} investigated the use of the GLQ in gravity forward modeling.
+Previous work by \citet{Ku1977} investigated the use of the GLQ
+in gravity forward modeling.
 \citet{Ku1977} numerically integrated the vertical component of the
@@
 Based on this, \citet{Ku1977} proposed an empirical criteria that the distance
-between adjacent point masses should be less than the distance to the computation point.
+between adjacent point masses should be less than
+the distance to the computation point.
 \citet{Asgharzadeh2007} used  this criteria for the GLQ integration of
@@
 integration when applying the criteria of \citet{Ku1977}.
-\todo[inline]{LEO: Valeria, I've made some modifications to your correction.}
 
@@
 is smaller than the largest dimension of the tesseroid.
-\todo[inline]{VAL:
-Leo, I have modified the above sentence because I have not understood the original meaning.
-}
-\todo[inline]{LEO: There was a mistake in that sentence.
-I've rewritten it. Is it better now?}
 This division is repeated recursively until all tesseroids obey the criterion.
@@
 Braitenberg2011, Fullea2014}.
-
-\todo[inline]{Valeria: Leo,  what does the above sentence mean? Does it mean that these mentioned authors have already used a former version of this software?}
-\todo[inline]{Leo: exactly. These are the papers I found that mentioned using Tesseroids.}
 
@@
 evenly throughout the whole tesseroid.
-\todo[inline]{LEO: Valeria, this correction you made is wrong.
-It is not ``in regions where the tesseroids are closer''.
-It is actually ``in parts of the tesseroid that are closer''.
-This is all happening inside a single tesseroid.
-I have corrected the text above.}
 
@@
 rectangular prisms (in both Cartesian and spherical coordinates).
-\todo[inline]{Valeria: Leo, I did not understand "as well as model and computation grid generation"}
-\todo[inline]{Leo: I mean that there are programs to generate the model files and the grid files. This can be removed without any problem.}
 All programs receive input through
@@
 \end{equation}
-\todo[inline]{Leo: the condition was wrong. It had a $<$ when it should be a $>$. I fixed it.}
 
@@
         dimension of the tesseroid.
-
-\todo[inline]{  Regarding the Item 5  - I think that it is confusing.
-
-In the item 5 is written " If none of the dimensions fail the above condition, ...."
-
-This  means that all the dimensions hold the inequality (21). Ok ?
-
-So, Item 5 could be written directly,
-
-5. If all the dimensions hold the inequality (21), hence the tesseroid is not divided and the gravitational effect of the tesseroid is computed by using the
-        Gauss-Legendre Quadrature.....}
-\todo[inline]{Leo: I agree completely. I have made the modifications below.}
-
     \item If all dimensions hold the inequality \ref{eq:condition},
@@
              the condition.
-
-\todo[inline]{   Regarding the Item 6b  - I am not sure, but I think that it might be wrong or written in an awkward style
-
-Item 6b is written - "Check if the number of new elements plus \emph{stktop} is smaller than the maximum stack size."
-Based on item 6c, I think that in 6b you are check if the stack is full, ok ?
-
-My suggestion is to merge items 6b and 6c, for example:
-
-If the stack is full, place the smaller tesseroids into the stack; otherwise warn the user of "stak overflow" and compute the effect of the larger tesseroid (step 5)
-
-}
-\todo[inline]{Leo: Yes, that sentence was a bit confusing.
-The idea is checking if the newly generated tesseroids
-fit in the stack.
-I just wanted to explain how to check for that
-(the number of new elements ...).
-I'll rephrase it.}
-\todo[inline]{Leo: By the way, the proper way of writing " in LaTeX is ``like this''.}
-
         \item Check if there is room in the stack
@@
 
-\todo[inline]{ VAL: The code was written in C., was not it? I think that there is a  version in Python. However, the readers do not known. }
-
-\todo[inline]{Leo: The code that we are discussing in this paper was written in C.
-The figures were generated in IPython notebooks in Python.
-The error analysis was done in IPython notebooks as well.
-However, it does NOT use the Python version of the tesseroid code.
-It runs the C programs and uses Python to calculate the differences
-The modeling that is written in Python is not included or discussed here.
-Readers do not need to know that it exists.}
-
-\todo[inline]{VAL: The above sentence is confusing:
-1) "These files combine the analysis and figure.... ",
-What are the files? Probably, they are the IPython notebooks. Ok?
-Why are you saying "the analysis"
-In the last paragraph, you said "running the analysis".
-What is the analysis? Is the test to evaluate the accuracy?
-I think that you want to say
-"The Ipython nootebook files combine source codes in python, code executions, text, equations, dnd the figures generated by the code into a single document"
-You need keeping in mind that you know about Ipython nootebook. However, the readers cannot know.}
-
-\todo[inline]{LEO: Yes, that part is a bit confusing now that I read it again.
-By analysis, I mean the error analysis.
-That is done in \href{http://nbviewer.ipython.org/github/pinga-lab/paper-tesseroids/blob/master/notebooks/tesseroid_vs_spherical_shell.ipynb}{this IPython notebook}.
-This file will be  available  with the article.
-There is a link to this online version in \url{https://github.com/pinga-lab/paper-tesseroids}
-which is mentioned in the text.
-There are two other notebooks that make the figures for the methods and the final results.
-There are links to those two in the above link  as well.
-I'll rewrite the paragraph to make it more clear.}
-
 The IPython notebooks
@@
 
-\todo[inline]{ VAL: The above paragraph is great !}
-
 The reference against which we compare the computed tesseroid fields
@@
 (see Table~\ref{tbl:experiment}).
-\todo[inline]{ VAL: A grid of 10 X  10  point covers a are of ??? X ???? km}
-\todo[inline]{LEO: The grid location varied in each experiment of the numerical analysis. Table 1 shows the locations of the grids.
-This is discussed below.
-I'll add this information to the text.}
 Fortunately, the symmetry of the shell allows us to consider the computation
@@
 
-\todo[inline]{ VAL: I rewrite the above sentence by:
-"We compute the differences between the effects of the shell and the tesseroid model on the grid; however, we consider  the largest error only.  }
-\todo[inline]{LEO: Rewritten. Is it better now?}
 We placed the grid on top of a particular tesseroid
 to increase the chances of capturing the true largest integration error.
-\todo[inline]{ VAL:I rewrite the above sentence as:
-"We decided for the nonconservative procedure of placing the grid on top of a particular tesseroid to increase the chances
-of capturing the true largest integration error"}
-\todo[inline]{LEO: Changed. I don't think this is really
-``nonconservative'' so I left it out.}
-
 We calculate the errors for values of the distance-size ratio
@@
 tesseroid model.
-
- \todo[inline]{ Val: Regarding the visual aspect of Figure 4.
-I think that  the dashed lines are yielding a visual pollution.
-However, they are helping us to identify the values in the axes.
-Regarding the visual aspect of Figure 4.
-A possibility would be keep the dashed lines, but with  other color (e.g., light gray  dashed lines) }
 
@@
 accuracy.
-
-\todo[inline]{ VAL: I think that we discussed that similar results that are achieved in gzz (Figure 5) also are I achieved in gxx, gyy.... I think that this information could be written.}
 
@@
     \includegraphics[width=0.4\textwidth]{figs/distance-size-curves}
-    %\missingfigure{}
     \caption{
@@
     \includegraphics[width=0.4\textwidth]{figs/gzz-with-height}
-    %\missingfigure{}
     \caption{
@@
 \section{Conclusions}
-
-\todo[inline]{ I am here. I did not read the captions of the Figures 4 and 5 and the caption of the Table 1}
 

### 2015-03-19 manuscript.tex
Condition for division should be >=
Included the pre-reserved DOIs
Say that D=0 means no divisions
Removed  hyperref because it was causing problems

@@
 \usepackage{amsmath}
-\usepackage{hyperref}
 \newcommand{\mitbf}[1]{\mathbf{{#1}}}
@@
 
-\hypersetup{
-    colorlinks=true,
-    citecolor=blue,
-    linkcolor=blue,
-    urlcolor=blue,
-    pdfauthor={Uieda et al.},
-    pdftitle={Tesseroids: forward modeling gravitational fields in spherical
-              coordinates}
-}
 \title{
@@
 It is freely available online\footnote{
-\url{http://tesseroids.leouieda.com}}\footnote{
-\url{}\todo[inline]{Zenodo DOI}}
+http://tesseroids.leouieda.com}\footnote{
+http://dx.doi.org/10.5281/zenodo.16033}
 under the BSD 3-clause open-source license.
@@
 \begin{equation}
-    \frac{d}{L_i} > D,
+    \frac{d}{L_i} \geq D,
     \label{eq:condition}
@@
 in online repositories\footnote{
-\url{https://github.com/pinga-lab/paper-tesseroids}}\footnote{
-\url{}\todo[inline]{Replace with the Zenodo/figshare DOI for the repository.}}.
+https://github.com/pinga-lab/paper-tesseroids}\footnote{
+http://dx.doi.org/10.5281/zenodo.16032}.
 
@@
         of $0.1\%$.
+        A value of $D=0$ means that no divisions are made.
     }

### 2015-03-20 manuscript.tex
Comments by Carla

@@
 
-Satellite missions dedicated to measuring the Earths gravity field
+Satellite missions dedicated to measuring the Earth's gravity field
 (like CHAMP, GRACE, and GOCE)
@@
 adjacent point masses.
-Based on this, \citet{Ku1977} proposed an empirical criteria that the distance
+Based on this, \citet{Ku1977} proposed an empirical criterion that the distance
+% * <berg@units.it> 2015-03-19T13:46:11.161Z:
+%
+%  an criterion or proposed empirical criteria
+%
 between adjacent point masses should be less than
@@
 because the number of points masses per tesseroid is fixed.
+% * <berg@units.it> 2015-03-19T13:48:33.551Z:
+%
+%  sentence is repeated later.
+%
 A tesseroid is divided if the minimum distance to the computation point
@@
 The Cartesian formula are numerically integrated using a Taylor series
+% * <berg@units.it> 2015-03-19T13:49:40.239Z:
+%
+%  singular or plural?
+%
 expansion as per \citet{Heck2007}.
@@
 integration kernel \citep{Hildebrand1987},
+
 \begin{equation}
@@
 the command to generate a regular grid with $NLON \times NLAT$ points,
-calculate on it $g_z$ and $g_{zz}$ caused by
+ $g_z$ and $g_{zz}$ caused by
 the tesseroids in a file ``MODELFILE'',
@@
 These points are roots of Legendre polynomials.
-Precomputed values are available low order polynomials,
+Precomputed values are available for low order polynomials,
 typically up to order five.
@@
 \begin{equation}
-    L_r = r_2 - r_2.
+    L_r = r_2 - r_1.
     \label{eq:sizer}
@@
 Then, the stack is initialized with the input tesseroid.
+% * <berg@units.it> 2015-03-19T21:04:01.768Z:
+%
+%  the description of the operation of stack is not clear- is it the input tesseroid array?
+%
 The initialization is done by copying the tesseroid into the stack
@@
 decreases as the computation height increases.
+% * <berg@units.it> 2015-03-19T21:07:33.445Z:
+%
+%  what is the height dependence due to?
+Should the distance not include height already?
+Is the height a second control parameter, next to the distance ratio?
+What is that due to?
+Is it only for gzz or also for the other gradient components?
+What about the gx?
+%
 For example, computation at 260 km height requires $D=2.5$ whereas at 10 km
@@
 Additional support for the authors was provided by
-the Brazilian agency FAPERJ (grant E-26/103.175/2011).
+the Brazilian agency FAPERJ (grant E-26/103.175/2011) and by the GOCE-Italy project (ASI).
 

### 2015-03-20 manuscript.tex
My replies to carla

@@
 \\
-\footnotemark[3]Department of Mathematics and Geosciences, University of
-Trieste, Trieste, Italy
+\footnotemark[3]Department of Mathematics and Geosciences,
+University of Trieste, Trieste, Italy
 }
@@
 %
+% ^ <leouieda@gmail.com> 2015-03-20T17:41:12.218Z:
+%
+%  It was an criterion
+%
 between adjacent point masses should be less than
@@
 the number of point masses per tesseroid.
-This division effectively increases the total number of point masses
-because the number of points masses per tesseroid is fixed.
 % * <berg@units.it> 2015-03-19T13:48:33.551Z:
@@
 %  sentence is repeated later.
+%
+% ^ <leouieda@gmail.com> 2015-03-20T17:40:17.120Z:
+%
+%  I removed it.
 %
@@
 like their spherical counterparts.
-The Cartesian formula are numerically integrated using a Taylor series
+The Cartesian formulae are numerically integrated using a Taylor series
 % * <berg@units.it> 2015-03-19T13:49:40.239Z:
@@
 %  singular or plural?
+%
+% ^ <leouieda@gmail.com> 2015-03-20T17:41:54.790Z:
+%
+%  It should be the plural (formulae)
 %
@@
 the command to generate a regular grid with $NLON \times NLAT$ points,
- $g_z$ and $g_{zz}$ caused by
+ calculate $g_z$ and $g_{zz}$ caused by
 the tesseroids in a file ``MODELFILE'',
@@
 
-We will describe the stack-based algorithm
-for computing the effect of a single tesseroid
+The discretization algorithm is implemented in
+function \emph{calc\_tess\_model\_adapt}
+of the file \emph{src/lib/grav\_tess.c}.
+This function calculates the effect of a single tesseroid
 on a single computation point.
-The algorithm is implemented in function \emph{calc\_tess\_model\_adapt}
-of the file \emph{src/lib/grav\_tess.c}.
 The stack of tesseroids is represented by
@@
 
+Below, we describe the algorithm to calculate
+the effect of a single tesseroid from the input model
+on a single computation point.
 The algorithm starts by creating an empty stack of tesseroids.
-Then, the stack is initialized with the input tesseroid.
+Then, the stack is initialized with the single input tesseroid.
 % * <berg@units.it> 2015-03-19T21:04:01.768Z:
@@
 %  the description of the operation of stack is not clear- is it the input tesseroid array?
+%
+% ^ <leouieda@gmail.com> 2015-03-20T18:02:14.018Z:
+%
+%  I modified the description above and added more text below as well. Is it clearer now?
 %
@@
 and setting \emph{stktop} to zero (the first element).
+It is important to note that the stack is not the input tesseroid model.
+Instead, it is a buffer used to temporarily store
+each stage of the discretization algorithm.
+
 Once the stack is initialized, the steps of the algorithm are:
@@
 
+The algorithm above is repeated
+for every tesseroid of the input model
+and the results are summed.
+This will yield the gravitational effect
+of the input tesseroid model on a single point.
+Thus, the computations must be repeated
+for every computation point.
+The whole algorithm can be summarized
+in the following pseudo-code.
+
+\begin{verbatim}
+Initialize the output array with zeros.
+for tesseroid in model:
+  for point in grid:
+    Initialize the stack with tesseroid.
+    stktop = 0
+    while stktop >= 0:
+      Perform steps 1-6 of the algorithm.
+    Sum the calculated value to the output.
+\end{verbatim}
+
 This stack-based implementation
@@
 What about the gx?
+%
+% ^ <leouieda@gmail.com> 2015-03-20T18:23:53.901Z:
+%
+%  The height is included in the distance and is not a control variable.
+This happens for all tensor components but not to the potential, gx, gy, and gz.
+I think this is because the decay factor for the tensor components is 1/r^3 whereas the discreatization algorithm compares D to r/size.
+So as the  computation point becomes closer.
+the field increases more rapidly than the algorithm increases the  amount of discretization.
+That is why you need higher D (more discretization) when you are closer to the tesseroid.
 %
@@
 the distance-size ratio required for 0.1\% error decreases with height.
-This can be attributed to the higher sensitivity of gravity gradients
-to height variations because they decay with the distance cubed.
-This behavior causes the integration kernel to have more high frequency
-components, thus requiring more point masses for numerical integration.
+We believe this is because the decay factor for
+the gravity gradient components is $d^{-3}$,
+whereas the discretization algorithm uses $d/L_i$.
+As the computation point becomes closer to the tesseroid,
+the field increases more rapidly than
+the algorithm increases the amount of discretization.
+Hence, a higher value of D (i.e., more discretization)
+is required.
+
+% * <leouieda@gmail.com> 2015-03-20T18:43:18.571Z:
+%
+%  I have modified the above paragraph to try to explain why the ratio for gzz varies with height.
+%
 
@@
 Additional support for the authors was provided by
-the Brazilian agency FAPERJ (grant E-26/103.175/2011) and by the GOCE-Italy project (ASI).
+the Brazilian agency FAPERJ (grant E-26/103.175/2011)
+and by the GOCE-Italy project (ASI).
 

### 2015-03-30 manuscript.tex
White edges in split figure
Comments by Valeria and Carla

@@
 shape at the poles.
-The Gauss-Legendre Quadrature (GLQ) integration consists of approximating
+The Gauss-Legendre Quadrature (GLQ) integration consists in approximating
 the volume integral by a weighted sum of the effect of point masses.
@@
 the ratio between distance to the computation point and the distance between
+% * <valeria.valcris@gmail.com> 2015-03-23T17:28:28.769Z:
+%
+%  Leo, the sentence "Based on this, \citet{Ku1977} proposed an empirical criterion that the distance the distance to the computation point."
+%
+% ^ <valeria.valcris@gmail.com> 2015-03-23T17:57:28.416Z.
 adjacent point masses.
-Based on this, \citet{Ku1977} proposed an empirical criterion that the distance
-% * <berg@units.it> 2015-03-19T13:46:11.161Z:
-%
-%  an criterion or proposed empirical criteria
-%
-% ^ <leouieda@gmail.com> 2015-03-20T17:41:12.218Z:
-%
-%  It was an criterion
-%
-between adjacent point masses should be less than
+% * <valeria.valcris@gmail.com> 2015-03-23T17:29:37.800Z:
+%
+%  is wrong. The criterion disappeared.
+%
+% ^ <valeria.valcris@gmail.com> 2015-03-23T17:57:27.222Z.
+Based on this, \citet{Ku1977} proposed an empirical criterion that
+the distance between point masses should be greater than
 the distance to the computation point.
-\citet{Asgharzadeh2007} used  this criteria for the GLQ integration of
+\citet{Asgharzadeh2007} used  this criterion for the GLQ integration of
 the gravity gradient tensor of tesseroids.
@@
 the number of point masses per tesseroid.
-% * <berg@units.it> 2015-03-19T13:48:33.551Z:
-%
-%  sentence is repeated later.
-%
-% ^ <leouieda@gmail.com> 2015-03-20T17:40:17.120Z:
-%
-%  I removed it.
-%
 A tesseroid is divided if the minimum distance to the computation point
@@
 The Cartesian formulae are numerically integrated using a Taylor series
-% * <berg@units.it> 2015-03-19T13:49:40.239Z:
-%
-%  singular or plural?
-%
-% ^ <leouieda@gmail.com> 2015-03-20T17:41:54.790Z:
-%
-%  It should be the plural (formulae)
-%
 expansion as per \citet{Heck2007}.
@@
 in the computations.
-These results allow us calibrate the adaptive discretization algorithm
+These results allow us to calibrate the adaptive discretization algorithm
 separately for the gravitational potential, gravitational acceleration,
@@
 $\rho$ is the density,
-$G = 6.673\times10^{-11}\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
+$G = 6.674\times10^{-11}\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
+% * <berg@units.it> 2015-03-22T10:56:54.365Z:
+%
+%  Leo, on the link below the digits are 6.67384, so if you truncate it should be 6.674 not 6.673 as on the wikipedia article on G.
+%
+% * <berg@units.it> 2015-03-22T10:54:00.463Z:
+%
+%  http://physics.nist.gov/cgi-bin/cuu/Value?bg|search_for=universal_in!
+%
+% ^ <leouieda@gmail.com> 2015-03-22T21:20:28.891Z:
+%
+%  That's correct, thanks. It must have been a typing error.
+%
 and
@@
 using the Gauss-Legendre Quadrature (GLQ).
-The GLQ consists of approximating the integral by a weighted sum of the
+The GLQ consists in approximating the integral by a weighted sum of the
 integration kernel \citep{Hildebrand1987},
@@
 Our implementation of the adaptive discretization algorithm
-differs in a few ways from one proposed by \citet{Li2011}.
+differs in a few ways from the one proposed by \citet{Li2011}.
 In \citet{Li2011},
@@
 \end{enumerate}
-
+% * <valeria.valcris@gmail.com> 2015-03-23T19:53:50.437Z:
+%
+%  Leo, item 6b I suggest changing " warn the user of a ``stack overflow''              and compute the effect of the larger tesseroid (step 5)." by
+%
 The algorithm above is repeated
+% * <valeria.valcris@gmail.com> 2015-03-23T21:09:58.768Z:
+%
+%  by "warn the user of a ``stack overflow''  and compute the effect of the tesseroid computed on the step 5."
+%
 for every tesseroid of the input model
@@
 of $10 \times 10$ points at different geographic locations
+% * <berg@units.it> 2015-03-22T12:57:56.006Z:
+%
+%  and takethe average over the hundert points. Leo: is that correct? Should we add it to be more clear?
+%
+% ^ <leouieda@gmail.com> 2015-03-22T20:37:47.094Z:
+%
+%  No, I don't take the average.
+I take the largest error.
+I didn't mention that above because I only talk about the error further on ("Figure~\ref{fig:dist-size-curves} shows the maximum difference between the shell and tesseroid fields").
+%
 (see Table~\ref{tbl:experiment}).
@@
 $D$ varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
+% * <berg@units.it> 2015-03-22T13:00:05.803Z:
+%
+%  D=0 means we are sitting in the center of the tesseroid, more than no division. Maybe you could be more precise on this, as D=0 will not happen, as you are at 2km height?
+%
+% ^ <leouieda@gmail.com> 2015-03-22T20:45:02.247Z:
+%
+%  I'm not moving the computation grid during this test.
+What I do is I vary the value of D used in the adaptive discretization.
+That's what is shown in Fig 4.
+Each curve corresponds to a fixed grid.
+Changing D just means that the adaptive discretization is performed differently because of the inequality in eq 17.
+Notice that if D=0, any combination of distance/size is larger than D.
+So there won't be any divisions.
+%
 Furthermore, we repeated the error analysis in four different numerical experiments,
@@
 Only results for the gravitational potential, $g_z$, and $g_{zz}$ are shown.
+% * <berg@units.it> 2015-03-22T13:02:41.484Z:
+%
+%  Not sure I understand what you did for the 30° tesseroid.
+Errors are nearly the same.
+It means that at D=1 the tesseroid has been divided into many small tesseroids so as to have the correct distance-size ratio: Already for D=.5 the calculation is made on small tesseroids that make up the big 30° tesseroid.
+So actually the different curves represent very similar  tesseroids in the near field, only in the far field are the tesseroids left with the large size.
+We see that the result is nearly the same, with a great reduction in computer time in the case of the 30° tesseroid as for distant tesseroids no division is necessary.
+Leo: maybe we could add this comment, if my interpretation is correct.
+%
+% ^ <leouieda@gmail.com> 2015-03-22T21:19:24.057Z:
+%
+%  Carla, your interpretation is correct.
+In practice there is no advantage to using large tesseroids with this algorithm because I don't take into account a finer resolution model in the near zone (like a proper near-far zone separation).
+The difference is that I'm dividing a corse model into small but equivalent tesseroids.
+For a near-far zone separation I'd average out the fine model in the far zone.
+%
 The results for the other diagonal components of the gravity gradient tensor
@@
 That is why you need higher D (more discretization) when you are closer to the tesseroid.
+%
+% ^ <berg@units.it> 2015-03-22T13:32:27.725Z:
+%
+%  hello Leo: are you sure D<2 is what he is calculating at 260km height?
+if it is a 1° tesseroid and you are at 260km height, your distance is greater than the biggest sidelength of the tesseroid, and maybe that is why the error starts to decrease for D>2.5, because for smalle D your condition is satisfied by itself .
+%
+% ^ <leouieda@gmail.com> 2015-03-22T21:14:00.081Z:
+%
+%  That is exactly it. For 1º tesseroids at 260km, there are no divisions for D < 2 because the condition is already satisfied. That's why the error curve is flat.
 %

### 2015-03-30 manuscript.tex
Cleaned up comments from the source

--- 2015-03-20 manuscript.tex
@@
 adjacent point masses.
-Based on this, \citet{Ku1977} proposed an empirical criterion that the distance
-% * <berg@units.it> 2015-03-19T13:46:11.161Z:
-%
-%  an criterion or proposed empirical criteria
-%
-% ^ <leouieda@gmail.com> 2015-03-20T17:41:12.218Z:
-%
-%  It was an criterion
-%
-between adjacent point masses should be less than
+Based on this, \citet{Ku1977} proposed an empirical criterion that
+the distance between point masses should be greater than
 the distance to the computation point.
-\citet{Asgharzadeh2007} used  this criteria for the GLQ integration of
+\citet{Asgharzadeh2007} used  this criterion for the GLQ integration of
 the gravity gradient tensor of tesseroids.
@@
 $\rho$ is the density,
-$G = 6.673\times10^{-11}\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
+$G = 6.674\times10^{-11}\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
 and
@@
 Then, the stack is initialized with the single input tesseroid.
-% * <berg@units.it> 2015-03-19T21:04:01.768Z:
-%
-%  the description of the operation of stack is not clear- is it the input tesseroid array?
-%
-% ^ <leouieda@gmail.com> 2015-03-20T18:02:14.018Z:
-%
-%  I modified the description above and added more text below as well. Is it clearer now?
-%
 The initialization is done by copying the tesseroid into the stack
@@
              warn the user of a ``stack overflow''
-             and compute the effect of the larger tesseroid (step 5).
+             and compute the effect of the tesseroid, as in step 5.
              If there is room in the stack,
@@
 all accompanying material is available
-in online repositories\footnote{
-https://github.com/pinga-lab/paper-tesseroids}\footnote{
-http://dx.doi.org/10.5281/zenodo.16032}.
+in an online repository\footnote{
+https://github.com/pinga-lab/paper-tesseroids}.
 
@@
 decreases as the computation height increases.
-% * <berg@units.it> 2015-03-19T21:07:33.445Z:
-%
-%  what is the height dependence due to?
-Should the distance not include height already?
-Is the height a second control parameter, next to the distance ratio?
-What is that due to?
-Is it only for gzz or also for the other gradient components?
-What about the gx?
-%
-% ^ <leouieda@gmail.com> 2015-03-20T18:23:53.901Z:
-%
-%  The height is included in the distance and is not a control variable.
-This happens for all tensor components but not to the potential, gx, gy, and gz.
-I think this is because the decay factor for the tensor components is 1/r^3 whereas the discreatization algorithm compares D to r/size.
-So as the  computation point becomes closer.
-the field increases more rapidly than the algorithm increases the  amount of discretization.
-That is why you need higher D (more discretization) when you are closer to the tesseroid.
-%
 For example, computation at 260 km height requires $D=2.5$ whereas at 10 km
@@
 
-% * <leouieda@gmail.com> 2015-03-20T18:43:18.571Z:
-%
-%  I have modified the above paragraph to try to explain why the ratio for gzz varies with height.
-%
-
 The values of the distance-size ratio determined above were

### 2015-11-15 manuscript.tex
Make PDF in manuscript mode for submission
Remove figure size commands
Add link to software.seg.org in the paper
Clarify what LIFO stands for
Add D values to abstract + extra about the code

@@
 \pdfoutput=1
-\documentclass[paper,twocolumn,twoside]{geophysics}
+%\documentclass[paper,twocolumn,twoside]{geophysics}
 %\documentclass[manuscript]{geophysics}
-%\documentclass[manuscript,endfloat]{geophysics}
-%\documentclass[twocolumn]{article}
+\documentclass[manuscript,endfloat]{geophysics}
 \usepackage{natbib}
@@
 \righthead{Tesseroids: modeling in spherical coordinates}
-\ms{Submission} % manuscript number
-
-\address{
+\ms{GEO-2015-0204} % manuscript number
+
+\address{Peer-reviewed code related to this article can be found at
+http://software.seg.org/2016/000X.
+\\
 \footnotemark[1]Universidade do Estado do Rio de Janeiro,
@@
 of gravitational fields in spherical coordinates.
-The fields are calculated numerically using
+The software is implemented in the C programming language and uses tesseroids
+(spherical prisms) for the discretization of the subsurface mass distribution.
+The gravitational fields of tesseroids are calculated numerically using
 the Gauss-Legendre Quadrature (GLQ).
@@
 uses a ``stack'' based algorithm
-instead of recursion to achieve higher performance
-and more control over execution errors and corner cases.
+instead of recursion to achieve
+more control over execution errors and corner cases.
 The algorithm is controlled by
-a scalar called the distance-size ratio ($D$)
+a scalar value called the distance-size ratio ($D$)
 that determines the accuracy of the integration as well as
@@
 with those of a homogeneous spherical shell.
+The values required for a maximum relative error of 0.1\% of the shell effects
+are $D = 1$ for the gravitational potential, $D = 1.5$ for the gravitational
+acceleration, and $D = 8$ for the gravity gradients.
 Contrary to previous assumptions,
 our results show that the potential and its first and second derivatives
-require different values of $D$ to achieve a target accuracy of 0.1\%.
+require different values of $D$ to achieve the same accuracy.
 These values were incorporated as defaults in the software.
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figs/tesseroid}
+    \includegraphics{figs/tesseroid}
     \caption{
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figs/tesseroid-split}
+    \includegraphics{figs/tesseroid-split}
     \caption{
@@
 (commonly referred to as ``popping'' the stack).
-Stacks are also known as ``Last-in-first-out'', or LIFO, data structures.
+Because of these restrictions,
+stacks are also known as ``Last-In-First-Out'' (LIFO) data structures.
 
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figs/distance-size-curves}
+    \includegraphics{figs/distance-size-curves}
     \caption{
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figs/gzz-with-height}
+    \includegraphics{figs/gzz-with-height}
     \caption{

### 2015-11-15 manuscript.tex
Add missing punctuation in equations
Explain Kronecker's delta

@@
         I_{\alpha\beta}
-        \ \kappa\ dr^\prime d\phi^\prime d\lambda^\prime.
+        \ \kappa\ dr^\prime d\phi^\prime d\lambda^\prime ,
     \label{eq:tesstensor}
@@
         \frac{\delta_{\alpha\beta}}{\ell^3}
-    \right)
+    \right) ,
     \label{eq:tesstensorkernel}
@@
 $G = 6.674\times10^{-11}\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
+$\delta_{\alpha\beta}$ is Kronecker's delta
+($\delta_{\alpha\beta} = 1$ if $\alpha = \beta$
+and $\delta_{\alpha\beta} = 0$ if $\alpha \neq \beta$),
 and
@@
         r^2 + r_t^2 - 2 r r_t \cos\psi_t
-        \right]^{\frac{1}{2}}.
+        \right]^{\frac{1}{2}} ,
     \label{eq:distance}
@@
     \cos\psi_t =
-        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t)
+        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t) .
 \end{equation}

### 2015-11-15 manuscript.tex
Add 2 other refs for shell effect
Talk about analytical solutions in intro
Before saying that need to solve numerically, say that there are
analytical in special cases of a spherical cap

@@
 instead of rectangular prisms.
-However, the integral formula for the gravitational effects of a tesseroid
+An analytical solution exists when the computation point is along
+the polar axis and the tesseroid is extended into a spherical cap
+\citep{Lafehr1991, Mikuska2006, Grombein2013}.
+For more general cases,
+the integral formula for the gravitational effects of a tesseroid
 must be solved numerically.
@@
 The shell has analytical solutions along the polar axis
-\citep{Grombein2013}
+\citep{Lafehr1991, Mikuska2006, Grombein2013}
 and can be perfectly discretized into tesseroids.

### 2015-11-15 manuscript.tex
stack faster than recursion is not a result
Make it clear that the stack vs recursion speed claim is not based on
experimental results but as a rule of thumb

@@
 namely:
-(1) It will be faster because it bypasses the overhead of function calls.
-(2) It gives the developer more control over the recursion step.
+(1) It gives the developer more control over the recursion step.
+(2) In general, it is faster because it bypasses the overhead of function
+calls.
 In recursive implementations,
@@
 structure in place of the originally proposed recursive implementation.
-The stack-based algorithm is computationally faster and allows more control
+As a rule of thumb in procedural languages (like C),
+stack-base implementations are computationally faster
+than the equivalent code using function recursion.
+Furthermore, the stack-based algorithm allows more control
 over errors when too many divisions are necessary.

### 2016-05-02 manuscript.tex
Add correct link to SEG software archive
Fixes #8
Cite figure from figshare
Fixes #5
Separate the table to another file
Fixes #11

@@
 \address{Peer-reviewed code related to this article can be found at
-http://software.seg.org/2016/000X.
+http://software.seg.org/2016/0004.
 \\
@@
         and $\ell$ is the Cartesian distance between $P$ and $Q$.
+        After \citet{uieda2015}.
     }
@@
 
-\begin{table*}
-\centering
-\begin{tabular}{lccc}
-    \hline
-    ~             & Grid location & Grid height & Tesseroid size \\
-    \hline
-    Experiment 1 (pole)            & 89N--90N/0E--1E  & 2 km   & $1^\circ \times 1^\circ$ \\
-    Experiment 2 (equator)         & 0N--1N/0E--1E    & 2 km   & $1^\circ \times 1^\circ$ \\
-    Experiment 3 (260 km)          & 89N--90N/0E--1E  & 260 km & $1^\circ \times 1^\circ$ \\
-    Experiment 4 ($30^\circ$ size) & 60N--90N/0E--30E & 2 km   & $30^\circ \times 30^\circ$ \\
-    \hline
-\end{tabular}
-\caption{
-    Parameters of the numerical experiments to quantify the accuracy of
-    the numerical integration.
-    All grids had $10\times10$ regularly spaced computation points at a
-    constant height.
-    Tesseroids used to discretize the spherical shell had 1 km thickness and
-    the horizontal dimensions shown in the table.
-}
-\label{tbl:experiment}
-\end{table*}
+\input{table1.tex}
 

### 2016-05-02 table1.tex
Separate the table to another file
Fixes #11

\begin{table*}
\centering
\begin{tabular}{lccc}
    \hline
    ~             & Grid location & Grid height & Tesseroid size \\
    \hline
    Experiment 1 (pole)            & 89N--90N/0E--1E  & 2 km   & $1^\circ \times 1^\circ$ \\
    Experiment 2 (equator)         & 0N--1N/0E--1E    & 2 km   & $1^\circ \times 1^\circ$ \\
    Experiment 3 (260 km)          & 89N--90N/0E--1E  & 260 km & $1^\circ \times 1^\circ$ \\
    Experiment 4 ($30^\circ$ size) & 60N--90N/0E--30E & 2 km   & $30^\circ \times 30^\circ$ \\
    \hline
\end{tabular}
\caption{
    Parameters of the numerical experiments to quantify the accuracy of
    the numerical integration.
    All grids had $10\times10$ regularly spaced computation points at a
    constant height.
    Tesseroids used to discretize the spherical shell had 1 km thickness and
    the horizontal dimensions shown in the table.
}
\label{tbl:experiment}
\end{table*}

### 2016-05-02 manuscript-final.tex
Add tex file with bib inside

--- 2016-05-02 manuscript.tex
@@
-\pdfoutput=1
-%\documentclass[paper,twocolumn,twoside]{geophysics}
-%\documentclass[manuscript]{geophysics}
 \documentclass[manuscript,endfloat]{geophysics}
@@
 \newcommand{\mitbf}[1]{\mathbf{{#1}}}
-
-% To leave comments
-\usepackage[colorinlistoftodos]{todonotes}
 
@@
 
+We are indebted to the developers and maintainers of the open-source
+software without which this work would not have been possible.
+The authors thank associate editor Joe Dellinger, reviewer Roman Pasteka, and
+four anonymous reviewers for their hard work and helpful comments.
 The authors were supported in this research by
@@
 
-\bibliographystyle{seg}
-\bibliography{references.bib}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{thebibliography}{}
+\itemsep0pt
+
+\bibitem[{\'A}lvarez et~al., 2012]{Alvarez2012}
+{\'A}lvarez, O., M. Gimenez, C. Braitenberg, and A. Folguera,  2012, {GOCE}
+  satellite derived gravity and gravity gradient corrected for topographic
+  effect in the south central andes region: {GOCE} derivatives in the south
+  central andes: Geophysical Journal International, {\bf 190}, 941--959.
+
+\bibitem[Asgharzadeh et~al., 2007]{Asgharzadeh2007}
+Asgharzadeh, M.~F., R.~R.~B. von Frese, H.~R. Kim, T.~E. Leftwich, and J.~W.
+  Kim,  2007, Spherical prism gravity effects by {Gauss}-legendre quadrature
+  integration: Geophysical Journal International, {\bf 169}, 1--11.
+
+\bibitem[Barrera-Figueroa et~al., 2006]{Barrera-Figueroa2006}
+Barrera-Figueroa, V., J. Sosa-Pedroza, and J. L{\'o}pez-Bonilla,  2006,
+  Multiple root finder algorithm for legendre and {Chebyshev} polynomials via
+  {Newton\textquoteright}s method: Annales Mathematicae et Informaticae,
+  3{\textendash}13.
+
+\bibitem[Bouman et~al., 2013a]{Bouman2013}
+Bouman, J., J. Ebbing, and M. Fuchs,  2013a, Reference frame transformation of
+  satellite gravity gradients and topographic mass reduction: Journal of
+  Geophysical Research: Solid Earth, {\bf 118}, 759--774.
+
+\bibitem[Bouman et~al., 2013b]{Bouman2013a}
+Bouman, J., J. Ebbing, S. Meekes, R. Abdul~Fattah, M. Fuchs, S. Gradmann, R.
+  Haagmans, V. Lieb, M. Schmidt, D. Dettmering, and W. Bosch,  2013b, {GOCE}
+  gravity gradient data for lithospheric modeling: International Journal of
+  Applied Earth Observation and Geoinformation, {\bf 35}, 16--30.
+
+\bibitem[Braitenberg, 2015]{Braitenberg2015}
+Braitenberg, C.,  2015, Exploration of tectonic structures with {GOCE} in
+  {Africa} and across-continents: International Journal of Applied Earth
+  Observation and Geoinformation, {\bf 35}, 88--95.
+
+\bibitem[Braitenberg et~al., 2013]{Braitenberg2014}
+Braitenberg, C., P. Mariani, and A. De~Min,  2013, The {Europe}an alps and
+  nearby orogenic belts sensed by {GOCE}: Bollettino di Geofisica Teorica e
+  Applicata, {\bf 54}, 321--334.
+
+\bibitem[Braitenberg et~al., 2011]{Braitenberg2011}
+Braitenberg, C., P. Mariani, J. Ebbing, and M. Sprlak,  2011, The enigmatic
+  {Chad} lineament revisited with global gravity and gravity-gradient fields:
+  Geological Society, London, Special Publications, {\bf 357}, 329--341.
+
+\bibitem[Fullea et~al., 2014]{Fullea2014}
+Fullea, J., J. Rodr{\'i}guez-Gonz{\'a}lez, M. Charco, Z. Martinec, A. Negredo,
+  and A. Villase{\~n}or,  2014, Perturbing effects of sub-lithospheric mass
+  anomalies in {GOCE} gravity gradient and other gravity data modelling:
+  Application to the atlantic-mediterranean transition zone: International
+  Journal of Applied Earth Observation and Geoinformation, {\bf 35}, 54--69.
+
+\bibitem[Grombein et~al., 2013]{Grombein2013}
+Grombein, T., K. Seitz, and B. Heck,  2013, Optimized formulas for the
+  gravitational field of a tesseroid: Journal of Geodesy, {\bf 87}, 645--660.
+
+\bibitem[Heck and Seitz, 2007]{Heck2007}
+Heck, B., and K. Seitz,  2007, A comparison of the tesseroid, prism and
+  point-mass approaches for mass reductions in gravity field modelling: Journal
+  of Geodesy, {\bf 81}, 121--136.
+
+\bibitem[Hildebrand, 1987]{Hildebrand1987}
+Hildebrand, F.~B.,  1987, Introduction to numerical analysis: Dover
+  Publications.
+
+\bibitem[Hunter, 2007]{Hunter2007}
+Hunter, J.~D.,  2007, Matplotlib: {A} 2d graphics environment: Computing in
+  Science \& Engineering, {\bf 9}, 90--95.
+
+\bibitem[Ku, 1977]{Ku1977}
+Ku, C.~C.,  1977, A direct computation of gravity and magnetic anomalies caused
+  by 2-and 3-dimensional bodies of arbitrary shape and arbitrary magnetic
+  polarization by equivalent-point method and a simplified cubic spline:
+  Geophysics, {\bf 42}, 610{\textendash}622.
+
+\bibitem[{LaFehr}, 1991]{Lafehr1991}
+{LaFehr}, T.,  1991, An exact solution for the gravity curvature ({Bullard}
+  {B}) correction: {GEOPHYSICS}, {\bf 56}, 1179--1184.
+
+\bibitem[Li et~al., 2011]{Li2011}
+Li, Z., T. Hao, Y. Xu, and Y. Xu,  2011, An efficient and adaptive approach for
+  modeling gravity effects in spherical coordinates: Journal of Applied
+  Geophysics, {\bf 73}, 221--231.
+
+\bibitem[Mariani et~al., 2013]{Mariani2013}
+Mariani, P., C. Braitenberg, and N. Ussami,  2013, Explaining the thick crust
+  in paran{\'a} basin, {Brazil}, with satellite {GOCE} gravity observations:
+  Journal of South American Earth Sciences, {\bf 45}, 209--223.
+
+\bibitem[McKinney, 2010]{Mckinney2010}
+McKinney, W.,  2010, Data structures for statistical computing in python:
+  Proceedings of the 9th Python in Science Conference, 51 -- 56.
+
+\bibitem[Miku{\v s}ka et~al., 2006]{Mikuska2006}
+Miku{\v s}ka, J., R. {Pa}{\v s}teka, and I. {Maru}{\v s}iak,  2006, Estimation
+  of distant relief effect in gravimetry: {GEOPHYSICS}, {\bf 71}, J59--J69.
+
+\bibitem[P\'erez and Granger, 2007]{Perez2007}
+P\'erez, F., and B.~E. Granger,  2007, {IP}ython: a system for interactive
+  scientific computing: Computing in Science and Engineering, {\bf 9}, 21--29.
+
+\bibitem[Ramachandran and Varoquaux, 2011]{Ramachandran2011}
+Ramachandran, P., and G. Varoquaux,  2011, Mayavi: 3d visualization of
+  scientific data: Computing in Science \& Engineering, {\bf 13}, 40--51.
+
+\bibitem[Reguzzoni et~al., 2013]{Reguzzoni2013}
+Reguzzoni, M., D. Sampietro, and F. Sanso,  2013, Global {Moho} from the
+  combination of the {CRUST}2.0 model and {GOCE} data: Geophysical Journal
+  International, {\bf 195}, 222--237.
+
+\bibitem[{Uieda}, 2015]{uieda2015}
+{Uieda}, L.,  2015, A tesserioid (spherical prism) in a geocentric coordinate
+  system with a local-{{North}}-oriented coordinate system: figshare, available
+  from: http://dx.doi.org/10.6084/m9.figshare.1495525.
+\newblock (Accessed May 2016).
+
+\bibitem[Wild-Pfeiffer, 2008]{Wild-Pfeiffer2008}
+Wild-Pfeiffer, F.,  2008, A comparison of different mass elements for use in
+  gravity gradiometry: Journal of Geodesy, {\bf 82}, 637--653.
+
+\end{thebibliography}
+
 \end{document}

### 2016-05-03 manuscript.tex
Saying thanks
Fixes #6
Rename figures with numbers
To not confuse the bad submission system.

Fixes #12

@@
     \centering
-    \includegraphics{figs/tesseroid}
+    \includegraphics{figs/figure1}
     \caption{
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figs/vary-height-and-order}
+    \includegraphics[width=0.4\textwidth]{figs/figure2}
     \caption{
@@
     \centering
-    \includegraphics{figs/tesseroid-split}
+    \includegraphics{figs/figure3}
     \caption{
@@
     \centering
-    \includegraphics{figs/distance-size-curves}
+    \includegraphics{figs/figure4}
     \caption{
@@
     \centering
-    \includegraphics{figs/gzz-with-height}
+    \includegraphics{figs/figure5}
     \caption{

### 2016-05-18 manuscript-final.tex
Rename figures with numbers
To not confuse the bad submission system.

Fixes #12
Had to remove figures and tables for submission

@@
-\documentclass[manuscript,endfloat]{geophysics}
+% Final version submitted to the journal system. Had to remove all
+% cross-references to figures and tables and include only the captions.
+% Bibliography is included in this file, instead of the .bib file.
+
+\documentclass[manuscript]{geophysics}
 \usepackage{natbib}
@@
 A common approach to achieve this is
-to discretize the Earth into tesseroids (Figure~\ref{fig:tesseroid})
+to discretize the Earth into tesseroids (Figure~1)
 instead of rectangular prisms.
@@
 
-\begin{figure}
-    \centering
-    \includegraphics{figs/tesseroid}
-    \caption{
-        View of a tesseroid,
-        the integration point $Q$ inside the tesseroid,
-        a geocentric coordinate system $(X, Y, Z)$,
-        the computation $P$ and it's local coordinate system $(x, y, z)$.
-        $r$, $\phi$, $\lambda$ are
-        the radius, latitude, and longitude, respectively, of point $P$,
-        and $\ell$ is the Cartesian distance between $P$ and $Q$.
-        After \citet{uieda2015}.
-    }
-    \label{fig:tesseroid}
-\end{figure}
-
 A tesseroid is a mass element defined in geocentric spherical
 coordinates
-(Figure~\ref{fig:tesseroid}).
+(Figure~1).
 It is bounded by two meridians, two parallels, and two concentric circles.
@@
 are determined with respect to the local North-oriented coordinate system at
-P (x, y, z in Figure~\ref{fig:tesseroid}).
+P (x, y, z in Figure~1).
 \citet{Grombein2013} formulated Cartesian kernels for the volume integrals
@@
 the distance to the computation point and the distance between adjacent nodes.
-Figure~\ref{fig:glqerrorsample}
+Figure~2
 illustrates this effect on the $g_{xy}$ gravity gradient component.
@@
 with $100\times100$ points.
-Figure~\ref{fig:glqerrorsample}a shows the $g_{xy}$ component
+Figure~2a shows the $g_{xy}$ component
 calculated at 400 km height using
 GLQ with order two ($2 \times 2 \times 2 = 8$ point masses).
-Figure~\ref{fig:glqerrorsample}b shows $g_{xy}$ computed with order two
+Figure~2b shows $g_{xy}$ computed with order two
 GLQ as well but at 150 km height.
@@
 $N$ in all three directions.
-Figure~\ref{fig:glqerrorsample}c shows the $g_{xy}$ component calculated at
+Figure~2c shows the $g_{xy}$ component calculated at
 150km height but with a GLQ order of 30
@@
 the expect results for a single tesseroid \citep{Asgharzadeh2007}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\textwidth]{figs/vary-height-and-order}
-    \caption{
-        Example of the effect of varying
-        the computation height
-        and the number of point masses in the Gauss-Legendre Quadrature.
-        Black circles represent the horizontal location of the point masses.
-        a) $g_{xy}$ calculated at $400\ km$ height using GLQ order 2
-        ($2 \times 2 \times 2 = 8$ point masses).
-        b) At $150\ km$ height and GLQ order 2,
-        the result resembles that of
-        four point masses instead of a single tesseroid.
-        This effect was shown by \citet{Ku1977}.
-        c) At $150\ km$ but with a higher GLQ order of 30.
-        In (c) the horizontal locations of the point masses were not shown.
-        Notice that the results shown in (c) are similar to that expected
-        for a single mass source.
-    }
-    \label{fig:glqerrorsample}
-\end{figure}
 
@@
 along longitude, latitude, and radius, respectively, are
-(Figure~\ref{fig:division}a)
+(Figure~3a)
 
@@
 $L_\lambda$ and $L_\phi$ are arc-distances measured along the top surface of
-the tesseroid (Figure~\ref{fig:division}a).
+the tesseroid (Figure~3a).
 Specifically, $L_\lambda$ is measured long the middle latitude of the
@@
 
-Figure~\ref{fig:division} shows examples of
+Figure~3 shows examples of
 the resulting tesseroid models after adaptive discretization.
-Figure~\ref{fig:division}a shows
+Figure~3a shows
 the initial tesseroid and computation point P.
-Figures~\ref{fig:division}b-d are
+Figures~3b-d are
 the result of adaptive discretization using
@@
 4, 38, and 936.
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/tesseroid-split}
-    \caption{
-        Adaptive discretization
-        of the tesseroid shown in (a)
-        for a computation point P
-        using the distance-size ratio $D$ equal to
-        (b) 1, (c) 2, and (d) 6.
-        $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the tesseroid.
-        Note that increasing $D$
-        results in a fine division of the tesseroid
-        close the computation point
-        and a coarser division further away.
-    }
-    \label{fig:division}
-\end{figure}
 
@@
 The specific value chosen for $D$ determines how many divisions will be made
-(Figure~\ref{fig:division}).
+(Figure~3).
 Thus, $D$ indirectly controls both the accuracy of the integration
@@
 
-Figure~\ref{fig:glqerrorsample} shows that the largest errors are spread over
+Figure~2 shows that the largest errors are spread over
 on top of the tesseroid.
@@
 of $10 \times 10$ points at different geographic locations
-(see Table~\ref{tbl:experiment}).
+(see Table~1).
 Fortunately, the symmetry of the shell allows us to consider the computation
@@
 different tesseroid model sizes.
-Table~\ref{tbl:experiment} describes the different numerical
+Table~1 describes the different numerical
 experiments and the corresponding parameters of the computation grid and
@@
 
-Figure~\ref{fig:dist-size-curves} shows
+Figure~4 shows
 the maximum difference between the shell and tesseroid fields
@@
 We established a maximum tolerated error of $0.1\%$, represented by the
-horizontal solid lines in Figure~\ref{fig:dist-size-curves}.
+horizontal solid lines in Figure~4.
 Only results for the gravitational potential, $g_z$, and $g_{zz}$ are shown.
@@
 heights 2, 10, 50, 150, and 260 km.
-Figure~\ref{fig:gzz-with-height} shows the results for $g_{zz}$ at varying
+Figure~5 shows the results for $g_{zz}$ at varying
 computation heights.
@@
 accuracy.
-
-\input{table1.tex}
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/distance-size-curves}
-    \caption{
-        The maximum difference between the computed tesseroid and shell effects
-        as a function of the distance-size ratio $D$
-        for (a) the gravitational potential, (b) $g_z$, and (c) $g_{zz}$.
-        The difference is given as a percentage of the shell effect.
-        Curves correspond to the different tesseroid models and computation
-        grids shown in Table~\ref{tbl:experiment}.
-        The horizontal solid black line marks the established error threshold
-        of $0.1\%$.
-        A value of $D=0$ means that no divisions are made.
-    }
-    \label{fig:dist-size-curves}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/gzz-with-height}
-    \caption{
-        Difference between the computed $g_{zz}$ for the spherical shell and
-        the tesseroid model at different heights. Curves show the maximum
-        difference as a percentage of the shell value.
-        The horizontal solid black line marks the established error threshold
-        of $0.1\%$.
-        A value of $D=0$ means that no divisions are made.
-    }
-    \label{fig:gzz-with-height}
-\end{figure}
 
@@
 
+\newpage
+\section{Figure captions}
+
+\begin{enumerate}
+    \item
+    View of a tesseroid, the integration point $Q$ inside the tesseroid, a
+    geocentric coordinate system $(X, Y, Z)$, the computation $P$ and it's
+    local coordinate system $(x, y, z)$.  $r$, $\phi$, $\lambda$ are the
+    radius, latitude, and longitude, respectively, of point $P$, and $\ell$ is
+    the Cartesian distance between $P$ and $Q$.  After \citet{uieda2015}.
+
+    \item
+    Example of the effect of varying the computation height and the number of
+    point masses in the Gauss-Legendre Quadrature.  Black circles represent the
+    horizontal location of the point masses.  a) $g_{xy}$ calculated at $400\
+    km$ height using GLQ order 2 ($2 \times 2 \times 2 = 8$ point masses).  b)
+    At $150\ km$ height and GLQ order 2, the result resembles that of four
+    point masses instead of a single tesseroid.  This effect was shown by
+    \citet{Ku1977}.  c) At $150\ km$ but with a higher GLQ order of 30.  In (c)
+    the horizontal locations of the point masses were not shown.  Notice that
+    the results shown in (c) are similar to that expected for a single mass
+    source.
+
+    \item
+    Adaptive discretization of the tesseroid shown in (a) for a computation
+    point P using the distance-size ratio $D$ equal to (b) 1, (c) 2, and
+    (d) 6.  $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the
+    tesseroid.  Note that increasing $D$ results in a fine division of the
+    tesseroid close the computation point and a coarser division further
+    away.
+
+    \item
+    The maximum difference between the computed tesseroid and shell effects
+    as a function of the distance-size ratio $D$ for (a) the gravitational
+    potential, (b) $g_z$, and (c) $g_{zz}$.  The difference is given as a
+    percentage of the shell effect.  Curves correspond to the different
+    tesseroid models and computation grids shown in Table~1.  The
+    horizontal solid black line marks the established error threshold of
+    $0.1\%$.  A value of $D=0$ means that no divisions are made.
+
+    \item
+    Difference between the computed $g_{zz}$ for the spherical shell and
+    the tesseroid model at different heights. Curves show the maximum
+    difference as a percentage of the shell value.  The horizontal solid
+    black line marks the established error threshold of $0.1\%$.  A value
+    of $D=0$ means that no divisions are made.
+
+\end{enumerate}
+
+\newpage
+\section{Table captions}
+
+\begin{enumerate}
+    \item
+    Parameters of the numerical experiments to quantify the accuracy of the
+    numerical integration.  All grids had $10\times10$ regularly spaced
+    computation points at a constant height.  Tesseroids used to discretize the
+    spherical shell had 1 km thickness and the horizontal dimensions shown in
+    the table.
+\end{enumerate}
+
 \end{document}

### 2016-06-09 manuscript-final.tex
Minor corrections from reading the galley proof

@@
 Based on this, \citet{Ku1977} proposed an empirical criterion that
-the distance between point masses should be greater than
+the distance between point masses should not be greater than
 the distance to the computation point.
@@
 Figures for these components can be found in the supplementary material
-(see section "Reproducing the analysis and results").
+(see section "Code for figures and error analysis").
 

### 2016-06-09 manuscript.tex
Had to remove figures and tables for submission
Minor corrections from reading the galley proof

@@
+% A manuscript style format with figures included. This is one you can read.
+
 \pdfoutput=1
-%\documentclass[paper,twocolumn,twoside]{geophysics}
-%\documentclass[manuscript]{geophysics}
-\documentclass[manuscript,endfloat]{geophysics}
+\documentclass[paper,twocolumn]{geophysics}
 \usepackage{natbib}

### 2019-10-03 manuscript.tex
Remove Geophysics formatting from the article
Make a author version for archival.

@@
-% A manuscript style format with figures included. This is one you can read.
-
-\pdfoutput=1
-\documentclass[paper,twocolumn]{geophysics}
-\usepackage{natbib}
+\documentclass[twocolumn]{article}
 \usepackage{graphicx}
 \usepackage{amsmath}
+\usepackage[left=0.7in,right=0.7in,top=1in,bottom=1in]{geometry}
+\usepackage[utf8]{inputenc}
+\usepackage[round]{natbib}
+\usepackage{fixltx2e}
+\usepackage{url}
+\usepackage[pdftex,colorlinks=true]{hyperref}
+\usepackage{fancyhdr}
+
 \newcommand{\mitbf}[1]{\mathbf{{#1}}}
 
-% To leave comments
-\usepackage[colorinlistoftodos]{todonotes}
+\newcommand{\Title}{
+    Tesseroids: forward modeling gravitational fields in spherical coordinates
+}
+\newcommand{\Author}{Uieda, L., Barbosa, V.C.F., Braitenberg, C.}
+\newcommand{\AuthorAffil}{
+    {\large
+        Leonardo Uieda$^{1,2,*}$,
+        Val\'eria C. F. Barbosa$^{2}$,
+        and
+        Carla Braitenberg$^{3}$
+    }
+    \\[0.4cm]
+    {\small $^1$ Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil.}
+    \\
+    {\small $^2$ Observat\'orio Nacional, Rio de Janeiro, Brazil.}
+    \\
+    {\small $^3$ Department of Mathematics and Geosciences, University of Trieste, Trieste, Italy}
+    \\
+    {\small $^*$ e-mail: leouieda@gmail.com}
+}
+\newcommand{\DOI}{10.1190/geo2015-0204.1}
+\newcommand{\DOILink}{\href{https://doi.org/\DOI}{doi.org/\DOI}}
+
+\setlength{\columnsep}{2\columnsep}
+\hypersetup{
+    allcolors=blue,
+    pdftitle={\Title},
+    pdfauthor={\Author},
+}
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{
+    \fontsize{9pt}{12pt}\selectfont
+    \Author{}, 2016. \DOILink{}
+}
+\rhead{\fontsize{9pt}{12pt}\selectfont \thepage}
+\renewcommand{\headrulewidth}{0pt}
 
 \title{
-    Tesseroids: forward modeling gravitational fields in spherical coordinates
-}
-\author{
-    Leonardo Uieda\footnotemark[1]\footnotemark[2],
-    Val\'eria C. F. Barbosa\footnotemark[2],
-    and
-    Carla Braitenberg\footnotemark[3]
 }
@@
 
+\title{\Title}
+\author{\AuthorAffil}
+\date{
+    \normalsize
+    Accepted 2016 July 11. Received in original form 2015 March 30.
+    \\[0.4cm]
+    This is a pre-copyedited, author-produced PDF of an article accepted for
+    publication in \textit{Geophysics} following peer review.
+    The version of record
+    ``\textit{\Author{} (2016). \Title{}. Geophysics. doi:\DOI{}}''
+    is available online at: \DOILink{}.
+}
+
 \maketitle
 
-\righthead{Tesseroids: modeling in spherical coordinates}
-\ms{GEO-2015-0204} % manuscript number
-
-\address{Peer-reviewed code related to this article can be found at
-http://software.seg.org/2016/0004.
-\\
-\footnotemark[1]Universidade do Estado do Rio de Janeiro,
-Rio de Janeiro, Brazil.
-email: leouieda@gmail.com
-\\
-\footnotemark[2]Observat\'orio Nacional,
-Rio de Janeiro, Brazil.
-email: valcris@on.br
-\\
-\footnotemark[3]Department of Mathematics and Geosciences,
-University of Trieste, Trieste, Italy
-}
+\noindent Peer-reviewed code related to this article can be found at
+\href{http://software.seg.org/2016/0004}{software.seg.org/2016/0004} or
+\href{https://github.com/pinga-lab/paper-tesseroids}{github.com/pinga-lab/paper-tesseroids}.
 

### 2019-10-04 manuscript.tex
Finish formatting for author version publication

@@
 \usepackage{amsmath}
-\usepackage[left=0.7in,right=0.7in,top=1in,bottom=1in]{geometry}
+\usepackage[left=1in,right=1in,top=1.3in,bottom=1in]{geometry}
 \usepackage[utf8]{inputenc}
@@
     \fontsize{9pt}{12pt}\selectfont
-    \Author{}, 2016. \DOILink{}
+    Uieda \textit{et al.}, 2016, doi:\DOI{}
 }
@@
 \renewcommand{\headrulewidth}{0pt}
-
-\title{
-}
 
@@
     \\[0.4cm]
-    This is a pre-copyedited, author-produced PDF of an article accepted for
-    publication in \textit{Geophysics} following peer review.
+    This is a pre-copyedited, author-produced PDF of an article
+    published in \textit{Geophysics} following peer review.
     The version of record
@@
     \centering
-    \includegraphics{figs/figure1}
+    \includegraphics[width=\columnwidth]{figs/figure1}
     \caption{
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figs/figure2}
+    \includegraphics[width=0.85\columnwidth]{figs/figure2}
     \caption{
@@
     \centering
-    \includegraphics{figs/figure3}
+    \includegraphics[width=\columnwidth]{figs/figure3}
     \caption{
@@
 
+\input{table1.tex}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/figure4}
+    \caption{
+        The maximum difference between the computed tesseroid and shell effects
+        as a function of the distance-size ratio $D$
+        for (a) the gravitational potential, (b) $g_z$, and (c) $g_{zz}$.
+        The difference is given as a percentage of the shell effect.
+        Curves correspond to the different tesseroid models and computation
+        grids shown in Table~\ref{tbl:experiment}.
+        The horizontal solid black line marks the established error threshold
+        of $0.1\%$.
+        A value of $D=0$ means that no divisions are made.
+    }
+    \label{fig:dist-size-curves}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/figure5}
+    \caption{
+        Difference between the computed $g_{zz}$ for the spherical shell and
+        the tesseroid model at different heights. Curves show the maximum
+        difference as a percentage of the shell value.
+        The horizontal solid black line marks the established error threshold
+        of $0.1\%$.
+        A value of $D=0$ means that no divisions are made.
+    }
+    \label{fig:gzz-with-height}
+\end{figure}
+
 The key controlling point of the adaptive discretization algorithm
@@
 accuracy.
-
-\input{table1.tex}
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/figure4}
-    \caption{
-        The maximum difference between the computed tesseroid and shell effects
-        as a function of the distance-size ratio $D$
-        for (a) the gravitational potential, (b) $g_z$, and (c) $g_{zz}$.
-        The difference is given as a percentage of the shell effect.
-        Curves correspond to the different tesseroid models and computation
-        grids shown in Table~\ref{tbl:experiment}.
-        The horizontal solid black line marks the established error threshold
-        of $0.1\%$.
-        A value of $D=0$ means that no divisions are made.
-    }
-    \label{fig:dist-size-curves}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/figure5}
-    \caption{
-        Difference between the computed $g_{zz}$ for the spherical shell and
-        the tesseroid model at different heights. Curves show the maximum
-        difference as a percentage of the shell value.
-        The horizontal solid black line marks the established error threshold
-        of $0.1\%$.
-        A value of $D=0$ means that no divisions are made.
-    }
-    \label{fig:gzz-with-height}
-\end{figure}
 
